### 2019-08-20 Rmd/jse-2019-ms.Rmd
added Rmd for manuscript

---
title: 'Making Sense of Tweets: Student research projects in data science'
author: "Frederick J. Boehm and Bret M. Hanlon"
date: "8/20/2019"
output: html_document
---

### 2019-09-08 Rmd/jse-2019-ms.Rmd
updated bib

@@
 date: "8/20/2019"
+lastmod: "`r lubridate::now()`"
+bibliography: jse.bib
 output: html_document
 ---
+
+## Introduction
+
+## Methods
+
+- how did we/they conceive of projects?
+
+We mentored two data science researchers in the last year of undergraduate study in statistics at University of Wisconsin-Madison.
+Both students developed novel research questions, collected observational data, analyzed data, and interpreted results.
+
+- how did we balance our input with that of the students?
+- What data did we use?
+- How did we collect data? (which packages, code to sample twitter). How did Jinyu get her stock market data? 
+- What is the sampling scheme - for the 1% sample - from the entirety of Twitter? (Cite the webpage that documents the API)
+
+## Results
+
+## Discussion
+
+## References

### 2019-09-08 Rmd/abstract.Rmd
added abstract

## Abstract

Social media platforms have altered the way that we communicate. The tremendous volumes of streaming data that are available through platforms like Twitter and Facebook enable the investigation of new data science research questions. In many cases, these research questions may serve as engaging and scientifically meaningful studies for those who are new to data science research. 
In seeking to explore social media data science as a pedagogical tool, we guided two undergraduate researchers during the 2015-2016 academic year while they completed mentored research projects.

### 2019-09-08 Rmd/abstract.Rmd
elaborated abstract

@@
 
-Social media platforms have altered the way that we communicate.
+Social media platforms have altered the ways that we communicate.
 The tremendous volumes of streaming data that are available through platforms like Twitter and Facebook enable the investigation of new data science research questions.
-In many cases, these research questions may serve as engaging and scientifically meaningful studies for those who are new to data science research.
-In seeking to explore social media data science as a pedagogical tool, we guided two undergraduate researchers during the 2015-2016 academic year while they completed mentored research projects.
+In many cases, these research questions may motivate engaging, accessible, and scientifically meaningful studies for those who are new to data science research.
+In seeking to explore social media data science as a pedagogical tool, we guided two undergraduate researchers during the 2015-2016 academic year while they completed research projects.
+After discussing the nature of downloaded Twitter data, we encouraged the students to formulate novel research questions that both  draw on the time series structure of tweets and complement students' existing data science interests.
+One student elected to study relationships over time between tweet sentiments and stock market index values.
+The other chose to develop methods for using tweets to detect social media events, like the occurrence of the Super Bowl football game.

### 2019-09-09 Rmd/abstract.Rmd
updated abstr
minor update to abstr

@@
 In many cases, these research questions may motivate engaging, accessible, and scientifically meaningful studies for those who are new to data science research.
-In seeking to explore social media data science as a pedagogical tool, we guided two undergraduate researchers during the 2015-2016 academic year while they completed research projects.
+In seeking to explore social media data science as a pedagogical tool, we guided two undergraduate researchers while they completed research projects.
 After discussing the nature of downloaded Twitter data, we encouraged the students to formulate novel research questions that both draw on the time series structure of tweets and complement students' existing data science interests.
 One student elected to study relationships over time between tweet sentiments and stock market index values.
-The other chose to develop methods for using tweets to detect social media events, like the occurrence of the Super Bowl football game.
+The other chose to develop unsupervised methods to detect social media events, like the occurrence of the Super Bowl football game.
+Both students prepared manuscripts and posters to document their methods and findings.
+We found that research skills, self-confidence in research ability, and professional interest in data science increased for both students.

### 2019-09-17 Rmd/abstract.Rmd
rewrote abstract in response to BH comments

--- 2019-09-08 Rmd/abstract.Rmd
@@
 
-Social media platforms have altered the way that we communicate.
+Social media platforms have altered communication methods.
 The tremendous volumes of streaming data that are available through platforms like Twitter and Facebook enable the investigation of new data science research questions.
-In many cases, these research questions may serve as engaging and scientifically meaningful studies for those who are new to data science research.
-In seeking to explore social media data science as a pedagogical tool, we guided two undergraduate researchers during the 2015-2016 academic year while they completed mentored research projects.
+In many cases, these research questions may motivate engaging, accessible, and scientifically meaningful studies for data science research students.
+In seeking to explore social media data science as a pedagogical tool, we developed a teaching and mentoring framework to guide undergraduate researchers through individualized research projects.
+We demonstrated our approach by guiding two undergraduate researchers while they completed research projects.
+Our framework involves research question formulation, data acquisition, data analysis and visualization, and presentation of results.
+Our two student projects, which serve as case studies for our framework, completed all components of the invididualized research projects.
+We found that research skills, self-confidence in research ability, and professional interest in data science increased for both students.

### 2019-09-25 Rmd/abstract.Rmd
updated with BH edit, a concluding sentence for the abstract.
updates

@@
 In many cases, these research questions may motivate engaging, accessible, and scientifically meaningful studies for data science research students.
-In seeking to explore social media data science as a pedagogical tool, we developed a teaching and mentoring framework to guide undergraduate researchers through individualized research projects.
+To explore social media data science as a pedagogical tool, we developed a teaching and mentoring framework to guide undergraduate researchers through individualized research projects.
 We demonstrated our approach by guiding two undergraduate researchers while they completed research projects.
 Our framework involves research question formulation, data acquisition, data analysis and visualization, and presentation of results.
-Our two student projects, which serve as case studies for our framework, completed all components of the invididualized research projects.
+Our two student projects, which serve as case studies for our framework, completed all components of the individualized research projects.
 We found that research skills, self-confidence in research ability, and professional interest in data science increased for both students.
+Below, we describe our successes, lessons learned, and ideas for others to build similar frameworks for future honors projects.

### 2019-09-25 Rmd/abstract.Rmd
deleted first sentences of abstr

@@
 
-Social media platforms have altered communication methods.
-The tremendous volumes of streaming data that are available through platforms like Twitter and Facebook enable the investigation of new data science research questions.
-In many cases, these research questions may motivate engaging, accessible, and scientifically meaningful studies for data science research students.
 To explore social media data science as a pedagogical tool, we developed a teaching and mentoring framework to guide undergraduate researchers through individualized research projects.

### 2019-10-09 Rmd/abstract.Rmd
parsed text into one file per section.

@@
 
-To explore social media data science as a pedagogical tool, we developed a teaching and mentoring framework to guide undergraduate researchers through individualized research projects.
-We demonstrated our approach by guiding two undergraduate researchers while they completed research projects.
+To explore social media data science as a pedagogical tool, we develop
+a teaching and mentoring framework to guide undergraduate researchers
+through individualized research projects. We demonstrate our approach
+by describing our mentoring experience with two undergraduate researchers.
+These projects involved Twitter data, Facebook data, topic modeling, and sentiment analysis.
 Our framework involves research question formulation, data acquisition, data analysis and visualization, and presentation of results.

### 2019-10-09 Rmd/jse-2019-ms.Rmd
added abstract
fixed title
parsed text into one file per section.

--- 2019-08-20 Rmd/jse-2019-ms.Rmd
@@
 ---
-title: 'Making Sense of Tweets: Student research projects in data science'
+title: 'Making Sense of Tweets: A framework for individualized student research projects'
 author: "Frederick J. Boehm and Bret M. Hanlon"
 date: "8/20/2019"
+lastmod: "`r lubridate::now()`"
+bibliography: jse.bib
 output: html_document
 ---
+
+```{r child = "abstract.Rmd"}
+```
+
+```{r child = "introduction.Rmd"}
+```
+
+```{r child = "methods.Rmd"}
+```
+
+```{r child = "results.Rmd"}
+```
+
+```{r child = "discussion.Rmd"}
+```
+
+## Discussion
+
+## References

### 2019-10-09 Rmd/methods.Rmd
parsed text into one file per section.

## Methods

1. Overview of our framework
  a. student-led, hypothesis-driven research
  b. BH and FB helped with data acquisition
  c. helped with research question formulation
  d. how to describe roles for students v mentors?

Our framework for mentored student research projects involves these components:

1. State the questions and research hypotheses that the two students chose

1. Data description

- how did we/they conceive of projects?

We mentored two data science researchers in the last year of undergraduate study in statistics at University of Wisconsin-Madison. Both students developed novel research questions, collected observational data, analyzed data, and interpreted results.

- how did we balance our input with that of the students?
- What data did we use?
- How did we collect data? (which packages, code to sample twitter). How did Jinyu get her stock market data? 
- What is the sampling scheme - for the 1% sample - from the entirety of Twitter? (Cite the webpage that documents the API)

### 2019-10-26 Rmd/jse-2019-ms.Rmd
updated ms & bib file

@@
 ---
-title: 'Making Sense of Tweets: A framework for individualized student research projects'
+title: 'A framework for mentored data science research projects'
 author: "Frederick J. Boehm and Bret M. Hanlon"
 date: "8/20/2019"
-lastmod: "`r lubridate::now()`"
+params:
+  lastmod: !r lubridate::now()
 bibliography: jse.bib
-output: html_document
+output: bookdown::html_document2
 ---
+
+Last modified: `r params$lastmod`
 
@@
 
-## Discussion
-
 ## References

### 2019-10-26 Rmd/methods.Rmd
updated ms & bib file

@@
 ## Methods
+
+We designed and used a framework for mentored undergraduate data science research projects with big data.
+Our framework draws heavily on ideas in @nolan2010computing.
+Specifically, we incorporate the three key aspects that @nolan2010computing identified:
+
+1. broaden statistical computing to include emerging areas
+1. deepen computational reasoning skills
+1. combine computational topics with data analysis in the practice of statistics
+
+Our mentored research framework begins with brainstorming research ideas based on the student's interests.
+From there, we (mentor and student together) formulate the most promising ideas into scientific hypotheses.
+For the most appealing hypotheses, we encourage the student to translate the scientific question into a statistical question that may be addressed with data analysis.
 
@@
 
-We mentored two data science researchers in the last year of undergraduate study in statistics at University of Wisconsin-Madison.
-Both students developed novel research questions, collected observational data, analyzed data, and interpreted results.
-
 - how did we balance our input with that of the students?

### 2019-10-26 Rmd/notes.Rmd
started a notes file for reading notes

---
title: "Notes on @nolan2010computing"
author: "Frederick J. Boehm"
date: "10/25/2019"
output: html_document
bibliography: jse.bib
---

> The main message of this article is that the digital age is having a profound impact on statistics and the nature of data analysis, and these changes necessitate reevaluation of the training and education practices in statistics.

> Successful statisticians must be facile with the computer, for they are expected to be able to access data from various sources, apply the latest statistical methodologies, and communicate their findings to others in novel ways and via new media.

> When they graduate, what ought our students be able to do computationally, and are we preparing them adequately in this regard?

> Do we provide students the essential skills needed to engage in statistical problem solving and keep abreast of new technologies as they evolve?

> Do our students build the confidence needed to overcome computational challenges to, for example, reliably design and run a synthetic experiment or carry out a comprehensive data analysis?

> Overall, are we doing a good job preparing students who are ready to engage in and succeed at statistical inquiry?

@nolan2010computing identify "three key aspects":

1. broaden statistical computing to include emerging areas
1. deepen computational reasoning skills
1. combine computational topics with data analysis in the practice of statistics

### 2019-11-02 Rmd/discussion.Rmd
updated ms

## Discussion

### Benefits of our framework

Benefits of our framework

### How does the framework relate to ideas in Nolan and Temple Lang 2010??

### Critiques of our framework

Our measures of students' self-confidence in research ability was merely subjective. In future iterations of our framework, we would like to measure systematic and objective outcomes. One strategy for implementing this is to administer a survey, such as \_\_\_, both before and after the mentored research project. We would use validated survey questions that focused on student beliefs about themselves and their skills.

### 2019-11-02 Rmd/methods.Rmd
updated ms

@@
 
-Our mentored research framework begins with brainstorming research ideas based on the student's interests.
-From there, we (mentor and student together) formulate the most promising ideas into scientific hypotheses.
-For the most appealing hypotheses, we encourage the student to translate the scientific question into a statistical question that may be addressed with data analysis.
+Our mentored research framework begins with brainstorming scientific research ideas 
+based on the student's interests. This enables us to craft a project that excites the student. With the results of brainstorming sessions, we (mentor and student together) 
+formulate the most promising ideas into scientific hypotheses.
+
+For the most appealing 
+hypotheses, we encourage the student to translate the scientific question into a 
+statistical question that may be addressed with data.
+
+Examples may help to demonstrate our approach to identifying a statistical research question.
+One of our students had interests in acquiring and using social media posts.
+We helped her in brainstorming ideas for research involving social media like Facebook and Twitter.
+Through this brainstorming, we recognized that she had a parallel interest in financial markets.
+
+study relationships between social media trends and financial market trends. We helped her to brainstorm related research ideas. We then
+
+**justify each step**
+
+**Maybe use/cite Box's science and statistics article from... 1976?**
 

### 2019-11-02 Rmd/results.Rmd
updated ms

## Results

We applied the project framework to our mentoring of two students. Both engaged in 12 months of mentored research during their senior year of undergraduate studies in statistics in 2015 and 2016.

### Student Outcomes

We subjectively assessed student outcomes through conversations in our weekly student research meetings.

One student later enrolled in a statistics graduate program, while the other pursued employment in health care analytics.

### Mentor Outcomes

### Scholarly outcomes & deliverables

### 2019-11-02 Rmd/discussion.Rmd
updated methods

@@
 
-Benefits of our framework
+Benefits of our framework include enhanced
 
@@
 We would use validated survey questions that focused on student beliefs about themselves and their skills.
+
+### Areas for improvement
+
+1. reproducible research best practices
+
+One shortcoming of our initial framework was the relative lack of emphasis on best practices for computational reproducibility.
+This is one area that we would like to rectify in future iterations.
+The University of Wisconsin-Madison has periodically offered a semester course in best practices for computationally reproducible research.
+We especially see version control systems, such as Git and Github, as essential tools for the modern data scientist.

### 2019-11-02 Rmd/methods.Rmd
updated methods

@@
 1. combine computational topics with data analysis in the practice of statistics
+
+### Broaden statistical computing to include emerging areas
+
+Our framework broadens statistical computing by including the emerging areas of
+social media data analysis, sentiment analysis, and topic modeling. 
+Both of our students used Twitter tweets, which we accessed through a Twitter 
+streaming API.
+
+Our computational system for acquiring tweets involved several steps.
+We interacted with the API via the R package `twitteR` [@twitteR].
+We used the free Twitter streaming API that gave us access to approximately one percent stream of all tweets during the specified query time period.
+To ensure that we collected tweets continuously, we used the linux tool `crontab` to execute our R script every five minutes.
+Each execution of the R script performed a single streaming API query, lasting five minutes.
+Twitter's streaming API, at the time of our data collection, enforced rate limits on the frequency and duration of queries.
+
+### deepen computational reasoning skills
+
+Our framework, and the projects developed under it, encourage students to deepen 
+computational reasoning skills in several ways. First, they work with a variety of 
+internet-based data to answer research questions. In the two example cases, our 
+students collected tweets over time and gathered complementary data from other 
+resources, including daily closing prices of stock market indexes.
+
+Second, the students worked with a variety of data structures. 
+The Twitter streaming 
+API returns tweets as JSON files. Because distinct Twitter users may provide 
+different pieces of profile information, there is variability in the structure of 
+each tweet's JSON entry. Students needed to recognize this, and write code that 
+accommodated these variations in tweet data structure.
+
+### combine computational topics with data analysis in the practice of statistics
 

### 2019-11-08 Rmd/methods.Rmd
updated methods text

@@
 
-We designed and used a framework for mentored undergraduate data science research projects with big data.
-Our framework draws heavily on ideas in @nolan2010computing.
+We designed and used a framework for mentored undergraduate data science research 
+projects with big data. Our framework draws heavily on ideas in @nolan2010computing. 
 Specifically, we incorporate the three key aspects that @nolan2010computing identified:
@@
 
-Our computational system for acquiring tweets involved several steps.
-We interacted with the API via the R package `twitteR` [@twitteR].
-We used the free Twitter streaming API that gave us access to approximately one percent stream of all tweets during the specified query time period.
-To ensure that we collected tweets continuously, we used the linux tool `crontab` to execute our R script every five minutes.
-Each execution of the R script performed a single streaming API query, lasting five minutes.
-Twitter's streaming API, at the time of our data collection, enforced rate limits on the frequency and duration of queries.
+Our computational system for acquiring tweets involved several steps. We interacted 
+with the API via the R package `twitteR` [@twitteR]. We used the free Twitter streaming 
+API that gave us access to approximately one percent stream of all tweets during the 
+specified query time period. To ensure that we collected tweets continuously, we used 
+the linux tool `crontab` to execute our R script every five minutes. Each execution of 
+the R script performed a single streaming API query, lasting five minutes. Twitter's 
+streaming API, at the time of our data collection, enforced rate limits on the 
+frequency and duration of queries. With the above settings, we continuously collected 
+tweets.
 
@@
 students collected tweets over time and gathered complementary data from other 
-resources, including daily closing prices of stock market indexes.
+resources, including daily closing prices of stock market indexes. This gave students 
+opportunities to think creatively about what data to acquire and how to use multiple 
+data sources in a single cohesive project.
 
@@
 The Twitter streaming 
-API returns tweets as JSON files. Because distinct Twitter users may provide 
+API returns tweets as JSON (Javascript Object Notation, http://json.org). Because 
+distinct Twitter users may provide 
 different pieces of profile information, there is variability in the structure of 
-each tweet's JSON entry. Students needed to recognize this, and write code that 
+each tweet's JSON. Additionally, tweet metadata fields may appear in any order (https://developer.twitter.com/en/docs/tutorials/consuming-streaming-data). 
+Students needed to recognize this and write code that 
 accommodated these variations in tweet data structure.
+
+Students wrote code to parse and organize tweet JSONs. Each tweet's JSON included required fields, and, possibly, some optional fields. The required fields are:
+
+**see https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object**
+
+The evolving nature of JSON tweet structure 
+(https://developer.twitter.com/en/docs/tweets/data-dictionary/guides/tweet-timeline) 
+required students to write flexible code that could incorporate newly introduced 
+or deprecated metadata.
 
@@
 
+computational topics: LDA & topic modeling; time series analysis; sentiment analysis.
+
+data analysis and practice of statistics:
+
 Our mentored research framework begins with brainstorming scientific research ideas 
-based on the student's interests. This enables us to craft a project that excites the student. With the results of brainstorming sessions, we (mentor and student together) 
+based on the student's interests. This enables us to craft a project that excites the 
+student. With the results of brainstorming sessions, we (mentor and student together) 
 formulate the most promising ideas into scientific hypotheses.
@@
 
-Examples may help to demonstrate our approach to identifying a statistical research question.
-One of our students had interests in acquiring and using social media posts.
-We helped her in brainstorming ideas for research involving social media like Facebook and Twitter.
-Through this brainstorming, we recognized that she had a parallel interest in financial markets.
+Examples may help to demonstrate our approach to identifying a statistical research 
+question. One of our students had interests in acquiring and using social media posts. 
+We helped her in brainstorming ideas for research involving social media like Facebook 
+and Twitter. Through this brainstorming, we recognized that she had a parallel interest 
+in financial markets.
 
-study relationships between social media trends and financial market trends. We helped her to brainstorm related research ideas. We then
+study relationships between social media trends and financial market trends. We helped 
+her to brainstorm related research ideas. We then
 

### 2019-11-11 Rmd/introduction.Rmd
updated ms

## Introduction

- big data
- new areas of application of statistical methods - social media
- need for mentoring 
- need for students to work with real data
- motivate statistical analysis by real world research questions

### 2019-11-11 Rmd/methods.Rmd
updated ms

@@
 social media data analysis, sentiment analysis, and topic modeling. 
-Both of our students used Twitter tweets, which we accessed through a Twitter 
+Both students used Twitter tweets, which we accessed through a Twitter 
 streaming API.
@@
 
-Students wrote code to parse and organize tweet JSONs. Each tweet's JSON included required fields, and, possibly, some optional fields. The required fields are:
+Students wrote code to parse and organize tweet JSONs. They organized their R code into a package, https://github.com/rturn/parseTweetFiles. 
+Each tweet's JSON included required fields, and, possibly, some optional fields. 
+The required fields are:
 
@@
 computational topics: LDA & topic modeling; time series analysis; sentiment analysis.
+
+Both mentored students combined computing with data analysis in the practice of statistics.
+They used a combination of latent dirichlet allocation topic modeling, sentiment analysis, and time series analysis to reach conclusions about real world data.
+
+Both drew heavily on our collection of tweets. One student examined Standard and Poor's 500 index daily 
+closing prices over time. She also analyzed sentiments from each day's stock market-related tweets 
+to look for relationships between tweet sentiment and stock market prices.
+
+Our other student focused on developing detection methods for social media events through 
+topic modeling of tweets at different time periods. As a proof of principle, he fitted topic 
+models to collections of tweets preceding, during, and following the National Football 
+League's Super Bowl game. He hypothesized that topics would evolve over time, with football-related tweets appearing during the football game and disappearing soon after conclusion of the game.
 

### 2019-11-21 Rmd/jse-2019-ms.Rmd
updated ms

@@
 ---
-title: 'A framework for mentored data science research projects'
+title: 'A framework for mentored data science research'
 author: "Frederick J. Boehm and Bret M. Hanlon"
@@
 bibliography: jse.bib
-output: bookdown::html_document2
+output: 
+  - github_document
+  - pdf_document
 ---

### 2019-11-21 Rmd/methods.Rmd
recompiled pdf and md
updated ms

@@
 1. combine computational topics with data analysis in the practice of statistics
+
+Additionally, our projects gave students opportunities to develop and to practice skills in reproducible research.
+Given the growing imperative to document and share code, we feel that this skill set equals in importance the three points above.
+
+4. develop and practice skills in reproducible research
+
+Below, we describe how our framework enabled students to achieve in the four items listed above.
 
@@
 the linux tool `crontab` to execute our R script every five minutes. Each execution of 
-the R script performed a single streaming API query, lasting five minutes. Twitter's 
+the R script performed a single streaming API query for five minutes. Twitter's 
 streaming API, at the time of our data collection, enforced rate limits on the 
@@
 
-Our framework, and the projects developed under it, encourage students to deepen 
+Our framework encourages students to deepen 
 computational reasoning skills in several ways. First, they work with a variety of 
@@
 different pieces of profile information, there is variability in the structure of 
-each tweet's JSON. Additionally, tweet metadata fields may appear in any order (https://developer.twitter.com/en/docs/tutorials/consuming-streaming-data). 
-Students needed to recognize this and write code that 
+each tweet's JSON. 
+Additionally, tweet metadata fields may appear in any order (https://developer.twitter.com/en/docs/tutorials/consuming-streaming-data). 
+Students needed to recognize this and to write code that 
 accommodated these variations in tweet data structure.
@@
 **see https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object**
+**insert a 2-column table of fields here**
 

### 2019-11-21 Rmd/outline.Rmd
updated ms

---
title: "outline"
author: "Frederick J. Boehm"
date: "11/20/2019"
output: github_document
---

## What is the structure of Nolan and Temple Lang 2010??

They give one major section (2, 3, and 4) to each of the three bullet points that I highlight in my methods section.

## What are the important ideas that I need to convey in each section??

**Intro**

Mimic, maybe, the DTL2010 article in organization of the intro. have a subsection on *four* key components (vs DTL's three key components). These are the three of DTL plus reproducible research practices.

Maybe also try to write a paragraph about relevant background for each of me and Bret. Note that DTL puts a subsection on backgrounds in their intro.

**Methods**

Here, say more about our contributions - what we did and how we did it. Don't discuss DTL in depth in the Methods - ie, move DTL discussion (that's currently in my methods section) to intro.

Here is the place to discuss our framework and its implementation for two students JX and RT.

**Results**

### 2019-11-27 Rmd/introduction.Rmd
added citation
added intro text; added more bib citations

@@
 - need for students to work with real data
-- motivate statistical analysis by real world research questions
+- motivate statistical analysis and statistics research by real world scientific research questions (Box 1976)
+
+@nolan2010computing
+
+The need to analyze unprecedentedly large volumes of information combined with 
+the development of faster and more powerful computers has fueled advances in 
+data science methods for big data. We present below a framework for mentored 
+undergraduate data science research and our findings from its initial 
+implementation for two statistics students. We conclude with lessons learned 
+and suggestions for those who wish to reproduce and refine our framework.
+
+Social media data, including tweets from Twitter and posts from Facebook, are 
+available through website application product interfaces (APIs). Twitter 
+shares, via a streaming API, a sample of approximately one percent of all 
+tweets during an API query time period. Researchers have studied tweets for a 
+variety of purposes, including 
+inference of relationships and social networks among users [@lin2011joint]; determination of 
+authorship of specific tweets when multiple persons use a single account [@drob]; and 
+study of rhetoric in recruiting political supporters [@pelled2018little,@wells2016trump].

### 2019-11-28 Rmd/discussion.Rmd
updated methods text
updated intro

@@
 ### Benefits of our framework
+
+1. build self-confidence, self-assessed proficiency in data analysis and statistics
+
+1. greater interest in quantitative careers and/or grad school in quant fields
+
+1. Working with real data to solve real scientific questions
+
+1. Learn by example interplay between data science (statistics) and science research
 
@@
 
-Our measures of students' self-confidence in research ability was merely subjective.
-In future iterations of our framework, we would like to measure systematic and objective outcomes.
-One strategy for implementing this is to administer a survey, such as \_\_\_, both before and after the mentored research project.
-We would use validated survey questions that focused on student beliefs about themselves and their skills.
+Our measures of students' self-confidence in research ability was merely subjective. In 
+future iterations of our framework, we would like to measure systematic and objective 
+outcomes. One strategy for implementing this is to administer a survey, such as \_\_\_, 
+both before and after the mentored research project. We would use validated survey 
+questions that focused on student beliefs about themselves, their skills, and their 
+future careers. **maybe list resources for such survey questions**
 
@@
 We especially see version control systems, such as Git and Github, as essential tools for the modern data scientist.
+
+1. assessment of attitudes (pre and post survey??)
+
+1. survey questions might include those from @vance2017eight. I have the survey as a pdf in mendeley. It is part of the supp info for @vance2017eight.

### 2019-11-28 Rmd/introduction.Rmd
updated intro

@@
 - new areas of application of statistical methods - social media
-- need for mentoring 
 - need for students to work with real data
 - motivate statistical analysis and statistics research by real world scientific research questions (Box 1976)
-
-@nolan2010computing
+- need for mentoring
 
@@
 shares, via a streaming API, a sample of approximately one percent of all 
-tweets during an API query time period. Researchers have studied tweets for a 
+tweets during an API query time period [@tweet_stream]. 
+Researchers have studied tweets for a 
 variety of purposes, including 
-inference of relationships and social networks among users [@lin2011joint]; determination of 
-authorship of specific tweets when multiple persons use a single account [@drob]; and 
+inference of relationships and social networks among users [@lin2011joint]; 
+determination of 
+authorship of specific tweets when multiple persons share a single account [@drob]; and 
 study of rhetoric in recruiting political supporters [@pelled2018little,@wells2016trump]. 
+Recognizing the potential utility of tweets for data science research, we created a 
+collection of tweets over time by repeated querying of the Twitter streaming API.
+
+@nolan2010computing fervently argue for students to work with real data. 
+Working with real data allows them to develop skill not only in statistical analysis,
+but also in data transfer from online sources, in data storage, and in using data 
+from multiple file formats. In the case of Twitter data, tweets are stored in
+Javascript Object Notation (JSON) [@tweet_json, @json].
+
+A key component of our mentoring framework is the emphasis on using real data to answer real 
+scientific questions. We believe that this process develops the problem-solving skills that 
+students will need in their future careers in data science.
+
+We encouraged the student to articulate a scientific research question, translate that question into quantitative and statistical terms, determine which data could be used to address the question, acquire the data, analyze data, visualize results, and communicate what they learned.
+
+**Need to explain above paragraph with our two student examples. What types of questions did they articulate? How did they translate to quant terms? How did they determine data availability? etc.**

### 2019-11-28 Rmd/jse-2019-ms.Rmd
updated ms Rmd and bib file
updated intro

@@
 ---
-title: 'A framework for mentored data science research'
+title: 'A framework for undergraduate data science research'
 author: "Frederick J. Boehm and Bret M. Hanlon"
@@
   - github_document
-  - pdf_document
+  - bookdown::pdf_document2
+  - bookdown::html_document2
 ---

### 2019-11-28 Rmd/outline.Rmd
updated ms Rmd and bib file
updated intro

@@
 date: "11/20/2019"
-output: github_document
+output: bookdown::html_document2
 ---
@@
 Mimic, maybe, the DTL2010 article in organization of the intro.
-have a subsection on *four* key components (vs DTL's three key components).
-These are the three of DTL plus reproducible research practices.
+
+~~have a subsection on *four* key components (vs DTL's three key components). These are the three of DTL plus reproducible research practices.~~
 

### 2019-11-29 Rmd/introduction.Rmd
udpated intro & README

@@
 
-@nolan2010computing fervently argue for students to work with real data. 
+@nolan2010computing argue for students to work with real data. 
 Working with real data allows them to develop skill not only in statistical analysis,
@@
 
+Mentoring in the work place and in higher education can have many benefits, including [@higgins2001reconceptualizing].
 A key component of our mentoring framework is the emphasis on using real data to answer real 
@@
 students will need in their future careers in data science. 
+We encouraged the student to articulate a scientific research question, translate that question 
+into quantitative and statistical terms, determine which data could be used to address the 
+question, acquire the data, analyze data, visualize results, and communicate what they learned.
 
-We encouraged the student to articulate a scientific research question, translate that question into quantitative and statistical terms, determine which data could be used to address the question, acquire the data, analyze data, visualize results, and communicate what they learned.
+We provide guidance regarding selection of
 
 **Need to explain above paragraph with our two student examples. What types of questions did they  articulate? How did they translate to quant terms? How did they determine data availability? etc.**
+
+## Our backgrounds
+
+During the time when we first implemented our framework, we served as early-career instructors in 
+the statistics department at the University of Wisconsin-Madison. One of us (Hanlon) had prior 
+experience in mentoring undergraduates, while the other (Boehm) had none. Our initial 
+conceptualization of mentoring drew heavily on ideas we first encountered in professional development 
+courses, including the Delta Program's Entering Mentoring class (https://delta.wisc.edu). 
+Professor Erik Nordheim heavily 
+influenced our approach to and philosophy of teaching statistics. 
+We studied with Professor Nordheim early in our teaching 
+careers, and his guidance continues to influence our teaching practices.
+
+We both have experience in teaching undergraduate introductory statistics courses with enrollments over 
+100 students. Through our interactions with students in these classes, we've come to value not only 
+the ideas in a traditional introductory course, but also the need to prepare students with the essential skills needed for success in data science.
+@nolan2010computing summarizes these skill sets in the following three ideas:
+
+1. broaden statistical computing to include emerging areas
+1. deepen computational reasoning skills
+1. combine computational topics with data analysis in the practice of statistics
+
+To these three praiseworthy goals, we add a fourth:
+
+4. develop skills in reproducible research to promote the practice of open science
+
+We see the fourth goal as an equal with the first three from @nolan2010computing. 
+Data scientists are uniquely positioned to promote open science practices, including the free sharing of 
+data, code, and instructions for their use. The recently recognized need for science to be more transparent and more reproducible elevate this goal to the level of the first three.
+
+Below, we detail our methods for creating a reproducible framework for undergraduate data science 
+research. We describe our results before concluding with lessons learned, 
+things we could have done differently, and recommendations for future 
+mentors who may use and extend our framework.

### 2019-11-29 Rmd/methods.Rmd
udpated intro & README

@@
 Additionally, our projects gave students opportunities to develop and to practice skills in reproducible research.
-Given the growing imperative to document and share code, we feel that this skill set equals in importance the three points above.
+Given the growing imperative to document and share code to promote open science, we feel that this skill set equals in importance the three points above.
 
-4. develop and practice skills in reproducible research
+4. develop and practice skills in reproducible research to promote open science
 
-Below, we describe how our framework enabled students to achieve in the four items listed above.
+Below, we describe how our framework enabled students to achieve competence in the four areas listed above.
 
@@
 
-data analysis and practice of statistics:
+### data analysis and practice of statistics
 

### 2019-11-30 Rmd/methods.Rmd
fixed headers

@@
-## Methods
+# Methods
 
@@
 
-### Broaden statistical computing to include emerging areas
+## Broaden statistical computing to include emerging areas
 
@@
 
-### deepen computational reasoning skills
+## Deepen computational reasoning skills
 
@@
 
-### combine computational topics with data analysis in the practice of statistics
+## Combine computational topics with data analysis in the practice of statistics
 
@@
 
-### data analysis and practice of statistics
+## Implementation of the framework
 

### 2019-11-30 Rmd/introduction.Rmd
fixed headers
updated methods

@@
-## Introduction
+# Introduction
 
@@
 from multiple file formats. In the case of Twitter data, tweets are stored in
-Javascript Object Notation (JSON) [@tweet_json, @json].
+Javascript Object Notation (JSON) [@tweet_json,@json].
 
-Mentoring in the work place and in higher education can have many benefits, including [@higgins2001reconceptualizing].
-A key component of our mentoring framework is the emphasis on using real data to answer real 
-scientific questions. We believe that this process develops the problem-solving skills that 
+Mentoring in the work place and in higher education can have many 
+benefits, including improving students' development as thinkers and scholars, confidence in their 
+own abilities, integration into the campus community, and interest in graduate training 
+[@baker2010beyond,@higgins2001reconceptualizing].
+A key component of our data science mentoring framework is the emphasis on using real data to answer real 
+scientific questions. We believe that this process develops problem-solving skills that 
 students will need in their future careers in data science. 
@@
 the statistics department at the University of Wisconsin-Madison. One of us (Hanlon) had prior 
-experience in mentoring undergraduates, while the other (Boehm) had none. Our initial 
-conceptualization of mentoring drew heavily on ideas we first encountered in professional development 
-courses, including the Delta Program's Entering Mentoring class (https://delta.wisc.edu). 
+experience in mentoring students, while the other (Boehm) had none. Our initial 
+conceptualization of mentoring drew heavily on ideas we first encountered in professional 
+development 
+activities, including the Delta Program's mentoring class (https://delta.wisc.edu) and @handelsman2005entering. 
 Professor Erik Nordheim heavily 
@@
 We studied with Professor Nordheim early in our teaching 
-careers, and his guidance continues to influence our teaching practices.
+careers, and his emphasis on active learning continues to influence our teaching practices.
 
-We both have experience in teaching undergraduate introductory statistics courses with enrollments over 
-100 students. Through our interactions with students in these classes, we've come to value not only 
-the ideas in a traditional introductory course, but also the need to prepare students with the essential skills needed for success in data science.
-@nolan2010computing summarizes these skill sets in the following three ideas:
+We both have experience in teaching undergraduate introductory 
+statistics courses with enrollments over 
+100 students. Through our interactions with students in these 
+classes, we've grown to value not only 
+the ideas in a traditional introductory course, but also the need 
+to prepare students with the essential skills needed for success 
+in data science. @nolan2010computing summarizes these skill sets 
+in the following three goals:
 
@@
 
-4. develop skills in reproducible research to promote the practice of open science
+4. develop skills in reproducible research to promote open science practices
 
 We see the fourth goal as an equal with the first three from @nolan2010computing. 
-Data scientists are uniquely positioned to promote open science practices, including the free sharing of 
-data, code, and instructions for their use. The recently recognized need for science to be more transparent and more reproducible elevate this goal to the level of the first three.
+Data scientists are uniquely positioned to promote open science 
+practices, including the free sharing of 
+data, code, and instructions for their use. 
+The need for science to be more transparent and more reproducible
+elevate this goal to the level of the first three.
 
-Below, we detail our methods for creating a reproducible framework for undergraduate data science 
+Below, we detail our methods for creating a reproducible framework
+for undergraduate data science 
 research. We describe our results before concluding with lessons learned, 

### 2019-11-30 Rmd/methods.Rmd
updated methods

@@
 We designed and used a framework for mentored undergraduate data science research 
-projects with big data. Our framework draws heavily on ideas in @nolan2010computing. 
-Specifically, we incorporate the three key aspects that @nolan2010computing identified:
+projects with big data. Below, we describe our initial implementation of the framework before relating 
+it to three major ideas from @nolan2010computing.
+
+## Framework implementation
+
+Our mentored research framework begins with brainstorming scientific research ideas 
+based on the student's interests. This enables us to craft a project that excites the 
+student. With the results of brainstorming sessions, we (mentor and student together) 
+formulate the most promising ideas into scientific hypotheses.
+
+For the most appealing 
+hypotheses, we encourage the student to translate the scientific question into a 
+statistical question that may be addressed with data.
+
+Examples may help to demonstrate our approach to identifying a statistical research 
+question. One of our students had interests in acquiring and using social media posts. 
+We helped her in brainstorming ideas for research involving social media like Facebook 
+and Twitter. Through this brainstorming, we recognized that she had a parallel interest 
+in financial markets.
+
+study relationships between social media trends and financial market trends. We helped 
+her to brainstorm related research ideas. We then
+
+## Relating to three ideas from @nolan2010computing
+
+We incorporate three key aspects that @nolan2010computing identified:
 
@@
 
-Additionally, our projects gave students opportunities to develop and to practice skills in reproducible research.
-Given the growing imperative to document and share code to promote open science, we feel that this skill set equals in importance the three points above.
+Additionally, our projects gave students opportunities to develop and to practice skills in reproducible 
+research. Given the growing imperative to document and share code to promote open science, we feel that 
+this skill set equals in importance the three points above.
 
@@
 
-Below, we describe how our framework enabled students to achieve competence in the four areas listed above.
+Below, we describe how our framework enabled students to achieve competence in the four areas listed 
+above.
 
@@
 The Twitter streaming 
-API returns tweets as JSON (Javascript Object Notation, http://json.org). Because 
+API returns tweets as JSON (Javascript Object Notation). Because 
 distinct Twitter users may provide 
@@
 accommodated these variations in tweet data structure.
-
-Students wrote code to parse and organize tweet JSONs. They organized their R code into a package, https://github.com/rturn/parseTweetFiles. 
-Each tweet's JSON included required fields, and, possibly, some optional fields. 
-The required fields are:
-
-**see https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object**
-**insert a 2-column table of fields here**
-
+Additional variability in tweet structure arose due to changes in the API. 
 The evolving nature of JSON tweet structure 
@@
 
+Students wrote R code to parse and organize tweet JSON []. They organized their R code into a package, and shared it on Github (https://github.com/rturn/parseTweetFiles). 
+Each tweet's JSON included required fields, and, possibly, some optional fields.
+
 ## Combine computational topics with data analysis in the practice of statistics
@@
 
-Both mentored students combined computing with data analysis in the practice of statistics.
-They used a combination of latent dirichlet allocation topic modeling, sentiment analysis, and time series analysis to reach conclusions about real world data.
+Both mentored students combined computing with data analysis in 
+the practice of statistics. 
+They used a combination of latent dirichlet allocation topic 
+modeling, sentiment analysis, and time series analysis to reach 
+conclusions about real world data.
 
-Both drew heavily on our collection of tweets. One student examined Standard and Poor's 500 index daily 
-closing prices over time. She also analyzed sentiments from each day's stock market-related tweets 
+Both drew heavily on the collection of tweets. One student examined 
+Standard and Poor's 500 index daily 
+closing prices over time. She also analyzed sentiments 
+from each day's stock market-related tweets 
 to look for relationships between tweet sentiment and stock market prices.
@@
 models to collections of tweets preceding, during, and following the National Football 
-League's Super Bowl game. He hypothesized that topics would evolve over time, with football-related tweets appearing during the football game and disappearing soon after conclusion of the game.
+League's Super Bowl game. He hypothesized that topics would evolve 
+over time, with football-related tweets appearing during the football 
+game and disappearing soon after conclusion of the game.
 
-## Implementation of the framework
+Both students analyzed tweets as texts. This first required them to write code to parse the JSON
+that the API returns. Once they had isolated the tweet text from its metadata, they parsed the 
+tweet text into words for use in sentiment analysis and topic modeling. 
+For the stock market project, they analyzed only those tweets that contained finance-related 
+keywords. Sentiment analysis involved comparisons of tweet words to a dictionary that mapped words to 
+sentiments. This yielded a net sentiment score for each tweet. 
+They then treated tweet sentiment scores as a time series and compared them with 
+daily stock market index closing prices.
 
-Our mentored research framework begins with brainstorming scientific research ideas 
-based on the student's interests. This enables us to craft a project that excites the 
-student. With the results of brainstorming sessions, we (mentor and student together) 
-formulate the most promising ideas into scientific hypotheses.
-
-For the most appealing 
-hypotheses, we encourage the student to translate the scientific question into a 
-statistical question that may be addressed with data.
-
-Examples may help to demonstrate our approach to identifying a statistical research 
-question. One of our students had interests in acquiring and using social media posts. 
-We helped her in brainstorming ideas for research involving social media like Facebook 
-and Twitter. Through this brainstorming, we recognized that she had a parallel interest 
-in financial markets.
-
-study relationships between social media trends and financial market trends. We helped 
-her to brainstorm related research ideas. We then
+The second student project involved latent Dirichlet allocation modeling of
+tweet words at distinct time points to detect social media events [@blei2003latent]. 
+Latent Dirichlet allocation is a bayesian nonparametric method for modeling text corpora as the result 
+of words chosen from topics.
 

### 2019-12-02 Rmd/abstract.Rmd
fixed headers
updates

--- 2019-09-25 Rmd/abstract.Rmd
@@
-## Abstract
+# Abstract
 
-To explore social media data science as a pedagogical tool, we developed a teaching and mentoring framework to guide undergraduate researchers through individualized research projects.
-We demonstrated our approach by guiding two undergraduate researchers while they completed research projects.
-Our framework involves research question formulation, data acquisition, data analysis and visualization, and presentation of results.
-Our two student projects, which serve as case studies for our framework, completed all components of the individualized research projects.
-We found that research skills, self-confidence in research ability, and professional interest in data science increased for both students.
-Below, we describe our successes, lessons learned, and ideas for others to build similar frameworks for future honors projects.
+We develop
+a mentoring framework to guide undergraduate researchers
+through individualized research projects in social media data science. 
+Our framework involves research question formulation, data acquisition, data 
+analysis and visualization, and presentation and communication of results. Our two honors students, 
+whose projects serve as case studies for our framework, completed all components of the 
+individualized research projects. We found that data science research skills, self-confidence in 
+research ability, and professional interest in data science increased for both 
+students. We describe our successes, lessons learned, and ideas for others to 
+build similar frameworks.

### 2019-12-02 Rmd/methods.Rmd
updates

@@
 based on the student's interests. This enables us to craft a project that excites the 
-student. With the results of brainstorming sessions, we (mentor and student together) 
+student. With the results of brainstorming sessions, we (mentors and student together) 
 formulate the most promising ideas into scientific hypotheses.
 
-For the most appealing 
+For the most appealing scientific 
 hypotheses, we encourage the student to translate the scientific question into a 
-statistical question that may be addressed with data.
+statistical question that may be addressed with data. This is a crucial step in 
+data science research question formulation. Skill in translating in both directions between 
+scientific and statistical questions is a key communication skill that data science 
+researchers offer.
 
@@
 question. One of our students had interests in acquiring and using social media posts. 
-We helped her in brainstorming ideas for research involving social media like Facebook 
+We helped her in brainstorming ideas for research involving social media sources like Facebook 
 and Twitter. Through this brainstorming, we recognized that she had a parallel interest 
-in financial markets.
+in financial markets. Our student hypothesized that sentiment analysis of finance-related tweets 
+might reflect trends in financial market index prices. On days when the market index
+prices increase, sentiment analysis of finance-related tweets might reveal more use of 
+positive words, while days with decreasing prices might have more negative words 
+in finance-related tweets.
 
-study relationships between social media trends and financial market trends. We helped 
-her to brainstorm related research ideas. We then
+A second student wanted to study tweets over time and entertainment events that 
+garner lots of attention in social media. We encouraged this student to develop a 
+strategy for event detection from tweets over time. The rationale is that 
+a big entertainment event, such as the National Football League's Super Bowl game, 
+might generate enough tweets that Super Bowl-related words would appear with high 
+weights in results from latent Dirichlet allocation modeling of collections of tweets 
+at distinct time points. We reasoned that Super Bowl-related topics might appear during 
+the Super Bowl and vanish soon after the game's conclusion.
+
+The brainstorming sessions thus led to one set of questions for each student.
 

### 2019-12-03 Rmd/methods.Rmd
updated methods

@@
 
+We also incorporated data availability into our question formulation. We limited questions to those 
+that could be studied with publicly available data. This practice also enabled reproducibility of
+our analyses, since students could share the URL from which they accessed data.
+
+After identifying research questions and publicly available data, 
+the next step is to decide on informative data visualizations and quantitative analyses. Because 
+both projects primarily involved exploratory analyses of times series, 
+we encouraged students to think about
+visualizations that might reveal relationships over time.
+
+### Examples
+
 Examples may help to demonstrate our approach to identifying a statistical research 
@@
 the Super Bowl and vanish soon after the game's conclusion.
-
-The brainstorming sessions thus led to one set of questions for each student.
 

### 2019-12-04 Rmd/methods.Rmd
upd

@@
 
-We designed and used a framework for mentored undergraduate data science research 
-projects with big data. Below, we describe our initial implementation of the framework before relating 
-it to three major ideas from @nolan2010computing.
+We used backward design principles to create and implement a 
+framework for mentored undergraduate data science research 
+projects with big data. Below, we describe our use of 
+backward design and ideas from @nolan2010computing.
+We then detail our framework and its initial implementation.
+
+## Framework design
+
+We started with the backward design principles from @wiggins2005understanding. We prioritized
 
@@
 
-We also incorporated data availability into our question formulation. We limited questions to those 
-that could be studied with publicly available data. This practice also enabled reproducibility of
+We also incorporated data availability into our question formulation. 
+We limited questions to those 
+that could be studied with publicly available data. 
+This practice also enabled reproducibility of
 our analyses, since students could share the URL from which they accessed data.
@@
 After identifying research questions and publicly available data, 
-the next step is to decide on informative data visualizations and quantitative analyses. Because 
+the next step is to decide on informative data visualizations and quantitative analyses. 
+Because 
 both projects primarily involved exploratory analyses of times series, 

### 2019-12-04 Rmd/methods.Rmd
started adding stuff for backward design

@@
 
-We started with the backward design principles from @wiggins2005understanding. We prioritized
+We used the backward design principles from @wiggins2005understanding 
+to create a framework for undergraduate data science research. 
+Following @wiggins2005understanding, we identified desired results, 
+determined acceptable evidence, and planned learning experiences.
+
+In establishing framework priorities, we categorized all terms from Figure 1 of @nolan2010computing into 3 categories: 
+1. worth being familiar with 
+1. important to know and do
+1. enduring understanding
 

### 2019-12-04 Rmd/jse-2019-ms.Rmd
udpated intro & README
updates
upd
updated methods

--- 2019-11-21 Rmd/jse-2019-ms.Rmd
@@
 ---
-title: 'A framework for mentored data science research'
+title: 'Backward designing a reproducible framework for undergraduate data science research'
 author: "Frederick J. Boehm and Bret M. Hanlon"
@@
 output: 
-  - github_document
-  - pdf_document
+  - bookdown::pdf_document2
+  - bookdown::html_document2
 ---

### 2019-12-04 Rmd/methods.Rmd
updated methods

@@
 projects with big data. Below, we describe our use of 
-backward design and ideas from @nolan2010computing.
+backward design with ideas from @nolan2010computing.
 We then detail our framework and its initial implementation.
@@
 
-We used the backward design principles from @wiggins2005understanding 
+We used backward design principles from @wiggins2005understanding 
 to create a framework for undergraduate data science research. 
@@
 
-In establishing framework priorities, we categorized all terms from Figure 1 of @nolan2010computing into 3 categories: 
+Before we identified desired results, we prioritized topics from 
+@nolan2010computing. Specifically, we assigned all terms from Figure 1 
+of @nolan2010computing into one of three categories:
+
 1. worth being familiar with 
@@
 1. enduring understanding
+
+```{r table1, echo = FALSE}
+library(magrittr)
+circles <- tibble::tibble(Class = 1:3, Circle = c("Worth being familiar with", "Important to know and do", "Enduring understanding"))
+readr::read_csv("../data/nolan-figure1-circles.csv") %>%
+  dplyr::left_join(circles) %>%
+  dplyr::arrange(desc(Class)) %>%
+  dplyr::select(- Class) %>%
+  gt::gt() %>%
+  gt::tab_header(title = "Prioritizing Key Terms from Figure 1 of @nolan2010computing", 
+                 subtitle ="xxx" ) %>%
+  gt::as_latex()
+```
 

### 2019-12-04 Rmd/discussion.Rmd
fixed headers
minor updates to discussion

@@
-## Discussion
+# Discussion
 
-### Benefits of our framework
+## Benefits of our framework
 
-1. build self-confidence, self-assessed proficiency in data analysis and statistics
+1. build self-confidence, proficiency in data analysis and statistics
 
@@
 
-### How does the framework relate to ideas in Nolan and Temple Lang 2010??
+## How does the framework relate to ideas in Nolan and Temple Lang 2010??
 
-### Critiques of our framework
+## Critiques of our framework
 
@@
 
-### Areas for improvement
+## Areas for improvement
 
@@
 
-One shortcoming of our initial framework was the relative lack of emphasis on best practices for computational reproducibility.
-This is one area that we would like to rectify in future iterations.
-The University of Wisconsin-Madison has periodically offered a semester course in best practices for computationally reproducible research.
-We especially see version control systems, such as Git and Github, as essential tools for the modern data scientist.
+One shortcoming of our initial framework was the relative lack of emphasis on best practices for 
+computational reproducibility. This is one area that we would like to rectify in 
+future mentoring activities. The 
+University of Wisconsin-Madison has periodically offered a semester course in best practices for 
+computationally reproducible research (https://kbroman.org/Tools4RR/). We especially see collaborative 
+version control systems, such as Git and 
+Github, as essential tools for the modern data scientist.
+
+1. assessment of data science skills
 

### 2019-12-04 Rmd/introduction.Rmd
upd
fixed intro
minor updates to discussion

@@
 # Introduction
-
-- big data
-- new areas of application of statistical methods - social media
-- need for students to work with real data
-- motivate statistical analysis and statistics research by real world scientific research questions (Box 1976)
-- need for mentoring
 
@@
 @nolan2010computing argue for students to work with real data. 
-Working with real data allows them to develop skill not only in statistical analysis,
+Working with real data allows students to develop skill not only in statistical analysis,
 but also in data transfer from online sources, in data storage, and in using data 

### 2019-12-04 Rmd/results.Rmd
fixed headers
updates
minor updates to discussion

@@
-## Results
+# Outcomes
 
-We applied the project framework to our mentoring of two students. Both engaged in 12 months of mentored research during their senior year of undergraduate studies in statistics in 2015 and 2016.
+We applied the project framework to our mentoring of two students. 
+Both engaged in 12 months of mentored research during their senior year of 
+undergraduate studies in statistics.
 
-### Student Outcomes
+## Student Outcomes
 
-We subjectively assessed student outcomes through conversations in our weekly student research meetings.
+We subjectively assessed student outcomes through conversations in our weekly 
+student research meetings. Both students showed increases in confidence and ability to do 
+data science research.
 
-One student later enrolled in a statistics graduate program, while the other pursued employment in health care analytics.
+Both students secured positions in data science after graduation. 
+One student later enrolled in a statistics graduate program, 
+while the other pursued employment in health care analytics.
 
-### Mentor Outcomes
+## Scholarly outcomes
 
-### Scholarly outcomes & deliverables
+Our scholarly contributions include the `parseTweetFiles` R package on Github
+(https://github.com/rturn/parseTweetFiles) and presentations at conferences such as useR! 2016 and local poster sessions. Additionally, both students prepared end-of-project reports on their research.
+
+We consider below

### 2019-12-05 Rmd/discussion.Rmd
fixed methods & recompiled pdf (with line numbers)

@@
 
-1. build self-confidence, proficiency in data analysis and statistics
-
-1. greater interest in quantitative careers and/or grad school in quant fields
-
-1. Working with real data to solve real scientific questions
-
-1. Learn by example interplay between data science (statistics) and science research
-
-Benefits of our framework include enhanced
-
-## How does the framework relate to ideas in Nolan and Temple Lang 2010??
+The student test cases for our framework demonstrated greater 
+self-confidence and greater proficiency in data science skills over the 
+course of the research projects. They used real-world data sources to 
+address real scientific research questions.
+Additionally, they showed great interest
+in quantitative and data science careers. After graduation, one student immediately enrolled
+in statistics graduate training, while the other sought employment in 
+health care analytics.
 
@@
 
-Our measures of students' self-confidence in research ability was merely subjective. In 
+From our current perspective, we offer a number of framework critiques and opportunities for improvement. Our measure of students' self-confidence in research ability was merely subjective. In 
 future iterations of our framework, we would like to measure systematic and objective 
-outcomes. One strategy for implementing this is to administer a survey, such as \_\_\_, 
+outcomes. One strategy for implementing this is to administer a survey, including questions from @vance2017eight, 
 both before and after the mentored research project. We would use validated survey 
 questions that focused on student beliefs about themselves, their skills, and their 
-future careers. **maybe list resources for such survey questions**
+future careers.
 
-## Areas for improvement
-
-1. reproducible research best practices
-
-One shortcoming of our initial framework was the relative lack of emphasis on best practices for 
-computational reproducibility. This is one area that we would like to rectify in 
+One shortcoming of our initial framework was the relative lack of emphasis 
+on best practices for 
+computational reproducibility. This is one area that we would like to 
+rectify in 
 future mentoring activities. The 
-University of Wisconsin-Madison has periodically offered a semester course in best practices for 
-computationally reproducible research (https://kbroman.org/Tools4RR/). We especially see collaborative 
+University of Wisconsin-Madison has periodically offered a semester course 
+in best practices for 
+computationally reproducible research (https://kbroman.org/Tools4RR/). We 
+especially see collaborative 
 version control systems, such as Git and 
@@
 
-1. survey questions might include those from @vance2017eight. I have the survey as a pdf in mendeley. It is part of the supp info for @vance2017eight.
+### Framework development with backward design
+
+In future research, we will continue to develop our framework for 
+undergraduate data science research by explicitly 
+incorporating backward design principles [@wiggins2005understanding]. 
+Following @wiggins2005understanding, we will identify desired results, 
+determined acceptable evidence, and planned learning experiences.
+
+Before identifying desired results, we will prioritize topics from 
+@nolan2010computing. Specifically, we will assign all terms from Figure 1 
+of @nolan2010computing into one of three categories:
+
+1. worth being familiar with 
+1. important to know and do
+1. enduring understanding
+
+We've tabulated below the @nolan2010computing terms for the current 
+framework and its student projects.
+
+```{r table1, echo = FALSE, message = FALSE}
+library(magrittr)
+circles <- tibble::tibble(Class = 1:3, Circle = c("Worth being familiar with", "Important to know and do", "Enduring understanding"))
+readr::read_csv("../data/nolan-figure1-circles.csv") %>%
+  dplyr::left_join(circles) %>%
+  dplyr::arrange(desc(Class)) %>%
+  dplyr::select(- Class) %>%
+  gt::gt() %>%
+  gt::tab_header(title = "Prioritizing Key Terms from Figure 1 of @nolan2010computing", 
+                 subtitle ="xxx" ) %>%
+  gt::as_latex()
+```
+
+Potential benefits of incorporating backward design ideas include clearer 
+articulation of goals and better assessment of goal achievement.
+
+We see our framework as one contribution to scholarship on improving data 
+science training programs. Given the increasing economic need, in the USA 
+and abroad, for 
+data scientists and other researchers with quantitative training, we 
+anticipate that our framework and its future iterations will continue
+to prepare students for data science careers by offering training in tangible and 
+transferable analytic skills in the context of solving scientific questions.

### 2019-12-05 Rmd/jse-2019-ms.Rmd
fixed methods & recompiled pdf (with line numbers)

@@
 output: 
-  - bookdown::pdf_document2
-  - bookdown::html_document2
+  bookdown::pdf_document2:
+    includes:
+      in_header: head.tex
 ---

### 2019-12-05 Rmd/methods.Rmd
fixed methods & recompiled pdf (with line numbers)

--- 2019-12-04 Rmd/methods.Rmd
@@
 
-We used backward design principles to create and implement a 
+We designed and implemented a 
 framework for mentored undergraduate data science research 
-projects with big data. Below, we describe our use of 
-backward design and ideas from @nolan2010computing.
-We then detail our framework and its initial implementation.
-
-## Framework design
-
-We started with the backward design principles from @wiggins2005understanding. We prioritized
+projects with big data. Below, we describe our initial framework 
+and connect it to ideas from @nolan2010computing.
 
@@
 
-**justify each step**
-
-**Maybe use/cite Box's science and statistics article from... 1976?**
-
-1. Overview of our framework
-  a. student-led, hypothesis-driven research
-  b. BH and FB helped with data acquisition
-  c. helped with research question formulation
-  d. how to describe roles for students v mentors?
-
 Our framework for mentored student research projects involves these components:
-
-1. State the questions and research hypotheses that the two students chose
-
-1. Data description
-
-- how did we/they conceive of projects?
-
-- how did we balance our input with that of the students?
-- What data did we use?
-- How did we collect data? (which packages, code to sample twitter). How did Jinyu get her stock market data? 
-- What is the sampling scheme - for the 1% sample - from the entirety of Twitter? (Cite the webpage that documents the API)

### 2019-12-05 Rmd/introduction.Rmd
updates to methods

@@
 the development of faster and more powerful computers has fueled advances in 
-data science methods for big data. We present below a framework for mentored 
-undergraduate data science research and our findings from its initial 
-implementation for two statistics students. We conclude with lessons learned 
-and suggestions for those who wish to reproduce and refine our framework.
+data science methods for big data. Similar causes have led to a
+need for greater 
+numbers of scientists with quantitative skills. In efforts to enhance training
+and mentoring for trainees, we created a program that emphasizes many 
+transferable skills that contribute to career success in data science.
 
-Social media data, including tweets from Twitter and posts from Facebook, are 
+We elected to work with social media data. This choice was deliberate. In making 
+this decision, we recognized that social media data, such as tweets from 
+Twitter, can be acquired with little cost and that there is growing research 
+interest in social media in many social science disciplines, including political 
+science, communication studies, and sociology. We also anticipated 
+that our undergraduate trainees might be intrigued by the possibility of 
+analyzing social media data, since many young adults use Facebook, Twitter, 
+Instagram, and related sites.
+
+Some social media data, including tweets from Twitter, are 
 available through website application product interfaces (APIs). Twitter 
@@
 authorship of specific tweets when multiple persons share a single account [@drob]; and 
-study of rhetoric in recruiting political supporters [@pelled2018little,@wells2016trump]. 
+study of rhetoric in recruiting political supporters [@pelled2018little;@wells2016trump]. 
 Recognizing the potential utility of tweets for data science research, we created a 
@@
 from multiple file formats. In the case of Twitter data, tweets are stored in
-Javascript Object Notation (JSON) [@tweet_json,@json].
+Javascript Object Notation (JSON) [@tweet_json;@json].
 
@@
 own abilities, integration into the campus community, and interest in graduate training 
-[@baker2010beyond,@higgins2001reconceptualizing].
+[@baker2010beyond;@higgins2001reconceptualizing].
 A key component of our data science mentoring framework is the emphasis on using real data to answer real 
@@
 
-**Need to explain above paragraph with our two student examples. What types of questions did they  articulate? How did they translate to quant terms? How did they determine data availability? etc.**
-
 ## Our backgrounds
@@
 activities, including the Delta Program's mentoring class (https://delta.wisc.edu) and @handelsman2005entering. 
-Professor Erik Nordheim heavily 
+Professor Erik Nordheim
 influenced our approach to and philosophy of teaching statistics. 
 We studied with Professor Nordheim early in our teaching 
-careers, and his emphasis on active learning continues to influence our teaching practices.
+careers, and his emphasis on backward design and active learning continues to influence our teaching practices.
 

### 2019-12-05 Rmd/methods.Rmd
updates to methods

@@
 ## Framework implementation
+
+### Research question formulation
 
@@
 
+### Data acquisition
+
 We also incorporated data availability into our question formulation. 
@@
 
+### Data analysis and visualization
+
 After identifying research questions and publicly available data, 
@@
 Because 
-both projects primarily involved exploratory analyses of times series, 
+both projects involved exploratory analyses of times series, 
 we encouraged students to think about
@@
 
-### Examples
+In the case of the event detection project, we e
+
+### Presentation and communication of results
+
+Students presented their research in a variety of settings. 
+Each student presented at the annual undergraduate statistics poster session. 
+We also encouraged them to present at the annual university-wide 
+undergraduate research symposium.
+
+In planning with students for poster and slide presentations, 
+we (Hanlon and Boehm) emphasized
+the importance of succinctly stating the research question and its scientific 
+context. After clarifying the importance of the question, the student could 
+proceed with explaining many of the elements that we've described above. Namely, the 
+student would discuss the analyzed data and its acquisition while noting any shortcomings or 
+biases of the data. For oral presentations, we suggested that 
+students cautiously limit discussion of statistical methods, with the caveat that 
+they prepare to answer detailed methodological inquiries during the question and 
+answer session. 
+Our students created powerful data 
+visualizations for their projects. Their presentations also included their major results and future research directions.
+
+In efforts to develop student written communication, we encouraged both students to 
+prepare a written senior thesis document that detailed their research. In the senior 
+thesis, we suggested that the students describe in rigorous detail their statistical 
+methods. The rationale for this distinction, relative to the oral presentation, is 
+that a reader doesn't have access to a question and answer session, while a poster 
+session attendee may freely ask questions of the author.
+
+## Examples
 
@@
 Our framework for mentored student research projects involves these components:
+
+## Develop and practice skills in reproducible research to promote open science

### 2019-12-05 Rmd/methods.Rmd
updated & organized Methods

@@
 
-computational topics: LDA & topic modeling; time series analysis; sentiment analysis.
-
 Both mentored students combined computing with data analysis in 
@@
 
-Our framework for mentored student research projects involves these components:
+## Develop and practice skills in reproducible research to promote open science
 
-## Develop and practice skills in reproducible research to promote open science
+With the goal of promoting transparency in our research, we encouraged students to 
+use `git` for version control of their code and documents and to share their code via
+the website Github (https://github.com). One student also enrolled in Karl Broman's class on 
+tools for reproducible research. This class features `git` and Github throughout its lectures 
+and activities.

### 2019-12-06 Rmd/discussion.Rmd
expanded discussion

@@
 transferable analytic skills in the context of solving scientific questions.
+
+## Integrating more mentoring activities
+
+Our framework would benefit students more if we explicitly incorporate more 
+mentoring activities. Through professional development courses at the University
+of Wisconsin-Madison's Delta Program, we received training in how to offer 
+professional support to students. While we both informally supported our
+students, the Delta Program suggested
+ways to encourage the student's professional development through structured
+conversations and goal-setting. Additions like this would only enhance our 
+framework.
+
+@baker2010beyond discuss the role of faculty "developers" in 
+student success. A faculty "developer", as envisioned by @higgins2001reconceptualizing, 
+offers not only psychosocial and career support, like a mentor, but also 
+supports students' academic goals. Such relationships between developers 
+and students benefit both parties. The student gets support while the 
+developer refines her teaching and expands her scholarly network. We anticipate
+expanding our framework to more holistically support students.

### 2019-12-06 Rmd/methods.Rmd
expanded discussion

@@
 
+Our computational system for acquiring tweets involved several steps. We interacted 
+with the API via the R package `twitteR` [@twitteR]. We used the free Twitter streaming 
+API that gave us access to approximately one percent stream of all tweets during the 
+specified query time period. To ensure that we collected tweets continuously, we used 
+the linux tool `crontab` to execute our R script every five minutes. Each execution of 
+the R script performed a single streaming API query for five minutes. Twitter's 
+streaming API, at the time of our data collection, enforced rate limits on the 
+frequency and duration of queries. With the above settings, we continuously 
+collected tweets.
+
+We encouraged students to complement tweets with additional data from publicly 
+available sources.
+
 ### Data analysis and visualization
@@
 
-In the case of the event detection project, we e
+In the case of the event detection project, our student created 
+word clouds for every inferred "topic". 
+He also presented most probable words from each inferred topic, i.e., each distribution over words, as a bar plot.
+
+The student working on sentiment analysis and market index prices plotted a 
+daily sentiment "score" over time and presented it beside a plot of daily market 
+index prices and compared the two plots.
 
@@
 
-Our computational system for acquiring tweets involved several steps. We interacted 
-with the API via the R package `twitteR` [@twitteR]. We used the free Twitter streaming 
-API that gave us access to approximately one percent stream of all tweets during the 
-specified query time period. To ensure that we collected tweets continuously, we used 
-the linux tool `crontab` to execute our R script every five minutes. Each execution of 
-the R script performed a single streaming API query for five minutes. Twitter's 
-streaming API, at the time of our data collection, enforced rate limits on the 
-frequency and duration of queries. With the above settings, we continuously collected 
-tweets.
-
 ## Deepen computational reasoning skills
@@
 
-Students wrote R code to parse and organize tweet JSON []. They organized their R code into a package, and shared it on Github (https://github.com/rturn/parseTweetFiles). 
+Students wrote R code to parse and organize tweet JSON. 
+They organized their R code into a package, and shared it on Github (https://github.com/rturn/parseTweetFiles). 
 Each tweet's JSON included required fields, and, possibly, some optional fields. 
+Thus, students' code needed to accommodate variability in tweet structure.
 
@@
 modeling, sentiment analysis, and time series analysis to reach 
-conclusions about real world data.
+conclusions about real world phenomena.
 
@@
 Latent Dirichlet allocation is a bayesian nonparametric method for modeling text corpora as the result 
-of words chosen from topics.
+of words chosen from topics. The student treated tweets as "documents" (in the parlance of 
+topic modeling). The goal of topic modeling, then, was to infer the underlying "topics" (or probability distributions over words) from a collection of tweets.
 
@@
 the website Github (https://github.com). One student also enrolled in Karl Broman's class on 
-tools for reproducible research. This class features `git` and Github throughout its lectures 
-and activities.
+tools for reproducible research. This class features `git` and Github throughout 
+its lectures and activities.
+
+As we stated above, the students created an R package, `parseTweetFiles`, version controlled it with `git`, and shared it via Github.

### 2019-12-08 Rmd/introduction.Rmd
added markdown table

@@
 numbers of scientists with quantitative skills. In efforts to enhance training
-and mentoring for trainees, we created a program that emphasizes many 
+and mentoring for students, we created a program that emphasizes many 
 transferable skills that contribute to career success in data science.
 
-We elected to work with social media data. This choice was deliberate. In making 
+We elected to work with social media data. In making 
 this decision, we recognized that social media data, such as tweets from 
@@
 study of rhetoric in recruiting political supporters [@pelled2018little;@wells2016trump].
-Recognizing the potential utility of tweets for data science research, we created a 
+Recognizing the potential utility of tweets for data science research and teaching, 
+we created a 
 collection of tweets over time by repeated querying of the Twitter streaming API.
@@
 benefits, including improving students' development as thinkers and scholars, confidence in their 
-own abilities, integration into the campus community, and interest in graduate training 
+own abilities, integration into the campus community,
+and interest in graduate training 
 [@baker2010beyond;@higgins2001reconceptualizing].
-A key component of our data science mentoring framework is the emphasis on using real data to answer real 
+A key component of our data science mentoring framework is 
+the emphasis on using real data to answer 
 scientific questions. We believe that this process develops problem-solving skills that 
 students will need in their future careers in data science. 
-We encouraged the student to articulate a scientific research question, translate that question 
+We encouraged students to articulate a scientific research question, translate that question 
 into quantitative and statistical terms, determine which data could be used to address the 
 question, acquire the data, analyze data, visualize results, and communicate what they learned.
-
-We provide guidance regarding selection of
 

### 2019-12-08 Rmd/methods.Rmd
added markdown table
added backward design fig1

@@
 and connect it to ideas from @nolan2010computing.
+
+```{r t1, echo = FALSE}
+library(magrittr)
+step <- c(rep("Question formulation", 3),
+          rep("Data acquisition", 3), 
+          rep("Data analysis and visualization", 2), 
+          rep("Presentation and communication", 3)
+)
+details <- c("List ideas", "Prioritize ideas", "Check data availability",
+             "Download tweets and other data", "Parse tweet JSON", "Organize data as times series",
+             "Exploratory plots", "Summary statistics",
+             "Interpret and summarize graphs and summary statistics", "Prepare written materials and figures", "Share findings as posters, slides, and reports"
+             )
+tibble::tibble(step, details) %>%
+#  dplyr::group_by(step) %>%
+  gt::gt() %>%
+  gt::tab_style(
+    style = list(
+      gt::cell_fill(color = "lightcyan"),
+      gt::cell_text(style = "italic")
+      ),
+    locations = gt::cells_data(
+      columns = dplyr::vars(details),
+      rows = step == "Question formulation"
+      )
+  ) %>%
+    gt::tab_style(
+    style = list(
+      gt::cell_fill(color = "pink"),
+      gt::cell_text(style = "italic")
+      ),
+    locations = gt::cells_data(
+      columns = dplyr::vars(details),
+      rows = step == "Data acquisition"
+      )
+  ) %>%
+  gt::tab_header(
+    title = gt::md("Framework Overview"), subtitle = "xxx"
+  ) %>%
+  gt::tab_style(
+    style = list(
+      gt::cell_fill(color = "apricot"),
+      gt::cell_text(style = "italic")
+      ),
+    locations = gt::cells_data(
+      columns = dplyr::vars(details),
+      rows = step == "Data analysis and visualization"
+      )
+  )
+```
 

### 2019-12-09 Rmd/methods.Rmd
updated methods

--- 2019-12-06 Rmd/methods.Rmd
@@
 and connect it to ideas from @nolan2010computing.
+
+```{r t1, echo = FALSE}
+library(magrittr)
+step <- c(rep("Question formulation", 3),
+          rep("Data acquisition", 3), 
+          rep("Data analysis and visualization", 2), 
+          rep("Presentation and communication", 3)
+)
+details <- c("List ideas", "Prioritize ideas", "Check data availability",
+             "Download tweets and other data", "Parse tweet JSON", "Organize data as times series",
+             "Exploratory plots", "Summary statistics",
+             "Interpret and summarize graphs and summary statistics", "Prepare written materials and figures", "Share findings as posters, slides, and reports"
+             )
+tibble::tibble(step, details) %>%
+#  dplyr::group_by(step) %>%
+  gt::gt() %>%
+  gt::tab_style(
+    style = list(
+      gt::cell_fill(color = "lightcyan"),
+      gt::cell_text(style = "italic")
+      ),
+    locations = gt::cells_data(
+      columns = dplyr::vars(details),
+      rows = step == "Question formulation"
+      )
+  ) %>%
+    gt::tab_style(
+    style = list(
+      gt::cell_fill(color = "pink"),
+      gt::cell_text(style = "italic")
+      ),
+    locations = gt::cells_data(
+      columns = dplyr::vars(details),
+      rows = step == "Data acquisition"
+      )
+  ) %>%
+  gt::tab_header(
+    title = gt::md("Framework Overview"), subtitle = "xxx"
+  ) %>%
+  gt::tab_style(
+    style = list(
+      gt::cell_fill(color = "apricot"),
+      gt::cell_text(style = "italic")
+      ),
+    locations = gt::cells_data(
+      columns = dplyr::vars(details),
+      rows = step == "Data analysis and visualization"
+      )
+  ) %>%
+  gt::as_latex()
+```
 
@@
 Both students used Twitter tweets, which we accessed through a Twitter 
-streaming API.
+streaming API. One student chose to apply topic modeling to tweets. 
+The other elected to apply sentiment analysis to daily 
+tweet collections. By using these methods, sentiment analysis and 
+topic modeling, students learned an aspect of text analysis that 
+is not taught in traditional statistics courses.
 

### 2019-12-09 Rmd/results.Rmd
expanded discussion
fixed results

@@
-# Outcomes
+# Results
 
 We applied the project framework to our mentoring of two students. 
-Both engaged in 12 months of mentored research during their senior year of 
-undergraduate studies in statistics.
+Both engaged in 12 months of research during their senior year of 
+undergraduate studies in statistics. Below, we describe three categories of outcomes:
+
+1. student outcomes
+1. mentor outcomes
+1. scholarly outcomes
 
@@
 Both students secured positions in data science after graduation. 
-One student later enrolled in a statistics graduate program, 
+One student enrolled in a statistics graduate program, 
 while the other pursued employment in health care analytics.
+
+## Mentor outcomes
+
+We (Boehm and Hanlon) grew as mentors during our work with the two students. We successfully
+guided junior scientists through a productive, hands-on research experience, and we anticipate
+refining the framework in future iterations.
 
@@
 Our scholarly contributions include the `parseTweetFiles` R package on Github
-(https://github.com/rturn/parseTweetFiles) and presentations at conferences such as useR! 2016 and local poster sessions. Additionally, both students prepared end-of-project reports on their research.
-
-We consider below
+(https://github.com/rturn/parseTweetFiles) and presentations at conferences such as useR! 2016 (R users' conference)
+and local poster sessions. Additionally, both students prepared end-of-project reports on 
+their research.

### 2019-12-12 Rmd/jse-2019-ms.Rmd
updates to methods
added markdown table
many updates

--- 2019-11-28 Rmd/jse-2019-ms.Rmd
@@
 ---
-title: 'A framework for undergraduate data science research'
+title: 'A framework for mentored data science research'
 author: "Frederick J. Boehm and Bret M. Hanlon"
@@
 bibliography: jse.bib
+geometry: "left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm"
 output: 
-  - github_document
-  - bookdown::pdf_document2
-  - bookdown::html_document2
+  bookdown::pdf_book:
+    includes:
+      in_header: head.tex
 ---

### 2019-12-12 Rmd/methods.Rmd
many updates

--- 2019-12-06 Rmd/methods.Rmd
@@
 and connect it to ideas from @nolan2010computing.
+
+```{r t1, echo = FALSE}
+library(magrittr)
+step <- c(rep("Question formulation", 3),
+          rep("Data acquisition", 3), 
+          rep("Data analysis and visualization", 2), 
+          rep("Presentation and communication", 3)
+)
+details <- c("List ideas", "Prioritize ideas", "Check data availability",
+             "Download tweets and other data", "Parse tweet JSON", "Organize data as times series",
+             "Exploratory plots", "Summary statistics",
+             "Interpret and summarize graphs and summary statistics", "Prepare written materials and figures", "Share findings as posters, slides, and reports"
+             )
+tibble::tibble(step, details) %>%
+#  dplyr::group_by(step) %>%
+  gt::gt() %>%
+  gt::tab_style(
+    style = list(
+      gt::cell_fill(color = "lightcyan"),
+      gt::cell_text(style = "italic")
+      ),
+    locations = gt::cells_data(
+      columns = dplyr::vars(details),
+      rows = step == "Question formulation"
+      )
+  ) %>%
+    gt::tab_style(
+    style = list(
+      gt::cell_fill(color = "pink"),
+      gt::cell_text(style = "italic")
+      ),
+    locations = gt::cells_data(
+      columns = dplyr::vars(details),
+      rows = step == "Data acquisition"
+      )
+  ) %>%
+  gt::tab_header(
+    title = gt::md("Framework Overview"), subtitle = "t"
+  ) %>%
+  gt::tab_style(
+    style = list(
+      gt::cell_fill(color = "yellow"),
+      gt::cell_text(style = "italic")
+      ),
+    locations = gt::cells_data(
+      columns = dplyr::vars(details),
+      rows = step == "Data analysis and visualization"
+      )
+  ) %>%
+  gt::as_latex()
+```
 
@@
 the website Github (https://github.com). One student also enrolled in Karl Broman's class on 
-tools for reproducible research. This class features `git` and Github throughout 
+tools for reproducible research (https://kbroman.org/Tools4RR/). This class features `git` and Github throughout 
 its lectures and activities.

### 2019-12-12 Rmd/results.Rmd
many updates

@@
 
+Students benefited from our emphasis on the four central concepts, 
+three from @nolan2010computing plus reproducible research skills. 
+The research projects successfully drew on emerging areas of 
+statistical computing, namely text analysis. They combined 
+computational topics, including topic modeling and time series 
+methods, with data analysis in the practice of statistics. Although 
+we didn't formally measure them, our informal assessment indicates 
+that students' computational reasoning skills increased over the 
+duration of our projects. Students used a variety of computing tools 
+and methods to arrive at a practical solution to a selected task. 
+They became more skilled in computing with R and shell scripts and 
+more fluid in their verbal explanations during our regular meetings [@r].
+
+Our framework's emphasis on reproducible research skills is 
+evidenced by the students' R package, `parseTweetFiles`, which is 
+both version controlled with `git` and shared via Github.com.
+
 ## Mentor outcomes

### 2019-12-14 Rmd/discussion.Rmd
added backward design fig1
many updates
added bret's feedback

@@
 outcomes. One strategy for implementing this is to administer a survey, including questions from @vance2017eight, 
-both before and after the mentored research project. We would use validated survey 
+both before and after the mentored research project. We would use survey 
 questions that focused on student beliefs about themselves, their skills, and their 
@@
 determined acceptable evidence, and planned learning experiences.
+
+```{r backward, echo = FALSE}
+knitr::include_graphics("backward-design.jpg")
+```
 

### 2019-12-14 Rmd/introduction.Rmd
added bret's feedback

@@
 activities, including the Delta Program's mentoring class (https://delta.wisc.edu) and @handelsman2005entering. 
-Professor Erik Nordheim
+Professor Erik Nordheim's class, "Teaching Statistics", 
 influenced our approach to and philosophy of teaching statistics. 
 We studied with Professor Nordheim early in our teaching 
-careers, and his emphasis on backward design and active learning continues to influence our teaching practices.
+careers, and his course's emphasis on backward design and active learning continues to influence our teaching practices (https://delta.wisc.edu).
 
 We both have experience in teaching undergraduate introductory 
-statistics courses with enrollments over 
+statistics courses with enrollments of more than 
 100 students. Through our interactions with students in these 

### 2019-12-14 Rmd/jse-2019-ms.Rmd
added bret's feedback

@@
   bookdown::pdf_book:
+    toc: FALSE
+    number_sections: FALSE
     includes:
@@
 ---
-
-Last modified: `r params$lastmod`
 

### 2019-12-14 Rmd/methods.Rmd
added bret's feedback

@@
 
-## Relating to three ideas from @nolan2010computing
-
-We incorporate three key aspects that @nolan2010computing identified:
-
-1. broaden statistical computing to include emerging areas
-1. deepen computational reasoning skills
-1. combine computational topics with data analysis in the practice of statistics
-
-Additionally, our projects gave students opportunities to develop and to practice skills in reproducible 
-research. Given the growing imperative to document and share code to promote open science, we feel that 
-this skill set equals in importance the three points above.
-
-4. develop and practice skills in reproducible research to promote open science
-
-Below, we describe how our framework enabled students to achieve competence in the four areas listed 
-above.
-
 ## Broaden statistical computing to include emerging areas

### 2019-12-14 Rmd/cover-letter.Rmd
added submission materials.

---
title: ""
author: ""
date: "12/14/2019"
output: github_document
---

Dear editors:

We're writing to submit a manuscript for the Journal of Statistics Education's special issue on "Computing in the 
Statistics Curriculum". In it, we describe our mentoring framework for undergraduate research projects in data science.
We relate the framework to many ideas from Nolan & Temple Lang (2010) and offer minor amendments to Nolan & Temple Lang's 
arguments.

We developed the framework while guiding two undergraduate students through their senior projects in statistics.

Our framework integrates ideas from Nolan and Temple Lang (2010) with literature on mentoring and curriculum design.

We thank you for your careful consideration of our manuscript.

Sincerely,

Frederick Boehm and Bret Hanlon

### 2019-12-14 Rmd/methods.Rmd
added submission materials.

@@
 In planning with students for poster and slide presentations, 
-we (Hanlon and Boehm) emphasized
+we emphasized
 the importance of succinctly stating the research question and its scientific 

### 2019-12-18 Rmd/discussion.Rmd
updated anonymous version

@@
 rectify in 
-future mentoring activities. The 
-University of Wisconsin-Madison has periodically offered a semester course 
+future mentoring activities. The university has periodically offered a semester course 
 in best practices for 
-computationally reproducible research (https://kbroman.org/Tools4RR/). We 
+computationally reproducible research. We 
 especially see collaborative 
@@
 Github, as essential tools for the modern data scientist.
-
-1. assessment of data science skills
-
-1. assessment of attitudes (pre and post survey??)
 
@@
 Our framework would benefit students more if we explicitly incorporate more 
-mentoring activities. Through professional development courses at the University
-of Wisconsin-Madison's Delta Program, we received training in how to offer 
+mentoring activities. Through professional development courses at the university, we received training in how to offer 
 professional support to students. While we both informally supported our
-students, the Delta Program suggested
+students, the professional development program suggested
 ways to encourage the student's professional development through structured

### 2019-12-18 Rmd/introduction.Rmd
added submission materials.
updated anonymous version

@@
 During the time when we first implemented our framework, we served as early-career instructors in 
-the statistics department at the University of Wisconsin-Madison. One of us (Hanlon) had prior 
-experience in mentoring students, while the other (Boehm) had none. Our initial 
+the statistics department at a large university. One of us had prior 
+experience in mentoring students, while the other had none. Our initial 
 conceptualization of mentoring drew heavily on ideas we first encountered in professional 
 development 
-activities, including the Delta Program's mentoring class (https://delta.wisc.edu) and @handelsman2005entering. 
-Professor Erik Nordheim's class, "Teaching Statistics", 
+activities, including a professional development class and @handelsman2005entering. 
+A professional development class, "Teaching Statistics", 
 influenced our approach to and philosophy of teaching statistics. 
-We studied with Professor Nordheim early in our teaching 
-careers, and his course's emphasis on backward design and active learning continues to influence our teaching practices (https://delta.wisc.edu).
+This course's emphasis on backward design and active learning continues to influence our teaching practices.
 

### 2020-03-19 Rmd/introduction.Rmd
removed anonymous files

@@
 interest in social media in many social science disciplines, including political 
-science, communication studies, and sociology. We also anticipated 
+science, communication studies, public health, and sociology. We also anticipated 
 that our undergraduate trainees might be intrigued by the possibility of 

### 2020-03-19 Rmd/jse-2019-ms.Rmd
removed anonymous files

--- 2019-12-05 Rmd/jse-2019-ms.Rmd
@@
 ---
-title: 'Backward designing a reproducible framework for undergraduate data science research'
+title: 'Mentored data science research with tweets'
 author: "Frederick J. Boehm and Bret M. Hanlon"
-date: "8/20/2019"
+date: "May 15, 2020"
 params:
@@
 bibliography: jse.bib
+geometry: "left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm"
 output: 
-  bookdown::pdf_document2:
+  bookdown::pdf_book:
+    toc: FALSE
+    number_sections: TRUE
     includes:

### 2020-04-10 Rmd/tweets.Rmd
started Rmd for working with tweets

---
title: "Tweets"
author: "Frederick J. Boehm"
date: "4/9/2020"
output: html_document
---

## Goals

1. Write a R script to query twitter streaming api to get 1 percent sample over 5'.
1. Demonstrate how to set up a crontab to execute script every 5'.
1. Verify that crontab approach works by running it on my computer for, say, 15'.
    1. Cautions about twitter databases: deletions and amendments to tweets
1. Examine structure of JSON obtained from api. 
1. Create phony JSON to represent a collection of tweets.
    1. Use lorem ipsum for text of tweets

## R code to query twitter streaming api

Install `rtweet` package from CRAN.

```{r, eval = FALSE}
install.packages("rtweet")
install.packages("lubridate")
```

```{r}
rtweet::stream_tweets(
  timeout = (60 * 5),
  parse = TRUE,
  file_name = paste0("../data/", lubridate::now(), "-tweets")
)
```

### Colophon

This report was generated on `r Sys.time()` using the following computational environment and dependencies:

```{r colophon, cache = FALSE}
# which R packages and versions?
if ("devtools" %in% installed.packages()) devtools::session_info()
```

The current Git commit details are:

```{r}
# what commit is this file at? 
if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
```

### 2020-04-11 Rmd/tweets.Rmd
added code to parse tweet texts

@@
 
+## Parse json file
+
+```{r}
+library(magrittr)
+```
+
+```{r}
+tweets <- rtweet::parse_stream("../data/2020-04-09 16:31:44-tweets.json") %>%
+  dplyr::filter(lang == "en") %>%
+  tidytext::unnest_tokens(output = word, input = text)
+```
+
+## Prepare document-term matrix from tweet words
+
+```{r}
+# create a twitter-specific stop word list
+twitter_stop <- tibble::tibble(word = c("https", "http", "t.co"))
+# get a list of english language words and names
+library(qdapDictionaries) # contains object GradyAugmented
+```
+
+```{r}
+# https://www.tidytextmining.com/topicmodeling.html#library-heist
+library(tidytext)
+word_counts <- tweets %>% 
+  dplyr::select(status_id, word) %>%
+  dplyr::anti_join(stop_words) %>%
+  dplyr::anti_join(twitter_stop) %>%
+  dplyr::filter(word %in% GradyAugmented) %>%
+  dplyr::count(status_id, word, sort = TRUE) %>%
+  dplyr::ungroup()
+```
+
 ### Colophon

### 2020-04-18 Rmd/jse-2019-ms.Rmd
updated manuscript; added todo file

@@
 
-```{r child = "introduction.Rmd"}
-```
+## Introduction
 
-```{r child = "methods.Rmd"}
-```
+Twitter has changed the way that we communicate. 
+In only 140 characters, users speak their piece on politics, 
+current events, sports, media, and many other topics. 
+Recent development of accessible statistical methods for text 
+analysis now enable instructors to use tweets as pedagogical 
+tools in mentoring undergraduate research projects.
+
+Some social media data, including tweets from Twitter, are 
+available through website application product interfaces (APIs). Twitter 
+shares, via a streaming API, a sample of approximately one percent of all 
+tweets during an API query time period [@tweet_stream].
+Researchers have studied tweets for a 
+variety of purposes, including 
+inference of relationships and social networks among users [@lin2011joint]; 
+determination of 
+authorship of specific tweets when multiple persons share a single account [@drob]; and 
+study of rhetoric in recruiting political supporters [@pelled2018little;@wells2016trump].
+Recognizing the potential utility of tweets for data science research and teaching, 
+we created a 
+collection of tweets over time by repeated querying of the Twitter streaming API.
+
+@nolan2010computing argue for students to work with real data. 
+Working with real data allows students to develop skill not only in statistical analysis,
+but also in related data science skills, including data transfer from online sources, data storage, and using data 
+from multiple file formats.
+
+Mentoring in the work place and in higher education can have many 
+benefits, including improving students' development as thinkers and scholars, confidence in their 
+own abilities, integration into the campus community,
+and interest in graduate training 
+[@baker2010beyond;@higgins2001reconceptualizing].
+
+We present below 1) an approach for collecting tweets in real 
+time and 2) statistical methods for detecting social media 
+events via topic modeling of tweets and 3) reflections on 
+using this data set in research mentoring for undergraduate 
+students.
+
+## Methods
 

### 2020-04-18 Rmd/jse-2019-ms.Rmd
ignored lit folder; added to todo list

@@
 
-@nolan2010computing argue for students to work with real data. 
-Working with real data allows students to develop skill not only in statistical analysis,
-but also in related data science skills, including data transfer from online sources, data storage, and using data 
+@nolan2010computing argue for students to work with real 
+data. 
+Working with real data allows students to develop skill not 
+only in statistical analysis,
+but also in related data science skills, including data 
+transfer from online sources, data storage, and using data 
 from multiple file formats.
 
-Mentoring in the work place and in higher education can have many 
+Mentoring in the work place and in higher education can have 
+many 
 benefits, including improving students' development as thinkers and scholars, confidence in their 
@@
 [@baker2010beyond;@higgins2001reconceptualizing].
+
+The content of tweets over time particularly intrigued us. We 
+hypothesized that high-profile social media events would generate lots of tweets, and that we'd detect social media events through changes in tweet topic content over time.
 
@@
 
+### Collecting tweets
+
+We collected tweets by automated, repeated querying of Twitter's 
+streaming API. This involved multiple steps. First, we created a 
+new account on Twitter. With these user credentials, we used the 
+R package `rtweet` to query the API. Because we work on computers 
+with linux operating systems, we made use of linux software 
+`crontab` to repeatedly execute R code to submit API queries. Each 
+query lasts a user-specified duration. We timed the API queries so 
+that there was no time lag between queries. We stored API query 
+results in their native JSON format. The R package `rtweet` 
+provides functions that parse tweet JSON to R data frames. We then 
+conducted all further analyses with the R data frames.
+
 ```{r child = "results.Rmd"}

### 2020-04-18 Rmd/jse-2019-ms.Rmd
added text to ms

@@
 
+As a proof of principle of our event detection strategy, we chose to 
+examine tweets before, during, and after the National Football League's 
+2015 Super Bowl.
+We fitted five latent Dirichlet allocation models.
+We examined five time periods, each lasting one hour in duration.
+The first period began approximately 48h before the Super Bowl halftime show.
+Subsequent time periods started at -24, 0, +24, and +48 hours after the approximate start time of the halftime show.
+
+We defined each time period to be a single corpus of tweets. We then fitted latent Dirichlet allocation models to each corpus.
+
 ```{r child = "results.Rmd"}

### 2020-05-03 Rmd/jse-2019-ms.Rmd
updated ms

@@
 ---
-title: 'Mentored data science research with tweets'
+title: 'Using tweet content to detect social media events'
 author: "Frederick J. Boehm and Bret M. Hanlon"
-date: "May 15, 2020"
+date: "June 15, 2020"
 params:
@@
 Twitter has changed the way that we communicate. 
-In only 140 characters, users speak their piece on politics, 
+In only 280 characters, users speak their piece on politics, 
 current events, sports, media, and many other topics. 
@@
 
-@nolan2010computing argue for students to work with real 
+Our collection of tweets over time served as a valuable resource in our mentoring 
+of undergraduate data science research. @nolan2010computing argue for students to 
+work with real 
 data. 
@@
 transfer from online sources, data storage, and using data 
-from multiple file formats.
+from multiple file formats. Analyzing our collection of tweets gave students 
+opportunities to develop skills in all of the above areas.
 
@@
 The content of tweets over time particularly intrigued us. We 
-hypothesized that high-profile social media events would generate lots of tweets, and that we'd detect social media events through changes in tweet topic content over time.
+hypothesized that high-profile social media events would generate lots of 
+tweets, and that we'd detect social media events through changes in tweet 
+topic content over time.
 
@@
 time and 2) statistical methods for detecting social media 
-events via topic modeling of tweets and 3) reflections on 
+events via latent Dirichlet allocation modeling of collections of tweets and 3) reflections on 
 using this data set in research mentoring for undergraduate 
@@
 
-### Collecting tweets
+### Study design
 
-We collected tweets by automated, repeated querying of Twitter's 
-streaming API. This involved multiple steps. First, we created a 
+We sought to validate our hypothesis that we could detect a major social 
+media event by examining tweet topic content at distinct time periods. We 
+chose the National Football League's Super Bowl as a case study.
+
+As a proof of principle of our event detection strategy, we chose to 
+examine tweets before, during, and after the National Football League's 
+2015 Super Bowl. We fitted latent Dirichlet allocation models for each of five 
+distinct one-hour time periods. The first period began approximately 48h 
+before the Super Bowl halftime show. Subsequent time periods started at -24, 
+0, +24, and +48 hours after the approximate start time of the halftime show.
+
+We defined each time period to be a single corpus of tweets. We then fitted latent Dirichlet allocation models to each corpus.
+
+### Collecting tweets over time
+
+We include here instructions for creating a tweet collection. First, we created a 
 new account on Twitter. With these user credentials, we used the 
 R package `rtweet` to query the API. Because we work on computers 
-with linux operating systems, we made use of linux software 
+with linux operating systems, we use the linux software 
 `crontab` to repeatedly execute R code to submit API queries. Each 
-query lasts a user-specified duration. We timed the API queries so 
-that there was no time lag between queries. We stored API query 
+query lasts a user-specified duration. We time the API queries so 
+that there is no time lag between queries. We store API query 
 results in their native JSON format. The R package `rtweet` 
@@
 
-As a proof of principle of our event detection strategy, we chose to 
-examine tweets before, during, and after the National Football League's 
-2015 Super Bowl.
-We fitted five latent Dirichlet allocation models.
-We examined five time periods, each lasting one hour in duration.
-The first period began approximately 48h before the Super Bowl halftime show.
-Subsequent time periods started at -24, 0, +24, and +48 hours after the approximate start time of the halftime show.
+### Querying Twitter API to get complete tweets
 
-We defined each time period to be a single corpus of tweets. We then fitted latent Dirichlet allocation models to each corpus.
+We queried a tweets database, created with the methods described above, to get ID numbers for 
+tweets from the desired time periods. We then submitted API queries to Twitter
+to get the full content of the tweets, including the tweet text.
+We provide below the R code that we used to query the Twitter API to obtain full tweet content.
+
+```{r, eval = FALSE}
+rtweet::lookup_tweets()
+```
+
+### Tweet structure
+
+Tweets are available as Javascript Object Notation (JSON) objects.
+Every tweet consists of multiple named fields, each of which is a key-value pair.
+The number of fields per tweet depends on user settings, retweet status, and other factors.
+
+**PLACE TWEET JSON HERE**
+
+### Parsing text of tweets
+
+We used functions from the `rtweet` package to parse tweet JSON into a data frame.
+From there, we used `tidytext` R package functions to break the tweet text into individual words for latent Dirichlet allocation.
+We discarded commonly used "stop words" and emojis.
+
+Latent Dirichlet allocation models require that the corpus be inputted as a 
+document by term matrix. Each row corresponds to a single document (a single 
+tweet), and each column is a single term (or word). Each cell contains a 
+count (the number of 
+occurrences of a term in the specified document). We created a document by 
+term matrix with the R functions `` from the `` R package.
 

### 2020-05-05 Rmd/jse-2019-ms.Rmd
fixed intro wording

@@
 ---
-title: 'Using tweet content to detect social media events'
+title: 'Detect social media events through tweet content analysis'
 author: "Frederick J. Boehm and Bret M. Hanlon"
@@
       in_header: head.tex
+  bookdown::word_document2:
+    toc: FALSE
 ---
 
-```{r child = "abstract.Rmd"}
+```{r child = "abstract.Rmd", eval = FALSE}
 ```
 
-## Introduction
+# Introduction
 
-Twitter has changed the way that we communicate. 
+Twitter has profoundly changed how we communicate. 
 In only 280 characters, users speak their piece on politics, 
@@
 Recent development of accessible statistical methods for text 
-analysis now enable instructors to use tweets as pedagogical 
-tools in mentoring undergraduate research projects.
+analysis now enable mentors to use tweets as pedagogical 
+tools in guiding undergraduate research projects.
 
-Some social media data, including tweets from Twitter, are 
+Some social media data, including tweets from Twitter, is
 available through website application product interfaces (APIs). Twitter 
 shares, via a streaming API, a sample of approximately one percent of all 
-tweets during an API query time period [@tweet_stream].
-Researchers have studied tweets for a 
-variety of purposes, including 
-inference of relationships and social networks among users [@lin2011joint]; 
-determination of 
+tweets during an API query time period [@tweet_stream]. Any Twitter user can 
+access this one percent sample. For a price, researchers can get access to larger 
+tweet samples.
+
+With large collections of tweets, researchers have studied
+inference of relationships and social networks among Twitter users [@lin2011joint]; 
 authorship of specific tweets when multiple persons share a single account [@drob]; and 
-study of rhetoric in recruiting political supporters [@pelled2018little;@wells2016trump].
+rhetoric in recruiting political supporters [@pelled2018little;@wells2016trump].
 Recognizing the potential utility of tweets for data science research and teaching, 
-we created a 
-collection of tweets over time by repeated querying of the Twitter streaming API.
+we created a collection of tweets over time by repeated querying of the Twitter streaming API.
 
@@
 of undergraduate data science research. @nolan2010computing argue for students to 
-work with real 
-data. 
+work with real data. 
 Working with real data allows students to develop skill not 
@@
 transfer from online sources, data storage, and using data 
-from multiple file formats. Analyzing our collection of tweets gave students 
-opportunities to develop skills in all of the above areas.
+from multiple file formats. Collaboratively asking and addressing novel questions with 
+our collection of tweets gave mentored students 
+opportunities to develop skills in all of these areas.
 
-Mentoring in the work place and in higher education can have 
-many 
-benefits, including improving students' development as thinkers and scholars, confidence in their 
-own abilities, integration into the campus community,
-and interest in graduate training 
+Mentoring in the work place and in higher education has many 
+benefits, including improving 1) students' development as thinkers and scholars, 2) confidence in their 
+own abilities, 3) integration into the campus community,
+and 4) interest in graduate training 
 [@baker2010beyond;@higgins2001reconceptualizing].
 
-The content of tweets over time particularly intrigued us. We 
+While our tweet collection enables us to address many possible research questions, the 
+content of tweets over time particularly intrigued us. We 
 hypothesized that high-profile social media events would generate lots of 
@@
 events via latent Dirichlet allocation modeling of collections of tweets and 3) reflections on 
-using this data set in research mentoring for undergraduate 
+using this data set in research mentoring of undergraduate 
 students.
@@
 We sought to validate our hypothesis that we could detect a major social 
-media event by examining tweet topic content at distinct time periods. We 
-chose the National Football League's Super Bowl as a case study.
-
+media event by examining tweet topic content at distinct time periods. 
 As a proof of principle of our event detection strategy, we chose to 
@@
 2015 Super Bowl. We fitted latent Dirichlet allocation models for each of five 
-distinct one-hour time periods. The first period began approximately 48h 
+distinct one-hour periods. The first period began approximately 48h 
 before the Super Bowl halftime show. Subsequent time periods started at -24, 
@@
 
-We defined each time period to be a single corpus of tweets. We then fitted latent Dirichlet allocation models to each corpus.
+We defined each time period to be a single corpus of tweets. Each tweet constituted a document. We then fitted latent Dirichlet allocation models to each of the five corpora.
 

### 2020-05-06 Rmd/jse-2019-ms.Rmd
incorporated EACD suggestions

@@
 ---
-title: 'Detect social media events through tweet content analysis'
+title: 'Detecting social media events through tweet content analysis'
 author: "Frederick J. Boehm and Bret M. Hanlon"
@@
 Twitter has profoundly changed how we communicate. 
-In only 280 characters, users speak their piece on politics, 
+In only 280 characters, users can instantly contribute to public conversations on politics, 
 current events, sports, media, and many other topics. 
 Recent development of accessible statistical methods for text 
-analysis now enable mentors to use tweets as pedagogical 
+analysis now enable mentors to use tweets as contemporary pedagogical 
 tools in guiding undergraduate research projects.
@@
 Some social media data, including tweets from Twitter, is
-available through website application product interfaces (APIs). Twitter 
-shares, via a streaming API, a sample of approximately one percent of all 
-tweets during an API query time period [@tweet_stream]. Any Twitter user can 
-access this one percent sample. For a price, researchers can get access to larger 
-tweet samples.
+available through website application product interfaces (APIs). 
+By way of a streaming API, Twitter shares a sample of approximately one percent of all 
+tweets during an API query time period [@tweet_stream]. Any Twitter user can freely 
+access this one percent sample, whereas access to a larger selection is available to researchers for a fee.
 
-With large collections of tweets, researchers have studied
+Using large collections of tweets, researchers have studied
 inference of relationships and social networks among Twitter users [@lin2011joint]; 
@@
 
-Our collection of tweets over time served as a valuable resource in our mentoring 
-of undergraduate data science research. @nolan2010computing argue for students to 
-work with real data. 
-Working with real data allows students to develop skill not 
+In line with recent calls for students to work with real data [@nolan2010computing],   our collection of tweets has served as a valuable resource in our mentoring 
+of undergraduate data science research. 
+Working with real data allows students to develop proficiency not 
 only in statistical analysis,
@@
 our collection of tweets gave mentored students 
-opportunities to develop skills in all of these areas.
+opportunities to develop competency in all of these areas.
 
@@
 own abilities, 3) integration into the campus community,
-and 4) interest in graduate training 
-[@baker2010beyond;@higgins2001reconceptualizing].
+and 4) interest in graduate training [@baker2010beyond;@higgins2001reconceptualizing]. 
+We strived to develop trusting, mutually respectful mentoring relationships with 
+our students while advising their senior projects. With input from us,
+the students selected appealing research questions.
 
@@
 content of tweets over time particularly intrigued us. We 
-hypothesized that high-profile social media events would generate lots of 
+hypothesized that high-profile social media events would generate a high volume of 
 tweets, and that we'd detect social media events through changes in tweet 
-topic content over time.
-
-We present below 1) an approach for collecting tweets in real 
+topic content over time. We present below 1) an approach for collecting tweets in real 
 time and 2) statistical methods for detecting social media 
@@
 As a proof of principle of our event detection strategy, we chose to 
-examine tweets before, during, and after the National Football League's 
+analyze tweets before, during, and after the National Football League's 
 2015 Super Bowl. We fitted latent Dirichlet allocation models for each of five 

### 2020-05-06 Rmd/jse-2019-ms.Rmd
minor reword
updated ms

@@
 ---
-title: 'Detecting social media events through tweet content analysis'
+title: 'Analyzing tweets to detect social media events'
 author: "Frederick J. Boehm and Bret M. Hanlon"
@@
 While our tweet collection enables us to address many possible research questions, the 
-content of tweets over time particularly intrigued us. We 
+dynamic content of tweets over time particularly intrigued us. We 
 hypothesized that high-profile social media events would generate a high volume of 
@@
 events via latent Dirichlet allocation modeling of collections of tweets and 3) reflections on 
-using this data set in research mentoring of undergraduate 
-students.
+using this data set in research mentoring of undergraduate students.
 
 ## Methods
-
-### Study design
-
-We sought to validate our hypothesis that we could detect a major social 
-media event by examining tweet topic content at distinct time periods. 
-As a proof of principle of our event detection strategy, we chose to 
-analyze tweets before, during, and after the National Football League's 
-2015 Super Bowl. We fitted latent Dirichlet allocation models for each of five 
-distinct one-hour periods. The first period began approximately 48h 
-before the Super Bowl halftime show. Subsequent time periods started at -24, 
-0, +24, and +48 hours after the approximate start time of the halftime show.
-
-We defined each time period to be a single corpus of tweets. Each tweet constituted a document. We then fitted latent Dirichlet allocation models to each of the five corpora.
 
@@
 
+### Study design
+
+We sought to validate our hypothesis that we could detect a major social 
+media event by examining tweet topic content at distinct time periods. 
+As a proof of principle of our event detection strategy, we chose to 
+analyze tweets before, during, and after the National Football League's 
+2015 Super Bowl. We fitted latent Dirichlet allocation models for each of five 
+distinct one-hour periods. The first period began approximately 48h 
+before the Super Bowl halftime show. Subsequent time periods started at -24, 
+0, +24, and +48 hours after the approximate start time of the halftime show. 
+We defined each time period to be a single collection, or corpus, of tweets. Each tweet 
+constituted a "document", in the terminology of text analysis. We then fitted latent 
+Dirichlet allocation models to each corpus.
+
+We used several criteria to evaluate latent Dirichlet allocation model fits, with 
+emphasis on choosing a reasonable number of topics per corpus.
+Our strategy involved both visualization and more quantitative approaches to model 
+evaluation. For every model, we created one word cloud per topic.
+
+We then inspected topic contents at each of the five time points.
+
 ```{r child = "results.Rmd"}

### 2020-05-10 Rmd/tweets.Rmd
updated tweets analysis code

@@
 rtweet::stream_tweets(
-  timeout = (60 * 5),
-  parse = TRUE,
+  timeout = (15),
+  parse = FALSE,
   file_name = paste0("../data/", lubridate::now(), "-tweets")
@@
 library(magrittr)
+source("https://gist.githubusercontent.com/JBGruber/dee4c44e7d38d537426f57ba1e4f84ab/raw/ab87bebb8d020c2f96c71a40a483dc96a4c80e54/recover_stream.R")
 ```
@@
 ```{r}
-tweets <- rtweet::parse_stream("../data/2020-04-09 16:31:44-tweets.json") %>%
+tweets <- recover_stream(path = "../data/2020-05-10 10:36:09-tweets.json", verbose = TRUE) %>%
   dplyr::filter(lang == "en") %>%
@@
 
+```{r}
+tweets_dtm <- word_counts %>%
+  tidytext::cast_dtm(status_id, word, n)
+```
+
+```{r}
+tweets_lda <- topicmodels::LDA(tweets_dtm, k = 8)
+```
+
+```{r}
+tweets_beta <- tidytext::tidy(tweets_lda, matrix = "beta")
+```
+
+```{r}
+tweets_top_terms <- tweets_beta %>%
+  dplyr::group_by(topic) %>%
+  dplyr::top_n(10, beta) %>%
+  dplyr::ungroup() %>%
+  dplyr::arrange(topic, -beta)
+```
+
+```{r}
+tweets_top_terms %>%
+  dplyr::mutate(term = reorder_within(term, beta, topic)) %>%
+  ggplot2::ggplot(ggplot2::aes(term, beta, fill = factor(topic))) +
+  ggplot2::geom_col(show.legend = FALSE) +
+  ggplot2::facet_wrap(~ topic, scales = "free") +
+  ggplot2::coord_flip()
+```
+
+```{r}
+tweets_gamma <- tidytext::tidy(tweets_lda, matrix = "gamma")
+tweets_gamma
+```
+
+```{r, eval = FALSE}
+# too many documents to plot all
+tweets_gamma %>%
+  dplyr::mutate(title = reorder(document, gamma * topic)) %>%
+  ggplot2::ggplot(ggplot2::aes(factor(topic), gamma)) +
+  ggplot2::geom_boxplot() +
+  ggplot2::facet_wrap(~ document)
+```
+
 ### Colophon

### 2020-05-15 Rmd/tweets.Rmd
updated tweets stop list

@@
 # create a twitter-specific stop word list
-twitter_stop <- tibble::tibble(word = c("https", "http", "t.co"))
+twitter_stop <- tibble::tibble(word = c("https", "http", "t.co", "amp"))
 # get a list of english language words and names

### 2020-05-15 Rmd/tweets.Rmd
updated tweets collection

--- 2020-05-10 Rmd/tweets.Rmd
@@
 1. Examine structure of JSON obtained from api.
-1. Create phony JSON to represent a collection of tweets.
-    1. Use lorem ipsum for text of tweets
 
@@
 
-```{r}
-rtweet::stream_tweets(
-  timeout = (15),
-  parse = FALSE,
-  file_name = paste0("../data/", lubridate::now(), "-tweets")
-)
-```
-
 ## Parse json file
@@
 ```{r}
-tweets <- recover_stream(path = "../data/2020-05-10 10:36:09-tweets.json", verbose = TRUE) %>%
-  dplyr::filter(lang == "en") %>%
-  tidytext::unnest_tokens(output = word, input = text)
+tweets <- recover_stream(path = "../data/2020-05-10 12:00:01-tweets.json", verbose = TRUE) %>%
+  dplyr::mutate(status_id = as.character(status_id))
+```
+
+## Dehydrate tweets for sharing
+
+```{r}
+broken_ids <- readr::read_csv("broken_tweets.txt", col_names = FALSE) %>%
+  dplyr::rename(status_id = X1) %>%
+  dplyr::mutate(status_id = format(status_id, scientific = FALSE))
+readr::write_csv(tibble::tibble(status_id = c(tweets$status_id, broken_ids$status_id)), col_names = FALSE, path = "../data/2020-05-10-status-ids.csv")
+```
+
+## Rehydrate tweets by querying Twitter
+
+```{r}
+# need multiple calls to get all tweets rehydrated
+tweets_rehydrated <- rtweet::lookup_tweets(statuses = tweets$status_id, parse = TRUE)
 ```
@@
 twitter_stop <- tibble::tibble(word = c("https", "http", "t.co"))
+remove_reg <- "&amp;|&lt;|&gt;"
 # get a list of english language words and names
 library(qdapDictionaries) # contains object GradyAugmented
+library(tidytext) # get stop_words
 ```
@@
 # https://www.tidytextmining.com/topicmodeling.html#library-heist
-library(tidytext)
-word_counts <- tweets %>% 
+# https://www.tidytextmining.com/twitter.html
+word_counts <- tweets_rehydrated %>% 
+  dplyr::filter(lang == "en") %>%
+  dplyr::filter(!stringr::str_detect(text, "^RT")) %>%
+  dplyr::mutate(text = stringr::str_remove_all(text, remove_reg)) %>%
+  tidytext::unnest_tokens(output = word, input = text) %>%
   dplyr::select(status_id, word) %>%
@@
 ```{r}
-tweets_lda <- topicmodels::LDA(tweets_dtm, k = 8)
+tweets_lda <- topicmodels::LDA(tweets_dtm, k = 15)
 ```

### 2020-05-15 Rmd/jse-2019-ms.Rmd
updated tweets analysis code

@@
 
-We used functions from the `rtweet` package to parse tweet JSON into a data frame.
-From there, we used `tidytext` R package functions to break the tweet text into individual words for latent Dirichlet allocation.
-We discarded commonly used "stop words" and emojis.
+We used functions from the `rtweet` package to parse tweet JSON into a data frame. From there, we used `tidytext` R package 
+functions to break the tweet text into individual words. We discarded commonly used "stop words" 
+and emojis.
 
@@
 
+### Latent Dirichlet allocation
+
+Latent Dirichlet allocation is a statistical method for inferring latent (unobservable) topics (or themes)
+from a collection (or corpus) of documents. It is a probabilistic, generative model 
+for a corpus of documents. It is generative in the following sense. We pretend that
+there's an imaginary process for creating documents in the corpus. For each document,
+we choose a discrete distribution over topics. For example, some Mother's Day tweets wish 
+mothers a happy celebration. This may constitute one topic in the corpus. 
+Having chosen a distribution over topics, we then select document words by first 
+drawing a topic from the distribution over topics, then drawing a word from the
+chosen topic. 
+Thus, a topic is technically defined as a distribution over words in a fixed vocabulary (or collection of words).
+
+The literature on latent Dirichlet allocation and related methods is vast, 
+and we won't attempt to review it here.
+
 ### Study design
@@
 As a proof of principle of our event detection strategy, we chose to 
-analyze tweets before, during, and after the National Football League's 
-2015 Super Bowl. We fitted latent Dirichlet allocation models for each of five 
-distinct one-hour periods. The first period began approximately 48h 
-before the Super Bowl halftime show. Subsequent time periods started at -24, 
-0, +24, and +48 hours after the approximate start time of the halftime show. 
-We defined each time period to be a single collection, or corpus, of tweets. Each tweet 
-constituted a "document", in the terminology of text analysis. We then fitted latent 
+analyze tweets during and after Mother's Day 2020. We fitted latent Dirichlet allocation models for each of four 
+distinct five-minute periods. The first period began at noon Eastern time on Mother's Day 2020. Subsequent time periods started 
+24, 48, and 72 hours later. 
+We defined each time period to be a single collection, or corpus, of tweets. We then fitted latent 
 Dirichlet allocation models to each corpus.
@@
 
-We then inspected topic contents at each of the five time points.
+We then inspected topic contents at each of the four time points.
 

### 2020-05-15 Rmd/tweets.Rmd

@@
 # create a twitter-specific stop word list
-twitter_stop <- tibble::tibble(word = c("https", "http", "t.co"))
 remove_reg <- "&amp;|&lt;|&gt;"
+twitter_stop <- tibble::tibble(word = c("https", "http", "t.co", "amp"))
 # get a list of english language words and names

### 2020-05-16 Rmd/jse-2019-ms.Rmd
removed section numbers; fixed intro text

@@
     toc: FALSE
-    number_sections: TRUE
+    number_sections: FALSE
     includes:
@@
 current events, sports, media, and many other topics. 
-Recent development of accessible statistical methods for text 
-analysis now enable mentors to use tweets as contemporary pedagogical 
+Recent development of accessible statistical methods for large-scale text 
+analysis now enable instructors to use tweets as contemporary pedagogical 
 tools in guiding undergraduate research projects.
+We report one instance of a mentored text analysis research project. 
+We share our data and computer code to encourage others to undertake 
+tweet text analysis research. We also describe our methods for creating a 
+collection of tweets.
 
@@
 
-Using large collections of tweets, researchers have studied
-inference of relationships and social networks among Twitter users [@lin2011joint]; 
+Using large collections of tweets, scholars have studied diverse research questions, including
+the inference of relationships and social networks among Twitter users [@lin2011joint]; 
 authorship of specific tweets when multiple persons share a single account [@drob]; and 
@@
 
-In line with recent calls for students to work with real data [@nolan2010computing], our collection of tweets has served as a valuable resource in our mentoring 
+In line with recent calls for students to work with real data [@nolan2010computing], our 
+collection of tweets has served as a valuable resource in our mentoring 
 of undergraduate data science research. 
@@
 but also in related data science skills, including data 
-transfer from online sources, data storage, and using data 
-from multiple file formats. Collaboratively asking and addressing novel questions with 
+transfer from online sources, data storage, using data 
+from multiple file formats, and communicating findings and their limitations. 
+Collaboratively asking and addressing novel questions with 
 our collection of tweets gave mentored students 
@@
 and 4) interest in graduate training [@baker2010beyond;@higgins2001reconceptualizing].
-We strived to develop trusting, mutually respectful mentoring relationships with 
-our students while advising their senior projects. With input from us,
-the students selected appealing research questions.
 
 While our tweet collection enables us to address many possible research questions, the 
-dynamic content of tweets over time particularly intrigued us. We 
-hypothesized that high-profile social media events would generate a high volume of 
+dynamic content of tweets over time particularly piqued our interest. Together, students and 
+mentors hypothesized that high-profile social media events would generate a high volume of 
 tweets, and that we'd detect social media events through changes in tweet 
@@
 new account on Twitter. With these user credentials, we used the 
-R package `rtweet` to query the API. Because we work on computers 
-with linux operating systems, we use the linux software 
-`crontab` to repeatedly execute R code to submit API queries. Each 
+R package `rtweet` to query the API. Because we work with linux operating systems, we use the `crontab` software 
+to repeatedly execute R code to submit API queries. Each 
 query lasts a user-specified duration. We time the API queries so 
@@
 provides functions that parse tweet JSON to R data frames. We then 
-conducted all further analyses with the R data frames.
+conducted all further analyses in R.
 
@@
 **PLACE TWEET JSON HERE**
+{"created_at":"Mon May 11 15:59:57 +0000 2020","id":1259875904551038978,"id_str":"1259875904551038978","text":"Ryan has done a great job (and is also a fantastic runner) follow him and stay informed!","source":"\u003ca href=\"http:\/\/twitter.com\/download\/iphone\" rel=\"nofollow\"\u003eTwitter for iPhone\u003c\/a\u003e","truncated":false,"in_reply_to_status_id":null,"in_reply_to_status_id_str":null,"in_reply_to_user_id":null,"in_reply_to_user_id_str":null,"in_reply_to_screen_name":null,"user":{"id":2826587136,"id_str":"2826587136","name":"David, Mr.
+Priestap","screen_name":"dvdpeters","location":"Stolen Tonkowa Land","url":"https:\/\/tinyurl.com\/y6cptytj","description":"Penitent IRQ Veteran.
+Vicar of Pf.
+Waves at dogs #gothforgod #WAT Writes @guardianus @churchpubinc @oxunipress & @Ermenfrid.
+he\/him.
+#See my 1\u2b50\ufe0f reviews link\ud83d\udc47","translator_type":"none","protected":false,"verified":false,"followers_count":6271,"friends_count":1407,"listed_count":59,"favourites_count":347917,"statuses_count":61450,"created_at":"Mon Sep 22 16:17:34 +0000 2014","utc_offset":null,"time_zone":null,"geo_enabled":true,"lang":null,"contributors_enabled":false,"is_translator":false,"profile_background_color":"000000","profile_background_image_url":"http:\/\/abs.twimg.com\/images\/themes\/theme1\/bg.png","profile_background_image_url_https":"https:\/\/abs.twimg.com\/images\/themes\/theme1\/bg.png","profile_background_tile":false,"profile_link_color":"981CEB","profile_sidebar_border_color":"000000","profile_sidebar_fill_color":"000000","profile_teiestap","screen_name":"dvdpeters","location":"Stolen Tonkowa Land","url":"https:\/\/tinyurl.com\/y6cptytj","description":"Penitent IRQ Veteran.
+Vicar of Pf.
+Waves at dogs #gothforgod #WAT Writes @guardianus @churchpubinc @oxunipress & @Ermenfrid.
+he\/him.
+#See my 1\u2b50\ufe0f reviews link\ud83d\udc47","translator_type":"none","protected":false,"verified":false,"followers_count":6271,"friends_count":1407,"listed_count":59,"favourites_count":347917,"statuses_count":61450,"created_at":"Mon Sep 22 16:17:34 +0000 2014","utc_offset":null,"time_zone":null,"geo_enabled":true,"lang":null,"contributors_enabled":false,"is_translator":false,"profile_background_color":"000000","profile_background_image_url":"http:\/\/abs.twimg.com\/images\/themes\/theme1\/bg.png","profile_background_image_url_https":"https:\/\/abs.twimg.com\/images\/themes\/theme1\/bg.png","profile_background_tile":false,"profile_link_color":"981CEB","profile_sidebar_border_color":"000000","profile_sidebar_fill_color":"000000","profile_text_color":"000000","profile_use_background_image":false,"profile_image_url":"http:\/\/pbs.twimg.com\/profile_images\/1251909563894370311\/WTOcxogh_normal.jpg","profile_image_url_https":"https:\/\/pbs.twimg.com\/profile_images\/1251909563894370311\/WTOcxogh_normal.jpg","profile_banner_url":"https:\/\/pbs.twimg.com\/profile_banners\/2826587136\/1565768901","default_profile":false,"default_profile_image":false,"following":null,"follow_request_sent":null,"notifications":null},"geo":null,"coordinates":null,"place":null,"contributors":null,"quoted_status_id":1259874118251483136,"quoted_status_id_str":"1259874118251483136","quoted_status":{"created_at":"Mon May 11 15:52:
+51 +0000 2020","id":1259874118251483136,"id_str":"1259874118251483136","text":"Professional announcement: after 
+nearly four years of covering local courts for the Statesman, I soon will transiti\u2026 https:\/\/t.co\/zppt42m
+pwV","source":"\u003ca href=\"https:\/\/mobile.twitter.com\" rel=\"nofollow\"\u003eTwitter Web App\u003c\/a\u003
+e","truncated":true,"in_reply_to_status_id":null,"in_reply_to_status_id_str":null,"in_reply_to_user_id":null,"in
+_reply_to_user_id_str":null,"in_reply_to_screen_name":null,"user":{"id":42271065,"id_str":"42271065","name":"Rya
+n Autullo","screen_name":"AutulloAAS","location":"Austin, Texas","url":"http:\/\/statesman.com","description":"R
+eporter for the Austin American-Statesman covering courts and criminal justice. After hours: scolding scooter ri
+ders on the hike-and-bike trail.","translator_type":"none","protected":false,"verified":false,"followers_count":
 

### 2020-05-16 Rmd/jse-2019-ms.Rmd
deleted JSON

--- 2020-05-15 Rmd/jse-2019-ms.Rmd
@@
 **PLACE TWEET JSON HERE**
-
 ### Parsing text of tweets

### 2020-05-18 Rmd/jse-2019-ms.Rmd
added text about crontab

@@
 
+Setting up the query task with `crontab` is straightforward.
+On our computer, with Ubuntu 20.04 linux operating system, we opened a terminal and typed `crontab -e`.
+This opened a text file containing user-specified tasks.
+We added the following line to the bottom of the file:
+
+```{bash}
+*/5 * * * * R -e 'rtweet::stream_tweets(timeout = (60 * 5), parse = FALSE, file_name = paste0("~/work/mentoring/mentoring-framework/data/", lubridate::now(), "-tweets"))'
+```
+
+Users may need to slightly amend the above line to conform to requirements of their operating system's `crontab`.
+
 ### Querying Twitter API to get complete tweets

### 2020-05-18 Rmd/tweets.Rmd
added text about crontab

@@
 ```{r}
-tweets <- recover_stream(path = "../data/2020-05-10 12:00:01-tweets.json", verbose = TRUE) %>%
-  dplyr::mutate(status_id = as.character(status_id))
+tweets <- recover_stream(path = "../data/2020-05-10 12:00:01-tweets.json", verbose = TRUE) 
+tweets %>%
+  dplyr::select(status_id) %>%
+  readr::write_csv(path = "../data/tweets-status-ids-2020-05-10.csv", col_names = FALSE)
 ```
@@
   dplyr::rename(status_id = X1) %>%
-  dplyr::mutate(status_id = format(status_id, scientific = FALSE))
-readr::write_csv(tibble::tibble(status_id = c(tweets$status_id, broken_ids$status_id)), col_names = FALSE, path = "../data/2020-05-10-status-ids.csv")
+  dplyr::mutate(status_id = trimws(format(status_id, scientific = FALSE)))
 ```
 
-## Rehydrate tweets by querying Twitter
+## Hydrate tweets by querying Twitter
 
 ```{r}
+ids <- readr::read_csv("../data/tweets-status-ids-2020-05-10.csv", col_names = FALSE) %>%
+  dplyr::rename(status_id = X1) %>%
+  dplyr::mutate(status_id = as.character(status_id))
 # need multiple calls to get all tweets rehydrated
-tweets_rehydrated <- rtweet::lookup_tweets(statuses = tweets$status_id, parse = TRUE)
+tweets_rehydrated1 <- rtweet::lookup_tweets(statuses = ids$status_id, 
+                                            parse = TRUE)
+tweets_rehydrated2 <- rtweet::lookup_tweets(statuses = broken_ids$status_id, parse = TRUE)
+tweets_rehydrated <- tweets_rehydrated1 %>%
+  dplyr::bind_rows(tweets_rehydrated2)
 ```

### 2020-05-18 Rmd/jse-2019-ms.Rmd
deleted a mentoring paragraph

@@
 
-Mentoring in the work place and in higher education has many 
-benefits, including improving 1) students' development as thinkers and scholars, 2) confidence in their 
-own abilities, 3) integration into the campus community,
-and 4) interest in graduate training [@baker2010beyond;@higgins2001reconceptualizing].
-
 While our tweet collection enables us to address many possible research questions, the 
@@
 
-**PLACE TWEET JSON HERE**
 ### Parsing text of tweets
 
-We used functions from the `rtweet` package to parse tweet JSON into a data frame. From there, we used `tidytext` R package 
+We used functions from the `rtweet` R package to parse tweet JSON into a data frame. From there, we used `tidytext` R package 
 functions to break the tweet text into individual words. We discarded commonly used "stop words" 

### 2020-05-18 Rmd/jse-2019-ms.Rmd
switched csl file for citations

@@
 bibliography: jse.bib
+csl: chicago-annotated-bibliography.csl
 geometry: "left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm"
@@
 new account on Twitter. With these user credentials, we used the 
-R package `rtweet` to query the API. Because we work with linux operating systems, we use the `crontab` software 
+R package `rtweet` to query the API. Because we work with linux operating systems, we used the `crontab` software 
 to repeatedly execute R code to submit API queries. Each 
-query lasts a user-specified duration. We time the API queries so 
-that there is no time lag between queries. We store API query 
-results in their native JSON format. The R package `rtweet` 
+query lasted five minutes. We timed the API queries so 
+that there was no time lag between queries. We stored tweets resulting from API queries in their native JSON format.
+
+The R package `rtweet` 
 provides functions that parse tweet JSON to R data frames. We then 
@@
 
-Setting up the query task with `crontab` is straightforward.
-On our computer, with Ubuntu 20.04 linux operating system, we opened a terminal and typed `crontab -e`.
-This opened a text file containing user-specified tasks.
-We added the following line to the bottom of the file:
+Setting up the query task with `crontab` is straightforward. On our computer, with 
+Ubuntu 20.04 linux operating system, we opened a terminal and typed `crontab -e`. This 
+opened a text file containing user-specified tasks. We added the following line to the 
+bottom of the file:
 
-```{bash}
-*/5 * * * * R -e 'rtweet::stream_tweets(timeout = (60 * 5), parse = FALSE, file_name = paste0("~/work/mentoring/mentoring-framework/data/", lubridate::now(), "-tweets"))'
+```{bash, eval = FALSE}
+*/5 * * * * R -e 'rtweet::stream_tweets(timeout = (60 * 5), 
+parse = FALSE, file_name = paste0("~/work/mentoring/mentoring-framework/data/",
+lubridate::now(), "-tweets"))'
 ```
@@
 
-We queried a tweets database, created with the methods described above, to get ID numbers for 
-tweets from the desired time periods. We then submitted API queries to Twitter
-to get the full content of the tweets, including the tweet text.
+Twitter API use agreements forbid users from sharing complete API query results. 
+However, Twitter enables users to share tweet identification numbers. A user can 
+then query a Twitter API to obtain complete tweet data. In our experience, this 
+process is incomplete; that is, many tweets submitted to the Twitter API return 
+no data. Additionally, on repeated querying of the API, different sets of tweets 
+return data. This complicates our goal of making all analyses computationally 
+reproducible.
+
+From our collection of tweets, we chose to analyze those sent on five consecutive 
+days from May 10, 2020 to May 14, 2020. We then submitted API queries to Twitter
+to get the full content of tweets, including the tweet text.
 We provide below the R code that we used to query the Twitter API to obtain full tweet content.
-
-```{r, eval = FALSE}
-rtweet::lookup_tweets()
-```
 
@@
 
-Tweets are available as Javascript Object Notation (JSON) objects.
-Every tweet consists of multiple named fields, each of which is a key-value pair.
-The number of fields per tweet depends on user settings, retweet status, and other factors.
+Tweets are available as Javascript Object Notation (JSON) objects. Every tweet consists 
+of multiple named fields, each of which is a key-value pair. The number of fields per 
+tweet depends on user settings, retweet status, and other factors [@tweet_json].
+
+```{bash, eval = FALSE}
+{
+  "created_at": "Thu Apr 06 15:24:15 +0000 2017",
+  "id_str": "850006245121695744",
+  "text": "1\/ Today we\u2019re sharing our vision for the future of the Twitter API platform!\nhttps:\/\/t.co\/XweGngmxlP",
+  "user": {
+    "id": 2244994945,
+    "name": "Twitter Dev",
+    "screen_name": "TwitterDev",
+    "location": "Internet",
+    "url": "https:\/\/dev.twitter.com\/",
+    "description": "Your official source for Twitter Platform news, updates & events. 
+    Need technical help? Visit https:\/\/twittercommunity.com\/ \u2328\ufe0f 
+    #TapIntoTwitter"
+  },
+  "place": {   
+  },
+  "entities": {
+    "hashtags": [      
+    ],
+    "urls": [
+      {
+        "url": "https:\/\/t.co\/XweGngmxlP",
+        "unwound": {
+          "url": "https:\/\/cards.twitter.com\/cards\/18ce53wgo4h\/3xo1c",
+          "title": "Building the Future of the Twitter API Platform"
+        }
+      }
+    ],
+    "user_mentions": [     
+    ]
+  }
+}
+```
 
@@
 Latent Dirichlet allocation is a statistical method for inferring latent (unobservable) topics (or themes)
-from a collection (or corpus) of documents. It is a probabilistic, generative model 
-for a corpus of documents. It is generative in the following sense. We pretend that
+from a collection (or corpus) of documents. We pretend that
 there's an imaginary process for creating documents in the corpus. For each document,

### 2020-05-19 Rmd/tweets-one.Rmd
copied recover_tweets code to .R file

--- 2020-05-18 Rmd/tweets.Rmd
@@
----
-title: "Tweets"
-author: "Frederick J. Boehm"
-date: "4/9/2020"
-output: html_document
----
-
-## Goals
-
-1. Write a R script to query twitter streaming api to get 1 percent sample over 5'.
-1. Demonstrate how to set up a crontab to execute script every 5'.
-1. Verify that crontab approach works by running it on my computer for, say, 15'.
-    1. Cautions about twitter databases: deletions and amendments to tweets
-1. Examine structure of JSON obtained from api.
-
-## R code to query twitter streaming api
-
-Install `rtweet` package from CRAN.
-
-```{r, eval = FALSE}
-install.packages("rtweet")
-install.packages("lubridate")
-```
-
-## Parse json file
-
 ```{r}
-library(magrittr)
-source("https://gist.githubusercontent.com/JBGruber/dee4c44e7d38d537426f57ba1e4f84ab/raw/ab87bebb8d020c2f96c71a40a483dc96a4c80e54/recover_stream.R")
-```
-
-```{r}
-tweets <- recover_stream(path = "../data/2020-05-10 12:00:01-tweets.json", verbose = TRUE) 
+tweets <- recover_stream(path = "{{tweet_json_file}}", verbose = TRUE) 
 tweets %>%
   dplyr::select(status_id) %>%
-  readr::write_csv(path = "../data/tweets-status-ids-2020-05-10.csv", col_names = FALSE)
+  readr::write_csv(path = paste0("../data/tweets-status-ids-", "{{date}}", ".csv"), 
+                   col_names = FALSE)
 ```
@@
 ```{r}
-ids <- readr::read_csv("../data/tweets-status-ids-2020-05-10.csv", col_names = FALSE) %>%
+ids <- readr::read_csv(paste0("../data/tweets-status-ids-", "{{date}}", ".csv"), col_names = FALSE) %>%
   dplyr::rename(status_id = X1) %>%
   dplyr::mutate(status_id = as.character(status_id))
-# need multiple calls to get all tweets rehydrated
 tweets_rehydrated1 <- rtweet::lookup_tweets(statuses = ids$status_id, 
@@
   dplyr::bind_rows(tweets_rehydrated2)
+```
+
+```{r}
+# save ids of tweets that we actually used
+readr::write_csv(tweets_rehydrated$status_id, path = paste0("../data/", "{{date}}", "-analyzed-tweet-ids.csv"), col_names = FALSE)
 ```
@@
 ```{r}
-tweets_lda <- topicmodels::LDA(tweets_dtm, k = 15)
+tweets_lda <- topicmodels::LDA(tweets_dtm, k = 10)
 ```
@@
 
-### Colophon
-
-This report was generated on `r Sys.time()` using the following computational environment and dependencies:
-
-```{r colophon, cache = FALSE}
-# which R packages and versions?
-if ("devtools" %in% installed.packages()) devtools::session_info()
+```{r}
+Sys.sleep(15 * 60) # sleep for 15 minutes to allow api limits to reset
 ```
-
-The current Git commit details are:
-
-```{r}
-# what commit is this file at? 
-if ("git2r" %in% installed.packages() & git2r::in_repository(path = ".")) git2r::repository(here::here())  
-```

### 2020-05-19 Rmd/tweets.Rmd
ignored json files and other data files
switched csl file for citations
copied recover_tweets code to .R file

--- 2020-05-15 Rmd/tweets.Rmd
@@
 library(magrittr)
-source("https://gist.githubusercontent.com/JBGruber/dee4c44e7d38d537426f57ba1e4f84ab/raw/ab87bebb8d020c2f96c71a40a483dc96a4c80e54/recover_stream.R")
+source("recover_tweets.R")
 ```
@@
 ```{r}
-tweets <- recover_stream(path = "../data/2020-05-10 12:00:01-tweets.json", verbose = TRUE) %>%
-  dplyr::mutate(status_id = as.character(status_id))
+date <- c("2020-05-10", "2020-05-11", "2020-05-12", "2020-05-13", "2020-05-14")
+tweet_json_file <- c("../data/2020-05-10 12:00:01-tweets.json",
+  "../data/2020-05-11 12:00:01-tweets.json", 
+  "../data/2020-05-12 12:00:02-tweets.json",
+  "../data/2020-05-13 12:00:02-tweets.json", 
+  "../data/2020-05-14 12:00:02-tweets.json")
+out <- purrr::pmap_dfc(tibble::tibble(date, tweet_json_file),
+                       .f = function(date, tweet_json_file){
+                         knitr::knit_expand(text = readLines("tweets-one.Rmd"))
+                       })
 ```
 
-## Dehydrate tweets for sharing
+## May 10, 2020
 
-```{r}
-broken_ids <- readr::read_csv("broken_tweets.txt", col_names = FALSE) %>%
-  dplyr::rename(status_id = X1) %>%
-  dplyr::mutate(status_id = format(status_id, scientific = FALSE))
-readr::write_csv(tibble::tibble(status_id = c(tweets$status_id, broken_ids$status_id)), col_names = FALSE, path = "../data/2020-05-10-status-ids.csv")
-```
+`r knitr::knit(text = out[[1]])`
 
-## Rehydrate tweets by querying Twitter
+## May 11, 2020
 
-```{r}
-# need multiple calls to get all tweets rehydrated
-tweets_rehydrated <- rtweet::lookup_tweets(statuses = tweets$status_id, parse = TRUE)
-```
-
-## Prepare document-term matrix from tweet words
-
-```{r}
-# create a twitter-specific stop word list
-twitter_stop <- tibble::tibble(word = c("https", "http", "t.co"))
-remove_reg <- "&amp;|&lt;|&gt;"
-# get a list of english language words and names
-library(qdapDictionaries) # contains object GradyAugmented
-library(tidytext) # get stop_words
-```
-
-```{r}
-# https://www.tidytextmining.com/topicmodeling.html#library-heist
-# https://www.tidytextmining.com/twitter.html
-word_counts <- tweets_rehydrated %>% 
-  dplyr::filter(lang == "en") %>%
-  dplyr::filter(!stringr::str_detect(text, "^RT")) %>%
-  dplyr::mutate(text = stringr::str_remove_all(text, remove_reg)) %>%
-  tidytext::unnest_tokens(output = word, input = text) %>%
-  dplyr::select(status_id, word) %>%
-  dplyr::anti_join(stop_words) %>%
-  dplyr::anti_join(twitter_stop) %>%
-  dplyr::filter(word %in% GradyAugmented) %>%
-  dplyr::count(status_id, word, sort = TRUE) %>%
-  dplyr::ungroup()
-```
-
-```{r}
-tweets_dtm <- word_counts %>%
-  tidytext::cast_dtm(status_id, word, n)
-```
-
-```{r}
-tweets_lda <- topicmodels::LDA(tweets_dtm, k = 15)
-```
-
-```{r}
-tweets_beta <- tidytext::tidy(tweets_lda, matrix = "beta")
-```
-
-```{r}
-tweets_top_terms <- tweets_beta %>%
-  dplyr::group_by(topic) %>%
-  dplyr::top_n(10, beta) %>%
-  dplyr::ungroup() %>%
-  dplyr::arrange(topic, -beta)
-```
-
-```{r}
-tweets_top_terms %>%
-  dplyr::mutate(term = reorder_within(term, beta, topic)) %>%
-  ggplot2::ggplot(ggplot2::aes(term, beta, fill = factor(topic))) +
-  ggplot2::geom_col(show.legend = FALSE) +
-  ggplot2::facet_wrap(~ topic, scales = "free") +
-  ggplot2::coord_flip()
-```
-
-```{r}
-tweets_gamma <- tidytext::tidy(tweets_lda, matrix = "gamma")
-tweets_gamma
-```
-
-```{r, eval = FALSE}
-# too many documents to plot all
-tweets_gamma %>%
-  dplyr::mutate(title = reorder(document, gamma * topic)) %>%
-  ggplot2::ggplot(ggplot2::aes(factor(topic), gamma)) +
-  ggplot2::geom_boxplot() +
-  ggplot2::facet_wrap(~ document)
-```
+`r knitr::knit(text = out[[2]])`
 

### 2020-05-19 Rmd/recover_tweets.R

@@ -1,0 +1,68 @@
+#' Recovers Twitter damaged stream data (JSON file) into parsed data frame.
+#'
+#' @param path Character, name of JSON file with data collected by
+#'   \code{\link{stream_tweets}}.
+#' @param dir Character, name of a directory where intermediate files are
+#'   stored.
+#' @param verbose Logical, should progress be displayed?
+#' 
+#' @family stream tweets
+# https://gist.githubusercontent.com/JBGruber/dee4c44e7d38d537426f57ba1e4f84ab/raw/ab87bebb8d020c2f96c71a40a483dc96a4c80e54/recover_stream.R
+recover_stream <- function(path, dir = NULL, verbose = TRUE) {
+  
+  # read file and split to tweets
+  lines <- readChar(path, file.info(path)$size, useBytes = TRUE)
+  tweets <- stringi::stri_split_fixed(lines, "\n{")[[1]]
+  tweets[-1] <- paste0("{", tweets[-1])
+  tweets <- tweets[!(tweets == "" | tweets == "{")]
+  
+  # remove misbehaving characters
+  tweets <- gsub("\r", "", tweets, fixed = TRUE)
+  tweets <- gsub("\n", "", tweets, fixed = TRUE)
+  
+  # write tweets to disk and try to read them in individually
+  if (is.null(dir)) {
+    dir <- paste0(tempdir(), "/tweets/")
+    dir.create(dir, showWarnings = FALSE)
+  }
+  
+  if (verbose) {
+    pb <- progress::progress_bar$new(
+      format = "Processing tweets [:bar] :percent, :eta remaining",
+      total = length(tweets), clear = FALSE
+    )
+    pb$tick(0)
+  }
+  
+  tweets_l <- lapply(tweets, function(t) {
+    pb$tick()
+    id <- unlist(stringi::stri_extract_first_regex(t, "(?<=id\":)\\d+(?=,)"))[1]
+    f <- paste0(dir, id, ".json")
+    writeLines(t, f, useBytes = TRUE)
+    out <- tryCatch(rtweet::parse_stream(f),
+                    error = function(e) {})
+    if ("tbl_df" %in% class(out)) {
+      return(out)
+    } else {
+      return(id)
+    }
+  })
+  
+  # test which ones failed
+  test <- vapply(tweets_l, is.character, FUN.VALUE = logical(1L))
+  bad_files <- unlist(tweets_l[test])
+  
+  # Let user decide what to do
+  if (length(bad_files) > 0) {
[…]

### 2020-05-20 Rmd/tweets-one.Rmd
added files produced from tweets analysis

@@
-```{r}
+```{r, recover-{{date}}}
 tweets <- recover_stream(path = "{{tweet_json_file}}", verbose = TRUE) 
@@
 
-```{r}
+```{r, broken-{{date}}}
 broken_ids <- readr::read_csv("broken_tweets.txt", col_names = FALSE) %>%
@@
 
-```{r}
+```{r, ids-{{date}}}
 ids <- readr::read_csv(paste0("../data/tweets-status-ids-", "{{date}}", ".csv"), col_names = FALSE) %>%
@@
 
-```{r}
+```{r, saveids-{{date}}}
 # save ids of tweets that we actually used
-readr::write_csv(tweets_rehydrated$status_id, path = paste0("../data/", "{{date}}", "-analyzed-tweet-ids.csv"), col_names = FALSE)
+readr::write_csv(tibble::as_tibble(tweets_rehydrated$status_id), path = paste0("../data/", "{{date}}", "-analyzed-tweet-ids.csv"), col_names = FALSE)
 ```
@@
 
-```{r}
+```{r, stops-{{date}}}
 # create a twitter-specific stop word list
@@
 
-```{r}
+```{r, wordcounts-{{date}}}
 # https://www.tidytextmining.com/topicmodeling.html#library-heist
@@
 
-```{r}
+```{r, cast-{{date}}}
 tweets_dtm <- word_counts %>%
@@
 
-```{r}
-tweets_lda <- topicmodels::LDA(tweets_dtm, k = 10)
+```{r, lda-{{date}}}
+tweets_lda <- topicmodels::LDA(tweets_dtm, k = 8)
 ```
 
-```{r}
+```{r, beta-{{date}}}
 tweets_beta <- tidytext::tidy(tweets_lda, matrix = "beta")
@@
 
-```{r}
+```{r, topterms-{{date}}}
 tweets_top_terms <- tweets_beta %>%
@@
 
-```{r}
+```{r, plot-{{date}}}
 tweets_top_terms %>%
@@
 
-```{r}
+```{r, gamma-{{date}}}
 tweets_gamma <- tidytext::tidy(tweets_lda, matrix = "gamma")
@@
 
-```{r, eval = FALSE}
-# too many documents to plot all
-tweets_gamma %>%
-  dplyr::mutate(title = reorder(document, gamma * topic)) %>%
-  ggplot2::ggplot(ggplot2::aes(factor(topic), gamma)) +
-  ggplot2::geom_boxplot() +
-  ggplot2::facet_wrap(~ document)
-```
-
-```{r}
+```{r, sleep-{{date}}}
 Sys.sleep(15 * 60) # sleep for 15 minutes to allow api limits to reset

### 2020-05-20 Rmd/abstract.Rmd
upd
updates to methods
many updates
rewrote abstract

@@
 
-We develop
-a mentoring framework to guide undergraduate researchers
-through individualized research projects in social media data science. 
-Our framework involves research question formulation, data acquisition, data 
-analysis and visualization, and presentation and communication of results. Our two honors students, 
-whose projects serve as case studies for our framework, completed all components of the 
-individualized research projects. We found that data science research skills, self-confidence in 
-research ability, and professional interest in data science increased for both 
-students. We describe our successes, lessons learned, and ideas for others to 
-build similar frameworks.
+We describe methods for collecting tweets from Twitter. With these methods, we 
+curate a large collection of tweets.
+We present methods for social media event detection through text analysis of tweets. 
+We describe our use of tweets in mentored undergraduate research. 
+We also offer other directions for other text analysis of tweets.

### 2020-05-20 Rmd/jse-2019-ms.Rmd
rewrote abstract

@@
 mentors hypothesized that high-profile social media events would generate a high volume of 
-tweets, and that we'd detect social media events through changes in tweet 
+tweets, and that we would detect social media events through changes in tweet 
 topic content over time. We present below 1) an approach for collecting tweets in real 

### 2020-05-21 Rmd/jse-2019-ms.Rmd
updated tweets document

@@
 From our collection of tweets, we chose to analyze those sent on five consecutive 
-days from May 10, 2020 to May 14, 2020. We then submitted API queries to Twitter
+days from May 10, 2020 to May 14, 2020. To minimize the computing requirements, we 
+limited our study to tweets sent during a five-minute period (12:00pm to 12:05pm Eastern time) every day.
+We then submitted API queries to Twitter
 to get the full content of tweets, including the tweet text.
-We provide below the R code that we used to query the Twitter API to obtain full tweet content.
+In supplementary files, we provide the R code that we used to query the Twitter API to obtain full tweet content.
 
@@
 occurrences of a term in the specified document). We created a document by 
-term matrix with the R functions `` from the `` R package.
+term matrix with the R function `cast_dtm` from the `tidytext` package.
 

### 2020-05-21 Rmd/tweets-one.Rmd
updated tweets document

@@
 ```{r, recover-{{date}}}
+fn <- paste0("../data/", {{date}}, "-analyzed-tweet-ids.csv")
+tr_fn <- paste0("../data/", {{date}}, "-rehydrated-tweets.csv")
+if (!file.exists(fn)){
+
   tweets <- recover_stream(path = "{{tweet_json_file}}", verbose = TRUE) 
@@
                      col_names = FALSE)
-```
-
-## Dehydrate tweets for sharing
-
-```{r, broken-{{date}}}
   broken_ids <- readr::read_csv("broken_tweets.txt", col_names = FALSE) %>%
@@
     dplyr::mutate(status_id = trimws(format(status_id, scientific = FALSE)))
-```
-
-## Hydrate tweets by querying Twitter
-
-```{r, ids-{{date}}}
   ids <- readr::read_csv(paste0("../data/tweets-status-ids-", "{{date}}", ".csv"), col_names = FALSE) %>%
@@
     dplyr::bind_rows(tweets_rehydrated2)
-```
-
-```{r, saveids-{{date}}}
   # save ids of tweets that we actually used
-readr::write_csv(tibble::as_tibble(tweets_rehydrated$status_id), path = paste0("../data/", "{{date}}", "-analyzed-tweet-ids.csv"), col_names = FALSE)
+  readr::write_csv(tibble::as_tibble(tweets_rehydrated$status_id), path = fn, col_names = FALSE)
+  readr::write_csv(tweets_rehydrated, path = tr_fn)
+}
 ```
@@
 ## Prepare document-term matrix from tweet words
-
-```{r, stops-{{date}}}
-# create a twitter-specific stop word list
-remove_reg <- "&amp;|&lt;|&gt;"
-twitter_stop <- tibble::tibble(word = c("https", "http", "t.co", "amp"))
-# get a list of english language words and names
-library(qdapDictionaries) # contains object GradyAugmented
-library(tidytext) # get stop_words
-```
 
@@
 tweets_lda <- topicmodels::LDA(tweets_dtm, k = 8)
-```
-
-```{r, beta-{{date}}}
 tweets_beta <- tidytext::tidy(tweets_lda, matrix = "beta")
@@
 tweets_top_terms %>%
-  dplyr::mutate(term = reorder_within(term, beta, topic)) %>%
+  dplyr::mutate(term = tidytext::reorder_within(term, beta, topic)) %>%
   ggplot2::ggplot(ggplot2::aes(term, beta, fill = factor(topic))) +
@@
   ggplot2::facet_wrap(~ topic, scales = "free") +
-  ggplot2::coord_flip()
+  ggplot2::coord_flip() + 
+  theme(axis.text.x = element_text(angle = 270))
+
 ```
 
-```{r, gamma-{{date}}}
-tweets_gamma <- tidytext::tidy(tweets_lda, matrix = "gamma")
-tweets_gamma
+```{r, ggsave-{{date}}}
+ggplot2::ggsave(paste0("../results/beta-", {{date}}, ".png"))
 ```

### 2020-05-21 Rmd/tweets.Rmd
added files produced from tweets analysis
updated tweets document

@@
 1. Demonstrate how to set up a crontab to execute script every 5'.
-1. Verify that crontab approach works by running it on my computer for, say, 15'.
 1. Cautions about twitter databases: deletions and amendments to tweets
-1. Examine structure of JSON obtained from api.
 
@@
 source("recover_tweets.R")
+```
+
+```{r, stops}
+# create a twitter-specific stop word list
+remove_reg <- "&amp;|&lt;|&gt;"
+twitter_stop <- tibble::tibble(word = c("https", "http", "t.co", "amp"))
+# get a list of english language words and names
+library(qdapDictionaries) # contains object GradyAugmented
+library(tidytext) # get stop_words
 ```
@@
 
-### Colophon
+## May 12, 2020
+
+`r knitr::knit(text = out[[3]])`
+
+## May 13, 2020
+
+`r knitr::knit(text = out[[4]])`
+
+## May 14, 2020
+
+`r knitr::knit(text = out[[5]])`
+
+## Colophon
 

### 2020-05-21 Rmd/tweets-one.Rmd
cleaned data subdir

@@
 ```{r, recover-{{date}}}
-fn <- paste0("../data/", {{date}}, "-analyzed-tweet-ids.csv")
-tr_fn <- paste0("../data/", {{date}}, "-rehydrated-tweets.csv")
-if (!file.exists(fn)){
-
+fn <- paste0("../data/", "{{date}}", "-analyzed-tweet-ids.csv")
+#fn <- paste0("../data/", "2020-05-10", "-analyzed-tweet-ids.csv")
+tr_fn <- paste0("../data/", "{{date}}", "-rehydrated-tweets.csv")
+#tr_fn <- paste0("../data/", "2020-05-10", "-rehydrated-tweets.csv")
+if (!file.exists(tr_fn)){
   tweets <- recover_stream(path = "{{tweet_json_file}}", verbose = TRUE) 
+  #tweets <- recover_stream(path = "../data/2020-05-10 12:00:01-tweets.json", verbose = TRUE) 
+  id_fn <- paste0("../data/tweets-status-ids-", "{{date}}", ".csv")
   tweets %>%
     dplyr::select(status_id) %>%
-    readr::write_csv(path = paste0("../data/tweets-status-ids-", "{{date}}", ".csv"), 
+    readr::write_csv(path = id_fn, 
                      col_names = FALSE)
@@
     dplyr::mutate(status_id = trimws(format(status_id, scientific = FALSE)))
-  ids <- readr::read_csv(paste0("../data/tweets-status-ids-", "{{date}}", ".csv"), col_names = FALSE) %>%
+  ids <- readr::read_csv(id_fn, col_names = FALSE) %>%
     dplyr::rename(status_id = X1) %>%
@@
   readr::write_csv(tweets_rehydrated, path = tr_fn)
+} else {
+  tweets_rehydrated <- readr::read_csv(tr_fn)
 }
@@
 ```{r, ggsave-{{date}}}
-ggplot2::ggsave(paste0("../results/beta-", {{date}}, ".png"))
+ggplot2::ggsave(paste0("../results/beta-", "{{date}}", ".png"))
 ```

### 2020-05-24 Rmd/jse-2019-ms.Rmd
added placeholder for data dictionary table.
updated ms

@@
 output: 
-  bookdown::pdf_book:
+  pdf_document:
+    latex_engine: xelatex
     toc: FALSE
@@
 
+## Research mentoring considerations
+
+We collaboratively developed research goals with students through a series of discussions throughout the academic year.
+As trainees begin their senior research projects, we suggest that mentors discuss with them:
+
+### Student research interests and goals
+Student experience with data analysis and statistical analysis software
+Software for statistical analysis and best practices for reproducibility
+
+### Student research interests and goals
+
+In our experience, student interests vary, and their initial ability to articulate research goals may be limited.
+An initial brainstorming session may clarify their interests and encourage them to think critically about goals under the time constraints of their academic schedules.
+Additionally, we anticipate that sharing completed student project reports will guide student thinking about the scope of possible projects.
+
+We mentored one student whose interest in financial time series guided her project. A second student formulated a project around event detection from tweet time series.
+
+### Student experience with data analysis and statistical software
+
+### Software for data
+
+We used the linux software package `crontab` to automate the execution of R code every five minutes. The tweets are available from the Twitter API as JSON. We saved the JSON in plain text files.
+
+We performed all analyses in the R statistical computing environment.
+We used the Rstudio software package to interface with R.
+We share our R code to enhance reproducibility of our findings.
+Users may also find our code helpful when creating their own tweet collections.
+
+### Backward design
+
+We sought to give our trainees opportunities to formulate and address a data science research question. We consulted the GAISE report for additional guidance. 
+The GAISE report executive summary highlights six goals for statistics training:
+
+1. Teach statistical thinking  
+1. Focus on conceptual understanding  
+1. Integrate real data with a context and purpose  
+1. Foster active learning  
+1. Use technology to explore concepts and analyze data  
+1. Use assessments to improve and evaluate student learning
+
+### Learning outcomes
+
+### Time period
+
+We began collecting tweets in 2013.
+Our two students conducted their research projects during the 2015-2016 academic year.
+We recommend a full academic year for projects of this magnitude, although a one-semester project is possible.
+Our students presented their findings at the statistics department’s undergraduate poster session near the end of the 2015-2016 academic year.
+
 ## Methods
@@
 
-Tweets are available as Javascript Object Notation (JSON) objects. Every tweet consists 
-of multiple named fields, each of which is a key-value pair. The number of fields per 
+```{r, echo = FALSE, message = FALSE}
+data_dictionary <- readr::read_csv("../data/tweets-data-dictionary.csv")
+```
+
+Tweets are available from the Twitter API as Javascript Object Notation (JSON) objects. Every tweet consists 
+of multiple key-value pairs. The number of fields per 
 tweet depends on user settings, retweet status, and other factors [@tweet_json]. 
+The 31 tweet key-value pairs belong to
+`r length(unique(data_dictionary$Type))` distinct classes (Appendix 1). The 
+classes are either vectors - numeric, logical, or character - or arrays assembled 
+from the vector classes.
 
-```{bash, eval = FALSE}
+Below is an example of Tweet JSON. Every tweet features the keys "created\_at" (the time stamp), 
+"id\_str" (a unique tweet identifier), and "text". We use these three keys in 
+our analyses.
+
+```{bash, eval = FALSE, fig.cap = "Example Tweet JSON"}
 {
@@
 
-We used functions from the `rtweet` R package to parse tweet JSON into a data frame. From there, we used `tidytext` R package 
-functions to break the tweet text into individual words. We discarded commonly used "stop words" 
+We used functions from the `rtweet` R package to parse tweet JSON into a data frame. 
+With `rtweet` functions, we parsed JSON arrays into their component vectors and added them
+to the data frame.
+
+We then divided tweet text into words with functions from the `tidytext` R package. We discarded commonly used "stop words" 
 and emojis.
@@
 Latent Dirichlet allocation is a statistical method for inferring latent (unobservable) topics (or themes)
-from a collection (or corpus) of documents. We pretend that
+from a corpus (or collection) of documents. We pretend that
 there's an imaginary process for creating documents in the corpus. For each document,
@@
 
-```{r child = "results.Rmd"}
-```
+## References
 
-```{r child = "discussion.Rmd"}
-```
+<div id="refs"></div>
 
-## References
+\newpage
+
+## Appendix 1: Tweet data dictionary
+
+Twitter shares a data dictionary for tweets (https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object,  (Accessed: May 23, 2020)).
+We have saved it as a supplementary file, "tweets-data-dictionary.csv".

### 2020-05-24 Rmd/tweets-one.Rmd
updated ms

--- 2020-05-21 Rmd/tweets-one.Rmd
@@
 ```{r, recover-{{date}}}
-fn <- paste0("../data/", {{date}}, "-analyzed-tweet-ids.csv")
-tr_fn <- paste0("../data/", {{date}}, "-rehydrated-tweets.csv")
-if (!file.exists(fn)){
-
+fn <- paste0("../data/", "{{date}}", "-analyzed-tweet-ids.csv")
+#fn <- paste0("../data/", "2020-05-10", "-analyzed-tweet-ids.csv")
+tr_fn <- paste0("../data/", "{{date}}", "-rehydrated-tweets.rds")
+#tr_fn <- paste0("../data/", "2020-05-10", "-rehydrated-tweets.rds")
+if (!file.exists(tr_fn)){
   tweets <- recover_stream(path = "{{tweet_json_file}}", verbose = TRUE) 
+  #tweets <- recover_stream(path = "../data/2020-05-10 12:00:01-tweets.json", verbose = TRUE) 
+  id_fn <- paste0("../data/tweets-status-ids-", "{{date}}", ".csv")
+  #id_fn <- paste0("../data/tweets-status-ids-", "2020-05-10", ".csv")
   tweets %>%
     dplyr::select(status_id) %>%
-    readr::write_csv(path = paste0("../data/tweets-status-ids-", "{{date}}", ".csv"), 
+    readr::write_csv(path = id_fn, 
                      col_names = FALSE)
@@
   readr::write_csv(tibble::as_tibble(tweets_rehydrated$status_id), path = fn, col_names = FALSE)
-  readr::write_csv(tweets_rehydrated, path = tr_fn)
+  saveRDS(tweets_rehydrated, tr_fn)
+} else {
+  tweets_rehydrated <- readRDS(tr_fn)
 }
@@
 ```{r, lda-{{date}}}
-tweets_lda <- topicmodels::LDA(tweets_dtm, k = 8)
+tweets_lda <- topicmodels::LDA(tweets_dtm, k = 6)
 tweets_beta <- tidytext::tidy(tweets_lda, matrix = "beta")
@@
   ggplot2::coord_flip() + 
-  theme(axis.text.x = element_text(angle = 270))
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 270))
 

### 2020-05-25 Rmd/jse-special-issue-2020.Rmd
switched to using a new Rmd file

--- 2020-05-24 Rmd/jse-2019-ms.Rmd
@@
 ---
-title: 'Analyzing tweets to detect social media events'
+title: "What is happening on Twitter? Tweet analysis and undergraduate research projects"
 author: "Frederick J. Boehm and Bret M. Hanlon"
@@
 
+Last modified: `r params$lastmod`.
+
 ```{r child = "abstract.Rmd", eval = FALSE}
@@
 
-# Introduction
+## Introduction
 
 Twitter has profoundly changed how we communicate. 
-In only 280 characters, users can instantly contribute to public conversations on politics, 
+In only 280 characters, users instantly contribute to public conversations on politics, 
 current events, sports, media, and many other topics. 
@@
 topic content over time. We present below 1) an approach for collecting tweets in real 
-time and 2) statistical methods for detecting social media 
-events via latent Dirichlet allocation modeling of collections of tweets and 3) reflections on 
-using this data set in research mentoring of undergraduate students.
+time and 2) methods for detecting social media 
+events via latent Dirichlet allocation modeling of tweets and 3) suggestions for 
+using tweets in research mentoring of undergraduate students.
 
-## Research mentoring considerations
+## Preliminary research mentoring considerations
 
@@
 
-### Student research interests and goals
-Student experience with data analysis and statistical analysis software
-Software for statistical analysis and best practices for reproducibility
+1. Student experience with data analysis software
+1. Student research interests and goals
+1. Best practices for reproducibility
+
+### Student experience with data analysis software
+
+Student experience with data analysis software varies. In our statistics department, 
+most students learn elementary R computing skills through class assignments. Some 
+students, by concentrating in computer science, learn other data analysis software 
+packages, such as Python. 
+Those who do undergraduate statistics research often learn advanced topics in R
+computing, such as R package assembly, documentation, and testing. Many develop 
+expertise in linux computing and cluster computing, too.
 
@@
 
-### Student experience with data analysis and statistical software
-
-### Software for data
-
-We used the linux software package `crontab` to automate the execution of R code every five minutes. The tweets are available from the Twitter API as JSON. We saved the JSON in plain text files.
-
-We performed all analyses in the R statistical computing environment.
-We used the Rstudio software package to interface with R.
-We share our R code to enhance reproducibility of our findings.
-Users may also find our code helpful when creating their own tweet collections.
+### Best practices for reproducibility
 
@@
 
-We began collecting tweets in 2013.
 Our two students conducted their research projects during the 2015-2016 academic year. 
-We recommend a full academic year for projects of this magnitude, although a one-semester project is possible.
-Our students presented their findings at the statistics department’s undergraduate poster session near the end of the 2015-2016 academic year.
+We recommend a full academic year for projects of this magnitude, although a one-semester 
+project is possible. Our students presented their findings at the statistics department’s 
+undergraduate poster session near the end of the 2015-2016 academic year.
 
@@
 Twitter API use agreements forbid users from sharing complete API query results. 
-However, Twitter enables users to share tweet identification numbers. A user can 
+However, Twitter permits users to share tweet identification numbers. A user can 
 then query a Twitter API to obtain complete tweet data. In our experience, this 
@@
 to get the full content of tweets, including the tweet text.
-In supplementary files, we provide the R code that we used to query the Twitter API to obtain full tweet content.
+In supplementary files, we provide the R code that we used to 
+query the Twitter API to obtain full tweet content.
 
@@
   "id_str": "850006245121695744",
-  "text": "1\/ Today we\u2019re sharing our vision for the future of the Twitter API platform!\nhttps:\/\/t.co\/XweGngmxlP",
+  "text": "1\/ Today we\u2019re sharing our vision for the future of the Twitter API platform!",
   "user": {
@@
 We used functions from the `rtweet` R package to parse tweet JSON into a data frame. 
-With `rtweet` functions, we parsed JSON arrays into their component vectors and added them
+We divided JSON arrays into their component vectors and added them
 to the data frame.
@@
 
+### Results
+
+```{r, fig.cap = "Top terms for LDA model from May 10, 2020 (Mother's Day)"}
+knitr::include_graphics("../results/beta-2020-05-10.png")
+```
+
+```{r, fig.cap = "Top terms for LDA model from May 11, 2020"}
+knitr::include_graphics("../results/beta-2020-05-11.png")
+```
+
+## Acknowledgements
+
+The authors thank Betsy Colby Davie and Rick Nordheim for helpful discussions and 
+feedback on preliminary versions of the manuscript. We thank the special issue 
+editors and anonymous reviewers for their constructive comments and suggestions.
+
+## Supplementary Material
+
 ## References
@@
 We have saved it as a supplementary file, "tweets-data-dictionary.csv".
+
+## Appendix 2: LDA results for May 12, 13, and 14
+
+```{r, fig.cap = "Top terms for LDA model from May 12, 2020"}
+knitr::include_graphics("../results/beta-2020-05-12.png")
+```
+
+```{r, fig.cap = "Top terms for LDA model from May 13, 2020"}
+knitr::include_graphics("../results/beta-2020-05-13.png")
+```
+
+```{r, fig.cap = "Top terms for LDA model from May 14, 2020"}
+knitr::include_graphics("../results/beta-2020-05-14.png")
+```

### 2020-05-25 Rmd/jse-special-issue-2020.tex
switched to using a new Rmd file

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={What is happening on Twitter? Tweet analysis and undergraduate research projects},
  pdfauthor={Frederick J. Boehm and Bret M. Hanlon},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{lineno}
\linenumbers
\usepackage{setspace}
\doublespacing
\usepackage{caption}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
%\usepackage[utf8x]{inputenc}
\usepackage{rotating}
\usepackage{wrapfig}

\title{What is happening on Twitter? Tweet analysis and undergraduate research
projects}
\author{Frederick J. Boehm and Bret M. Hanlon}
\date{June 15, 2020}

\begin{document}
\maketitle

Last modified: 2020-05-24 19:58:05.

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Twitter has profoundly changed how we communicate. In only 280
characters, users instantly contribute to public conversations on
politics, current events, sports, media, and many other topics. Recent
development of accessible statistical methods for large-scale text
analysis now enable instructors to use tweets as contemporary
pedagogical tools in guiding undergraduate research projects. We report
one instance of a mentored text analysis research project. We share our
data and computer code to encourage others to undertake tweet text
analysis research. We also describe our methods for creating a
collection of tweets.

Some social media data, including tweets from Twitter, is available
through website application product interfaces (APIs). By way of a
streaming API, Twitter shares a sample of approximately one percent of
all tweets during an API query time period.\footnote{``Sampled Stream.''}
Any Twitter user can freely access this one percent sample, whereas
access to a larger selection is available to researchers for a fee.

Using large collections of tweets, scholars have studied diverse
research questions, including the inference of relationships and social
networks among Twitter users;\footnote{Lin et al., ``The Joint Inference
  of Topic Diffusion and Evolution in Social Communities.''} authorship
of specific tweets when multiple persons share a single
account;\footnote{Robinson, ``Text Analysis of Trump's Tweets Confirms
  He Writes Only the (Angrier) Android Half.''} and rhetoric in
recruiting political supporters.\footnote{Pelled et al., ```Little
  Marco,'`Lyin'Ted,'`Crooked Hillary,' and the `Biased' Media''; Wells
  et al., ``How Trump Drove Coverage to the Nomination.''} Recognizing
the potential utility of tweets for data science research and teaching,
we created a collection of tweets over time by repeated querying of the
Twitter streaming API.

In line with recent calls for students to work with real data,\footnote{Nolan
  and Temple Lang, ``Computing in the Statistics Curricula.''} our
collection of tweets has served as a valuable resource in our mentoring
of undergraduate data science research. Working with real data allows
students to develop proficiency not only in statistical analysis, but
also in related data science skills, including data transfer from online
sources, data storage, using data from multiple file formats, and
communicating findings and their limitations. Collaboratively asking and
addressing novel questions with our collection of tweets gave mentored
students opportunities to develop competency in all of these areas.

While our tweet collection enables us to address many possible research
questions, the dynamic content of tweets over time particularly piqued
our interest. Together, students and mentors hypothesized that
high-profile social media events would generate a high volume of tweets,
and that we would detect social media events through changes in tweet
topic content over time. We present below 1) an approach for collecting
tweets in real time and 2) methods for detecting social media events via
latent Dirichlet allocation modeling of tweets and 3) suggestions for
using tweets in research mentoring of undergraduate students.

\hypertarget{preliminary-research-mentoring-considerations}{%
\subsection{Preliminary research mentoring
considerations}\label{preliminary-research-mentoring-considerations}}

We collaboratively developed research goals with students through a
series of discussions throughout the academic year. As trainees begin
their senior research projects, we suggest that mentors discuss with
them:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Student experience with data analysis software
\item
  Student research interests and goals
\item
  Best practices for reproducibility
\end{enumerate}

\hypertarget{student-experience-with-data-analysis-software}{%
\subsubsection{Student experience with data analysis
software}\label{student-experience-with-data-analysis-software}}

Student experience with data analysis software varies. In our statistics
department, most students learn elementary R computing skills through
class assignments. Some students, by concentrating in computer science,
learn other data analysis software packages, such as Python. Those who
do undergraduate statistics research often learn advanced topics in R
computing, such as R package assembly, documentation, and testing. Many
develop expertise in linux computing and cluster computing, too.

\hypertarget{student-research-interests-and-goals}{%
\subsubsection{Student research interests and
goals}\label{student-research-interests-and-goals}}

In our experience, student interests vary, and their initial ability to
articulate research goals may be limited. An initial brainstorming
session may clarify their interests and encourage them to think
critically about goals under the time constraints of their academic
schedules. Additionally, we anticipate that sharing completed student
project reports will guide student thinking about the scope of possible
projects.

We mentored one student whose interest in financial time series guided
her project. A second student formulated a project around event
detection from tweet time series.

\hypertarget{best-practices-for-reproducibility}{%
\subsubsection{Best practices for
reproducibility}\label{best-practices-for-reproducibility}}

\hypertarget{backward-design}{%
\subsubsection{Backward design}\label{backward-design}}

We sought to give our trainees opportunities to formulate and address a
data science research question. We consulted the GAISE report for
additional guidance. The GAISE report executive summary highlights six
goals for statistics training:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Teach statistical thinking
\item
  Focus on conceptual understanding
\item
  Integrate real data with a context and purpose
\item
  Foster active learning
\item
  Use technology to explore concepts and analyze data
\item
  Use assessments to improve and evaluate student learning
\end{enumerate}

\hypertarget{learning-outcomes}{%
\subsubsection{Learning outcomes}\label{learning-outcomes}}

\hypertarget{time-period}{%
\subsubsection{Time period}\label{time-period}}

Our two students conducted their research projects during the 2015-2016
academic year. We recommend a full academic year for projects of this
magnitude, although a one-semester project is possible. Our students
presented their findings at the statistics department's undergraduate
poster session near the end of the 2015-2016 academic year.

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

\hypertarget{collecting-tweets-over-time}{%
\subsubsection{Collecting tweets over
time}\label{collecting-tweets-over-time}}

We include here instructions for creating a tweet collection. First, we
created a new account on Twitter. With these user credentials, we used
the R package \texttt{rtweet} to query the API. Because we work with
linux operating systems, we used the \texttt{crontab} software to
repeatedly execute R code to submit API queries. Each query lasted five
minutes. We timed the API queries so that there was no time lag between
queries. We stored tweets resulting from API queries in their native
JSON format.

The R package \texttt{rtweet} provides functions that parse tweet JSON
to R data frames. We then conducted all further analyses in R.

Setting up the query task with \texttt{crontab} is straightforward. On
our computer, with Ubuntu 20.04 linux operating system, we opened a
terminal and typed \texttt{crontab\ -e}. This opened a text file
containing user-specified tasks. We added the following line to the
bottom of the file:

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{*/5}\NormalTok{ * * * * R -e }\StringTok{'rtweet::stream_tweets(timeout = (60 * 5), }
\StringTok{parse = FALSE, file_name = paste0("~/work/mentoring/mentoring-framework/data/",}
\StringTok{lubridate::now(), "-tweets"))'}
\end{Highlighting}
\end{Shaded}

Users may need to slightly amend the above line to conform to
requirements of their operating system's \texttt{crontab}.

\hypertarget{querying-twitter-api-to-get-complete-tweets}{%
\subsubsection{Querying Twitter API to get complete
tweets}\label{querying-twitter-api-to-get-complete-tweets}}

Twitter API use agreements forbid users from sharing complete API query
results. However, Twitter permits users to share tweet identification
numbers. A user can then query a Twitter API to obtain complete tweet
data. In our experience, this process is incomplete; that is, many
tweets submitted to the Twitter API return no data. Additionally, on
repeated querying of the API, different sets of tweets return data. This
complicates our goal of making all analyses computationally
reproducible.

From our collection of tweets, we chose to analyze those sent on five
consecutive days from May 10, 2020 to May 14, 2020. To minimize the
computing requirements, we limited our study to tweets sent during a
five-minute period (12:00pm to 12:05pm Eastern time) every day. We then
submitted API queries to Twitter to get the full content of tweets,
including the tweet text. In supplementary files, we provide the R code
that we used to query the Twitter API to obtain full tweet content.

\hypertarget{tweet-structure}{%
\subsubsection{Tweet structure}\label{tweet-structure}}

Tweets are available from the Twitter API as Javascript Object Notation
(JSON) objects. Every tweet consists of multiple key-value pairs. The
number of fields per tweet depends on user settings, retweet status, and
other factors.\footnote{``Introduction to Tweet Json.''} The 31 tweet
key-value pairs belong to 12 distinct classes (Appendix 1). The classes
are either vectors - numeric, logical, or character - or arrays
assembled from the vector classes.

Below is an example of Tweet JSON. Every tweet features the keys
``created\_at'' (the time stamp), ``id\_str'' (a unique tweet
identifier), and ``text''. We use these three keys in our analyses.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{\{}
  \StringTok{"created_at"}\NormalTok{: }\StringTok{"Thu Apr 06 15:24:15 +0000 2017"}\NormalTok{,}
  \StringTok{"id_str"}\NormalTok{: }\StringTok{"850006245121695744"}\NormalTok{,}
  \StringTok{"text"}\NormalTok{: }\StringTok{"1\textbackslash{}/ Today we\textbackslash{}u2019re sharing our vision for the future of the Twitter API platform!"}\NormalTok{,}
  \StringTok{"user"}\NormalTok{: }\KeywordTok{\{}
    \StringTok{"id"}\NormalTok{: }\ExtensionTok{2244994945}\NormalTok{,}
    \StringTok{"name"}\NormalTok{: }\StringTok{"Twitter Dev"}\NormalTok{,}
    \StringTok{"screen_name"}\NormalTok{: }\StringTok{"TwitterDev"}\NormalTok{,}
    \StringTok{"location"}\NormalTok{: }\StringTok{"Internet"}\NormalTok{,}
    \StringTok{"url"}\NormalTok{: }\StringTok{"https:\textbackslash{}/\textbackslash{}/dev.twitter.com\textbackslash{}/"}\NormalTok{,}
    \StringTok{"description"}\NormalTok{: }\StringTok{"Your official source for Twitter Platform news, updates & events. }
\StringTok{    Need technical help? Visit https:\textbackslash{}/\textbackslash{}/twittercommunity.com\textbackslash{}/ \textbackslash{}u2328\textbackslash{}ufe0f }
\StringTok{    #TapIntoTwitter"}
  \KeywordTok{\}}\NormalTok{,}
  \StringTok{"place"}\NormalTok{: }\KeywordTok{\{}   
  \KeywordTok{\}}\NormalTok{,}
  \StringTok{"entities"}\NormalTok{: }\KeywordTok{\{}
    \StringTok{"hashtags"}\NormalTok{:}\BuiltInTok{ [}      
\NormalTok{    ],}
    \StringTok{"urls"}\NormalTok{: [}
\NormalTok{      \{}
        \StringTok{"url"}\NormalTok{: }\StringTok{"https:\textbackslash{}/\textbackslash{}/t.co\textbackslash{}/XweGngmxlP"}\NormalTok{,}
        \StringTok{"unwound"}\NormalTok{: \{}
          \StringTok{"url"}\NormalTok{: }\StringTok{"https:\textbackslash{}/\textbackslash{}/cards.twitter.com\textbackslash{}/cards\textbackslash{}/18ce53wgo4h\textbackslash{}/3xo1c"}\NormalTok{,}
          \StringTok{"title"}\NormalTok{: }\StringTok{"Building the Future of the Twitter API Platform"}
\NormalTok{        \}}
\NormalTok{      \}}
\NormalTok{    ],}
    \StringTok{"user_mentions"}\NormalTok{: [     }
\NormalTok{    ]}
\NormalTok{  \}}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\hypertarget{parsing-text-of-tweets}{%
\subsubsection{Parsing text of tweets}\label{parsing-text-of-tweets}}

We used functions from the \texttt{rtweet} R package to parse tweet JSON
into a data frame. We divided JSON arrays into their component vectors
and added them to the data frame.

We then divided tweet text into words with functions from the
\texttt{tidytext} R package. We discarded commonly used ``stop words''
and emojis.

Latent Dirichlet allocation models require that the corpus be inputted
as a document by term matrix. Each row corresponds to a single document
(a single tweet), and each column is a single term (or word). Each cell
contains a count (the number of occurrences of a term in the specified
document). We created a document by term matrix with the R function
\texttt{cast\_dtm} from the \texttt{tidytext} package.

\hypertarget{latent-dirichlet-allocation}{%
\subsubsection{Latent Dirichlet
allocation}\label{latent-dirichlet-allocation}}

Latent Dirichlet allocation is a statistical method for inferring latent
(unobservable) topics (or themes) from a corpus (or collection) of
documents. We pretend that there's an imaginary process for creating
documents in the corpus. For each document, we choose a discrete
distribution over topics. For example, some Mother's Day tweets wish
mothers a happy celebration. This may constitute one topic in the
corpus. Having chosen a distribution over topics, we then select
document words by first drawing a topic from the distribution over
topics, then drawing a word from the chosen topic. Thus, a topic is
technically defined as a distribution over words in a fixed vocabulary
(or collection of words).

The literature on latent Dirichlet allocation and related methods is
vast, and we won't attempt to review it here.

\hypertarget{study-design}{%
\subsubsection{Study design}\label{study-design}}

We sought to validate our hypothesis that we could detect a major social
media event by examining tweet topic content at distinct time periods.
As a proof of principle of our event detection strategy, we chose to
analyze tweets during and after Mother's Day 2020. We fitted latent
Dirichlet allocation models for each of four distinct five-minute
periods. The first period began at noon Eastern time on Mother's Day
2020. Subsequent time periods started 24, 48, and 72 hours later. We
defined each time period to be a single collection, or corpus, of
tweets. We then fitted latent Dirichlet allocation models to each
corpus.

We used several criteria to evaluate latent Dirichlet allocation model
fits, with emphasis on choosing a reasonable number of topics per
corpus. Our strategy involved both visualization and more quantitative
approaches to model evaluation. For every model, we created one word
cloud per topic.

We then inspected topic contents at each of the four time points.

\hypertarget{results}{%
\subsubsection{Results}\label{results}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{knitr}\OperatorTok{::}\KeywordTok{include_graphics}\NormalTok{(}\StringTok{"../results/beta-2020-05-10.png"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\includegraphics[width=29.17in]{../results/beta-2020-05-10} \caption{Top terms for LDA model from May 10, 2020 (Mother's Day)}\label{fig:unnamed-chunk-5}
\end{figure}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{knitr}\OperatorTok{::}\KeywordTok{include_graphics}\NormalTok{(}\StringTok{"../results/beta-2020-05-11.png"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}
\includegraphics[width=29.17in]{../results/beta-2020-05-11} \caption{Top terms for LDA model from May 11, 2020}\label{fig:unnamed-chunk-6}
\end{figure}

\hypertarget{references}{%
\subsection{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-tweet_json}{}%
``Introduction to Tweet Json.''
\url{https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/intro-to-tweet-json},
2020\\
Accessed: 2020-05-18.

\leavevmode\hypertarget{ref-lin2011joint}{}%
Lin, Cindy Xide, Qiaozhu Mei, Jiawei Han, Yunliang Jiang, and Marina
Danilevsky. ``The Joint Inference of Topic Diffusion and Evolution in
Social Communities.'' In \emph{2011 Ieee 11th International Conference
on Data Mining}, 378--87. IEEE, 2011.

\leavevmode\hypertarget{ref-nolan2010computing}{}%
Nolan, Deborah, and Duncan Temple Lang. ``Computing in the Statistics
Curricula.'' \emph{The American Statistician} 64, no. 2 (2010): 97--107.

\leavevmode\hypertarget{ref-pelled2018little}{}%
Pelled, Ayellet, Josephine Lukito, Fred Boehm, JungHwan Yang, and Dhavan
Shah. ```Little Marco,'`Lyin'Ted,'`Crooked Hillary,' and the `Biased'
Media: How Trump Used Twitter to Attack and Organize.'' In \emph{Digital
Discussions}, 176--96. Routledge, 2018.

\leavevmode\hypertarget{ref-drob}{}%
Robinson, David. ``Text Analysis of Trump's Tweets Confirms He Writes
Only the (Angrier) Android Half.''
\url{http://varianceexplained.org/r/trump-tweets/}, 2016\\
Accessed: 2019-11-26.

\leavevmode\hypertarget{ref-tweet_stream}{}%
``Sampled Stream.''
\url{https://developer.twitter.com/en/docs/labs/sampled-stream/overview},
2019\\
Accessed: 2019-11-27.

\leavevmode\hypertarget{ref-wells2016trump}{}%
Wells, Chris, Dhavan V Shah, Jon C Pevehouse, JungHwan Yang, Ayellet
Pelled, Frederick Boehm, Josephine Lukito, Shreenita Ghosh, and Jessica
L Schmidt. ``How Trump Drove Coverage to the Nomination: Hybrid Media
Campaigning.'' \emph{Political Communication} 33, no. 4 (2016): 669--76.

\newpage

\hypertarget{appendix-1-tweet-data-dictionary}{%
\subsection{Appendix 1: Tweet data
dictionary}\label{appendix-1-tweet-data-dictionary}}

Twitter shares a data dictionary for tweets
(\url{https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object},
(Accessed: May 23, 2020)). We have saved it as a supplementary file,
``tweets-data-dictionary.csv''.

\end{document}

### 2020-05-26 Rmd/jse-special-issue-2020.Rmd
updated ms
removed material on best practices for reproducibility

@@
 1. Student research interests and goals
-1. Best practices for reproducibility
 
@@
 We mentored one student whose interest in financial time series guided her project. A second student formulated a project around event detection from tweet time series.
-
-### Best practices for reproducibility
 
@@
 
+Our analyses use three fields from each tweet: date ("created_at"), tweet identifier 
+("id_str"), and tweet text ("text"). The "created_at" field is a character string containing 
+the date and time of the tweet. Every tweet has a unique identifier, the "id_str" value. The 
+"text" field contains the unicode representation of the message. For our topic modeling, a single day defines a single "corpus", and a single tweet 
+corresponds to a single "document".
+
 ### Parsing text of tweets
@@
 We used functions from the `rtweet` R package to parse tweet JSON into a data frame. 
-We divided JSON arrays into their component vectors and added them
-to the data frame.
-
-We then divided tweet text into words with functions from the `tidytext` R package. We discarded commonly used "stop words" 
-and emojis.
+We then divided tweet text into words with functions from the `tidytext` R package. We 
+discarded commonly used "stop words" and emojis.
 
 Latent Dirichlet allocation models require that the corpus be inputted as a 
-document by term matrix. Each row corresponds to a single document (a single 
+document by term matrix. In a document by term matrix, each row corresponds to a single document (a single 
 tweet), and each column is a single term (or word). Each cell contains a 
@@
 
+\newpage
+
 ## Appendix 2: LDA results for May 12, 13, and 14
@@
 ```
+
+\newpage
+
+## Appendix 3: R analysis code

### 2020-05-26 Rmd/jse-special-issue-2020.Rmd
deleted backward design stuff

@@
 We mentored one student whose interest in financial time series guided her project. A second student formulated a project around event detection from tweet time series.
-
-### Backward design
-
-We sought to give our trainees opportunities to formulate and address a data science research question. We consulted the GAISE report for additional guidance. 
-The GAISE report executive summary highlights six goals for statistics training:
-
-1. Teach statistical thinking
-1. Focus on conceptual understanding
-1. Integrate real data with a context and purpose
-1. Foster active learning
-1. Use technology to explore concepts and analyze data
-1. Use assessments to improve and evaluate student learning
-
-### Learning outcomes
 
@@
 
-### Parsing text of tweets
+### Parsing tweet text
 
-We used functions from the `rtweet` R package to parse tweet JSON into a data frame. 
+We used functions from the `rtweet` R package to parse tweet JSON into a data frame [@]. 
 We then divided tweet text into words with functions from the `tidytext` R package. We 

### 2020-05-27 Rmd/jse-special-issue-2020.Rmd
added tuning plots

@@
 
-We used functions from the `rtweet` R package to parse tweet JSON into a data frame [@]. 
-We then divided tweet text into words with functions from the `tidytext` R package. We 
+We used functions from the `rtweet` R package to parse tweet JSON into a data frame [@rtweet-package]. 
+We then divided tweet text into words with functions from the `tidytext` R package [@tidytext]. We 
 discarded commonly used "stop words" and emojis.

### 2020-05-29 Rmd/jse-special-issue-2020.Rmd
switched to using memorial day tweets

@@
 
-In line with recent calls for students to work with real data [@nolan2010computing], our 
+In line with recent calls for students to work with real data [@nolan2010computing;@carver2016guidelines], our 
 collection of tweets has served as a valuable resource in our mentoring 
@@
 transfer from online sources, data storage, using data 
-from multiple file formats, and communicating findings and their limitations. 
+from multiple file formats, and communicating findings. 
 Collaboratively asking and addressing novel questions with 
@@
 
-In our experience, student interests vary, and their initial ability to articulate research goals may be limited.
-An initial brainstorming session may clarify their interests and encourage them to think critically about goals under the time constraints of their academic schedules.
-Additionally, we anticipate that sharing completed student project reports will guide student thinking about the scope of possible projects.
+In our experience, student interests vary, and students' ability to articulate research goals 
+may be limited. An initial brainstorming session may clarify their interests and encourage 
+them to think critically about goals under the time constraints of their academic schedules. 
+Additionally, we anticipate that sharing completed student project reports will guide student 
+thinking about the scope of possible projects.
 
-We mentored one student whose interest in financial time series guided her project. A second student formulated a project around event detection from tweet time series.
+We mentored one student whose interest in financial time series and tweet sentiment analysis 
+guided her project. A second student formulated a project around event detection from tweet 
+time series.
 
@@
 there's an imaginary process for creating documents in the corpus. For each document,
-we choose a discrete distribution over topics. For example, some Mother's Day tweets wish 
-mothers a happy celebration. This may constitute one topic in the corpus. 
+we choose a discrete distribution over topics. For example, some tweets from Memorial Day may refer to the holiday. This may constitute one topic in the corpus. 
 Having chosen a distribution over topics, we then select document words by first 
@@
 chosen topic. 
-Thus, a topic is technically defined as a distribution over words in a fixed vocabulary (or collection of words).
-
-The literature on latent Dirichlet allocation and related methods is vast, 
-and we won't attempt to review it here.
+Thus, a topic is technically defined as a distribution over words in a fixed vocabulary (or collection of words) [@blei2003latent].
 
@@
 As a proof of principle of our event detection strategy, we chose to 
-analyze tweets during and after Mother's Day 2020. We fitted latent Dirichlet allocation models for each of four 
-distinct five-minute periods. The first period began at noon Eastern time on Mother's Day 2020. Subsequent time periods started 
-24, 48, and 72 hours later. 
+analyze tweets before, during, and after Memorial Day 2020 (May 25, 2020). We fitted latent Dirichlet allocation models for each of three
+distinct five-minute periods. The first period began at noon Eastern time on May 24, 2020. Subsequent time periods started 
+24 and 48 hours later. 
 We defined each time period to be a single collection, or corpus, of tweets. We then fitted latent 
@@
 
+One challenge in latent Dirichlet allocation modeling is the choice of the 
+number of topics per corpus. To address this, we fitted each corpus to models
+
 We used several criteria to evaluate latent Dirichlet allocation model fits, with 
-emphasis on choosing a reasonable number of topics per corpus.
+emphasis on choosing a small number of topics per corpus.
 Our strategy involved both visualization and more quantitative approaches to model 
@@
 
-## Appendix 2: LDA results for May 12, 13, and 14
-
-```{r, fig.cap = "Top terms for LDA model from May 12, 2020"}
-knitr::include_graphics("../results/beta-2020-05-12.png")
-```
-
-```{r, fig.cap = "Top terms for LDA model from May 13, 2020"}
-knitr::include_graphics("../results/beta-2020-05-13.png")
-```
-
-```{r, fig.cap = "Top terms for LDA model from May 14, 2020"}
-knitr::include_graphics("../results/beta-2020-05-14.png")
-```
-
 \newpage
 
-## Appendix 3: R analysis code
+## Appendix 2: R analysis code

### 2020-05-29 Rmd/tweets-one.Rmd
switched to using memorial day tweets

--- 2020-05-21 Rmd/tweets-one.Rmd
@@
 ```{r, recover-{{date}}}
-fn <- paste0("../data/", {{date}}, "-analyzed-tweet-ids.csv")
-tr_fn <- paste0("../data/", {{date}}, "-rehydrated-tweets.csv")
-if (!file.exists(fn)){
-
+fn <- paste0("../data/", "{{date}}", "-analyzed-tweet-ids.csv")
+#fn <- paste0("../data/", "2020-05-10", "-analyzed-tweet-ids.csv")
+tr_fn <- paste0("../data/", "{{date}}", "-rehydrated-tweets.rds")
+#tr_fn <- paste0("../data/", "2020-05-10", "-rehydrated-tweets.rds")
+sleep <- FALSE
+if (!file.exists(tr_fn)){
+  sleep <- TRUE
   tweets <- recover_stream(path = "{{tweet_json_file}}", verbose = TRUE) 
+  #tweets <- recover_stream(path = "../data/2020-05-26 12:00:02-tweets.json", verbose = TRUE) 
+  id_fn <- paste0("../data/tweets-status-ids-", "{{date}}", ".csv")
+  #id_fn <- paste0("../data/tweets-status-ids-", "2020-05-26", ".csv")
   tweets %>%
     dplyr::select(status_id) %>%
-    readr::write_csv(path = paste0("../data/tweets-status-ids-", "{{date}}", ".csv"), 
+    readr::write_csv(path = id_fn, 
                      col_names = FALSE)
@@
 
-```{r, lda-{{date}}}
-tweets_lda <- topicmodels::LDA(tweets_dtm, k = 8)
+## Comparing models with distinct number of topics
+
+```{r, tuning-{{date}}}
+(result <- ldatuning::FindTopicsNumber(
+  tweets_dtm,
+  topics = seq(from = 5, to = 40, by = 1),
+  metrics = c("CaoJuan2009", "Arun2010", "Deveaud2014"),
+  method = "VEM",
+  mc.cores = 2L,
+  verbose = TRUE
+))
+```
+
+```{r, tuning2-{{date}}}
+ldatuning::FindTopicsNumber_plot(result)
+```
+
+```{r, tuning-fig-{{date}}}
+png(paste0("../results/ldatuning-", "{{date}}", ".png"))
+ldatuning::FindTopicsNumber_plot(result)
+dev.off()
+```
+
+```{r, optimal-k-{{date}}}
+(optimal_k_CaoJuan2009 <- result$topics[which.min(result$CaoJuan2009)])
+(optimal_k_Arun2010 <- result$topics[which.min(result$Arun2010)])
+(optimal_k_Deveaud2014 <- result$topics[which.max(result$Deveaud2014)])
+```
+
+```{r, lda-CJ09-{{date}}}
+tweets_lda <- topicmodels::LDA(tweets_dtm, k = optimal_k_CaoJuan2009)
 tweets_beta <- tidytext::tidy(tweets_lda, matrix = "beta")
@@
 
-```{r, topterms-{{date}}}
+```{r, topterms-CJ09-{{date}}}
 tweets_top_terms <- tweets_beta %>%
@@
 
-```{r, plot-{{date}}}
+```{r, plot-CJ09-{{date}}}
 tweets_top_terms %>%
@@
 
-```{r, ggsave-{{date}}}
-ggplot2::ggsave(paste0("../results/beta-", {{date}}, ".png"))
+```{r, ggsave-CJ09-{{date}}}
+ggplot2::ggsave(paste0("../results/beta-CJ09", "{{date}}", ".png"))
 ```
 
+### Arun2010
+
+```{r, lda-Arun2010-{{date}}}
+tweets_lda <- topicmodels::LDA(tweets_dtm, k = optimal_k_Arun2010)
+tweets_beta <- tidytext::tidy(tweets_lda, matrix = "beta")
+```
+
+```{r, topterms-Arun2010-{{date}}}
+tweets_top_terms <- tweets_beta %>%
+  dplyr::group_by(topic) %>%
+  dplyr::top_n(10, beta) %>%
+  dplyr::ungroup() %>%
+  dplyr::arrange(topic, -beta)
+```
+
+```{r, plot-Arun2010-{{date}}}
+tweets_top_terms %>%
+  dplyr::mutate(term = tidytext::reorder_within(term, beta, topic)) %>%
+  ggplot2::ggplot(ggplot2::aes(term, beta, fill = factor(topic))) +
+  ggplot2::geom_col(show.legend = FALSE) +
+  ggplot2::facet_wrap(~ topic, scales = "free") +
+  ggplot2::coord_flip() + 
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 270))
+
+```
+
+```{r, ggsave-Arun2010-{{date}}}
+ggplot2::ggsave(paste0("../results/beta-Arun2010", "{{date}}", ".png"))
+```
+
+### Deveaud2014
+
+```{r, lda-Deveaud2014-{{date}}}
+tweets_lda <- topicmodels::LDA(tweets_dtm, k = optimal_k_Deveaud2014)
+tweets_beta <- tidytext::tidy(tweets_lda, matrix = "beta")
+```
+
+```{r, topterms-Deveaud2014-{{date}}}
+tweets_top_terms <- tweets_beta %>%
+  dplyr::group_by(topic) %>%
+  dplyr::top_n(10, beta) %>%
+  dplyr::ungroup() %>%
+  dplyr::arrange(topic, -beta)
+```
+
+```{r, plot-Deveaud2014-{{date}}}
+tweets_top_terms %>%
+  dplyr::mutate(term = tidytext::reorder_within(term, beta, topic)) %>%
+  ggplot2::ggplot(ggplot2::aes(term, beta, fill = factor(topic))) +
+  ggplot2::geom_col(show.legend = FALSE) +
+  ggplot2::facet_wrap(~ topic, scales = "free") +
+  ggplot2::coord_flip() + 
+  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 270))
+
+```
+
+```{r, ggsave-Deveaud2014-{{date}}}
+ggplot2::ggsave(paste0("../results/beta-Deveaud2014", "{{date}}", ".png"))
+```
+
+### Wait
+
 ```{r, sleep-{{date}}}
-Sys.sleep(15 * 60) # sleep for 15 minutes to allow api limits to reset
+if (sleep) Sys.sleep(15 * 60) # sleep for 15 minutes to allow api limits to reset
 ```

### 2020-05-29 Rmd/tweets.Rmd
Removed 'goals' section
switched to using memorial day tweets

@@
 ---
-
-## Goals
-
-1. Write a R script to query twitter streaming api to get 1 percent sample over 5'.
-1. Demonstrate how to set up a crontab to execute script every 5'.
-1. Cautions about twitter databases: deletions and amendments to tweets
 
@@
 ```{r}
-date <- c("2020-05-10", "2020-05-11", "2020-05-12", "2020-05-13", "2020-05-14")
-tweet_json_file <- c("../data/2020-05-10 12:00:01-tweets.json",
-  "../data/2020-05-11 12:00:01-tweets.json", 
-  "../data/2020-05-12 12:00:02-tweets.json",
-  "../data/2020-05-13 12:00:02-tweets.json", 
-  "../data/2020-05-14 12:00:02-tweets.json")
+date <- c("2020-05-24", "2020-05-25", "2020-05-26")
+tweet_json_file <- c("../data/2020-05-24 12:00:02-tweets.json",
+  "../data/2020-05-25 12:00:02-tweets.json", 
+  "../data/2020-05-26 12:00:02-tweets.json"
+  )
 out <- purrr::pmap_dfc(tibble::tibble(date, tweet_json_file),
@@
 
-## May 10, 2020
+## May 24, 2020
 
@@
 
-## May 11, 2020
+## May 25, 2020
 
@@
 
-## May 12, 2020
+## May 26, 2020
 
 `r knitr::knit(text = out[[3]])`
-
-## May 13, 2020
-
-`r knitr::knit(text = out[[4]])`
-
-## May 14, 2020
-
-`r knitr::knit(text = out[[5]])`
 

### 2020-05-29 Rmd/tweets-one.Rmd
removed code for comparing models with dift k values

--- 2020-05-24 Rmd/tweets-one.Rmd
@@
 #tr_fn <- paste0("../data/", "2020-05-10", "-rehydrated-tweets.rds")
+sleep <- FALSE
 if (!file.exists(tr_fn)){
+  sleep <- TRUE
   tweets <- recover_stream(path = "{{tweet_json_file}}", verbose = TRUE) 
-  #tweets <- recover_stream(path = "../data/2020-05-10 12:00:01-tweets.json", verbose = TRUE) 
+  #tweets <- recover_stream(path = "../data/2020-05-26 12:00:02-tweets.json", verbose = TRUE) 
   id_fn <- paste0("../data/tweets-status-ids-", "{{date}}", ".csv")
-  #id_fn <- paste0("../data/tweets-status-ids-", "2020-05-10", ".csv")
+  #id_fn <- paste0("../data/tweets-status-ids-", "2020-05-26", ".csv")
   tweets %>%
@@
 
+### Get beta plots
+
 ```{r, lda-{{date}}}
-tweets_lda <- topicmodels::LDA(tweets_dtm, k = 6)
+tweets_lda <- topicmodels::LDA(tweets_dtm, k = 10)
 tweets_beta <- tidytext::tidy(tweets_lda, matrix = "beta")
@@
 
+### Wait
+
 ```{r, sleep-{{date}}}
-Sys.sleep(15 * 60) # sleep for 15 minutes to allow api limits to reset
+if (sleep) Sys.sleep(15 * 60) # sleep for 15 minutes to allow api limits to reset
 ```

### 2020-05-29 Rmd/abstract.Rmd
updated ms
revised abstract. added section numbers

@@
 
-We describe methods for collecting tweets from Twitter. With these methods, we 
-curate a large collection of tweets.
-We present methods for social media event detection through text analysis of tweets. 
-We describe our use of tweets in mentored undergraduate research. 
-We also offer other directions for other text analysis of tweets.
+We present a freely available, large, rich data set from Twitter and describe how we collected 
+it. We then describe mentored student research that uses tweets to address novel research
+questions. We use backward design principles to develop learning objectives
+for student researchers. We conduct formative and summative assessments of student learning.
+To illustrate the value of our data set, we 
+present details of one student project. We conclude by discussing future research 
+directions.

### 2020-05-29 Rmd/jse-special-issue-2020.Rmd
revised abstract. added section numbers

@@
 ---
-title: "What is happening on Twitter? Tweet analysis and undergraduate research projects"
+title: "What is happening on Twitter? Tweet analysis and student research projects"
 author: "Frederick J. Boehm and Bret M. Hanlon"
@@
 output: 
-  pdf_document:
+  bookdown::pdf_document2:
     latex_engine: xelatex
     toc: FALSE
-    number_sections: FALSE
+    number_sections: TRUE
     includes:
@@
 
-```{r child = "abstract.Rmd", eval = FALSE}
+```{r child = "abstract.Rmd", eval = TRUE}
+knitr::opts_chunk$set(echo = FALSE)
 ```
 
-## Introduction
+# Introduction
 
@@
 
-## Preliminary research mentoring considerations
+# Preliminary research mentoring considerations
 
@@
 
-### Student experience with data analysis software
+## Student experience with data analysis software
 
@@
 
-### Student research interests and goals
+## Student research interests and goals
 
@@
 
-### Time period
+## Time period
 
@@
 
-## Methods
+# Methods
 
-### Collecting tweets over time
+## Collecting tweets over time
 
@@
 
-### Querying Twitter API to get complete tweets
+## Querying Twitter API to get complete tweets
 
@@
 
-From our collection of tweets, we chose to analyze those sent on five consecutive 
-days from May 10, 2020 to May 14, 2020. To minimize the computing requirements, we 
-limited our study to tweets sent during a five-minute period (12:00pm to 12:05pm Eastern time) every day.
+From our collection of tweets, we chose to analyze those sent on three consecutive 
+days from May 24, 2020 to May 26, 2020. We wanted to see if we could use text analysis 
+tools to detect a transient change in topics for Memorial Day (May 25, 2020).
+
+To minimize the computing requirements, we 
+limited our analysis to tweets sent during a five-minute period (12:00pm to 12:05pm Eastern time) every day. However, our methods are appropriate for much larger data sets.
 We then submitted API queries to Twitter
@@
 
-### Tweet structure
+## Tweet structure
 
-```{r, echo = FALSE, message = FALSE}
+```{r, echo = TRUE, message = FALSE}
 data_dictionary <- readr::read_csv("../data/tweets-data-dictionary.csv")
@@
 
-### Parsing tweet text
+## Parsing tweet text
 
@@
 
-### Latent Dirichlet allocation
+## Latent Dirichlet allocation
 
 Latent Dirichlet allocation is a statistical method for inferring latent (unobservable) topics (or themes)
-from a corpus (or collection) of documents. We pretend that
-there's an imaginary process for creating documents in the corpus. For each document,
-we choose a discrete distribution over topics. For example, some tweets from Memorial Day may refer to the holiday. This may constitute one topic in the corpus. 
+from a large corpus (or collection) of documents [@blei2003latent].
+We pretend that there's an imaginary process for creating documents in the corpus. For each document,
+we choose a discrete distribution over topics. For example, some tweets from Memorial Day may 
+refer to the holiday. This may constitute one topic in the corpus. 
 Having chosen a distribution over topics, we then select document words by first 
@@
 
-### Study design
+## Study design
 
-We sought to validate our hypothesis that we could detect a major social 
+We sought to validate our hypothesis that we could detect a social 
 media event by examining tweet topic content at distinct time periods. 
 As a proof of principle of our event detection strategy, we chose to 
-analyze tweets before, during, and after Memorial Day 2020 (May 25, 2020). We fitted latent Dirichlet allocation models for each of three
+analyze tweets before, during, and after Memorial Day (May 25, 2020). We fitted latent Dirichlet allocation models for each of three
 distinct five-minute periods. The first period began at noon Eastern time on May 24, 2020. Subsequent time periods started 
@@
 
-One challenge in latent Dirichlet allocation modeling is the choice of the 
-number of topics per corpus. To address this, we fitted each corpus to models
+# Results
 
-We used several criteria to evaluate latent Dirichlet allocation model fits, with 
-emphasis on choosing a small number of topics per corpus.
-Our strategy involved both visualization and more quantitative approaches to model 
-evaluation. For every model, we created one word cloud per topic.
+We identified the top ten terms for each of ten topics in our models. We plotted 
+the within-topic word probabilities as bar graphs.
 
-We then inspected topic contents at each of the four time points.
-
-### Results
-
-```{r, fig.cap = "Top terms for LDA model from May 10, 2020 (Mother's Day)"}
-knitr::include_graphics("../results/beta-2020-05-10.png")
+```{r, fig.cap = "Top terms for LDA model from May 24, 2020"}
+knitr::include_graphics("../results/beta-2020-05-24.png")
 ```
 
-```{r, fig.cap = "Top terms for LDA model from May 11, 2020"}
-knitr::include_graphics("../results/beta-2020-05-11.png")
+```{r, fig.cap = "Top terms for LDA model from May 25, 2020 (Memorial Day)"}
+knitr::include_graphics("../results/beta-2020-05-25.png")
 ```
 
-## Acknowledgements
+```{r, fig.cap = "Top terms for LDA model from May 26, 2020"}
+knitr::include_graphics("../results/beta-2020-05-26.png")
+```
+
+Assigning meaning to topics is an active research area [@chang2009reading]. Since
+our interest is in the transient appearance of a new topic, we don't attempt to
+assign meaning to topics in our models.
+
+We see that topic 7 from May 25 has several words that suggest Memorial Day: 
+memorial, remember, honor, country. A similar topic is not seen on May 24 or May 26.
+
+Some topics persist, possibly with distinct word probabilities, across the three days. For example, we see that President Trump features prominently in all three models.
+
+We also note, on May 26, topic 10, which reflects discussion of the Amy Cooper Central Park incident.
+
+We found evidence that latent Dirichlet allocation summarizes elements of 
+large collections of Twitter conversations.
+
+# Assessment of learning, exploring more advanced topics, and concluding remarks
+
+## Assessment of learning
+
+We examined student learning through both formative and summative assessments.
+We conducted formative assessments through weekly discussions with students.
+In these discussions, we developed action items to advance research progress and overcome challenges.
+
+We summatively assessed student achievement at the end of the academic year. Both
+students wrote a thesis and presented a poster to our statistics department. 
+We asked questions at the poster session to probe student understanding and critically
+evaluated the theses.
+
+In future iterations, we will use a written rubric to grade student theses. We'll
+share the rubric with our students at the start of the academic year.
+
+## Exploring more advanced topics
+
+Twitter data over time offers a wealth of potential research projects. Supplementing 
+tweets with public data from other sources multiplies the possibilities. For example, 
+one of our two students supplemented tweets with daily stock market index prices. 
+She studied sentiment of finance-related tweets and daily stock market index
+closing prices.
+
+Latent Dirichlet allocation modeling and related methods are a major research area in the 
+quantitative social sciences. Advanced students with interest in statistical computing might
+compare inferential methods for topic models. Those with interests in event detection
+and time series analysis could build on the findings of our student by explicitly accounting
+for topic evolution with dynamic topic models [@blei2006dynamic].
+
+## Concluding remarks
+
+Tweet collections over time are a rich, large, authentic data set that offer 
+many opportunities for student research projects. We provided 
+instructions to enable readers to 
+establish their own tweet collections. We also presented details for one 
+mentored research project that made use of our stored tweets.
+
+Tweet analysis gives students practical experience in the data science process of 
+formulating a research question, gathering data to address it, summarizing the 
+data, visualizing results, and communicating findings.
+
+# Acknowledgements
 
@@
 
-## Supplementary Material
-
-## References
+# References
 
@@
 
-## Appendix 1: Tweet data dictionary
+# Appendix 1: Tweet data dictionary
 
@@
 
-## Appendix 2: R analysis code
+# Appendix 2: R analysis code

### 2020-05-30 Rmd/jse-special-issue-2020.Rmd
updated package installations
added backward design stuff

@@
 ```{r child = "abstract.Rmd", eval = TRUE}
+```
+
+```{r setup, echo = FALSE}
 knitr::opts_chunk$set(echo = FALSE)
@@
 using tweets in research mentoring of undergraduate students.
+
+# Backward design and learning objectives
+
+## Backward design
+
+Backward design principles guided our planning and informed the writing of
+learning objectives [@wiggins2005understanding]. Following @wiggins2005understanding, 
+we began by listing 
+what students, at the end of their thesis research, should be able to do, 
+understand, and know. We then classified each of these items into 
+one of three categories: enduring understanding, important to know and do, 
+and worth being familiar with [@wiggins2005understanding] (Table \@ref(tab:circle-table)).
+
+```{r, circle-table, echo = FALSE, message=FALSE}
+library(magrittr)
+readr::read_csv("../data/concepts-backward-design.csv") %>%
+  knitr::kable(caption = "Classifying project skills")
+```
+
+## Learning objectives
+
+We translated into learning objectives our prioritized list of skills that 
+students should be able to do, understand, and know (Table \@ref(tab:circle-table)). We 
+phrased learning objectives in a manner that enabled their subsequent assessment.
+
+1. Write R code to perform text analysis of large volumes of tweets.
+1. Communicate results in a written report and poster presentation.
+1. Translate statistical findings into scientific conclusions.
+1. Develop data science strategies to address a scientific research question.
+
+Each objective is amenable to formative or summative assessment.
 
@@
 
-```{r, echo = TRUE, message = FALSE}
+```{r, echo = FALSE, message = FALSE}
 data_dictionary <- readr::read_csv("../data/tweets-data-dictionary.csv")
@@
 
-\newpage
-
 # Appendix 2: R analysis code

### 2020-05-30 Rmd/jse-special-issue-2020.Rmd
switched to using ASA citation style file

@@
 bibliography: jse.bib
-csl: chicago-annotated-bibliography.csl
+csl: american-statistical-association.csl
 geometry: "left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm"
@@
 
-```{bash, eval = FALSE, fig.cap = "Example Tweet JSON"}
+```{bash, eval = FALSE, fig.cap = "Example Tweet JSON", echo = TRUE}
 {

### 2020-05-31 Rmd/jse-special-issue-2020.Rmd
added paragraph on inference for LDA

@@
 tools in guiding undergraduate research projects.
-We report one instance of a mentored text analysis research project. 
+We report mentored text analysis projects. 
 We share our data and computer code to encourage others to undertake 
@@
 tweets, and that we would detect social media events through changes in tweet 
-topic content over time. We present below 1) an approach for collecting tweets in real 
-time and 2) methods for detecting social media 
-events via latent Dirichlet allocation modeling of tweets and 3) suggestions for 
-using tweets in research mentoring of undergraduate students.
+topic content over time. We discuss in detail below one approach to studying this
+question.
+
+In the sections that follow, we detail our backward design-inspired approach to 
+writing learning objectives, preliminary research mentoring considerations, 
+data science methods for collecting and analyzing tweets, analysis results, 
+and thoughts on assessment and advanced topics.
 
@@
 
-1. Write R code to perform text analysis of large volumes of tweets.
+1. Write R code to perform text analysis of large volumes of tweets [@r].
 1. Communicate results in a written report and poster presentation.
@@
 
-We collaboratively developed research goals with students through a series of discussions throughout the academic year.
+We collaboratively developed research goals with students in a series of discussions during the academic year.
 As trainees begin their senior research projects, we suggest that mentors discuss with them:
@@
 
-```{bash, eval = FALSE}
+```{bash, eval = FALSE, echo = TRUE}
 */5 * * * * R -e 'rtweet::stream_tweets(timeout = (60 * 5), 
@@
 
-Below is an example of Tweet JSON. Every tweet features the keys "created\_at" (the time stamp), 
-"id\_str" (a unique tweet identifier), and "text". We use these three keys in 
-our analyses.
+Below is an example of Tweet JSON.
 
@@
 
-Latent Dirichlet allocation models require that the corpus be inputted as a 
+Latent Dirichlet allocation models require that the corpus be organized as a 
 document by term matrix. In a document by term matrix, each row corresponds to a single document (a single 
@@
 chosen topic. 
-Thus, a topic is technically defined as a distribution over words in a fixed vocabulary (or collection of words) [@blei2003latent].
+The goal for latent Dirichlet allocation is to infer both the distribution over topics and the topics [@blei2003latent]. A topic, in this setting, is a distribution over 
+the vocabulary (the collection of all words in a corpus).
+
+Inference for latent Dirichlet allocation model is performed by either 
+sampling from the posterior distribution or through variational methods. Researchers
+have devised a variety of Gibbs sampling techniques for latent Dirichlet allocation models. 
+Variational methods, while using approximations to the posterior distribution, offer the 
+advantage of computational speed. We used variational methods below.
 
@@
 formulating a research question, gathering data to address it, summarizing the 
-data, visualizing results, and communicating findings.
+data, visualizing results, and communicating findings. By considering first student 
+research interests and integrating them with our senior thesis learning objectives, 
+we successfully guided two undergraduate researchers in data science research with
+tweets.
 

### 2020-06-01 Rmd/abstract.Rmd
fixed abstract

--- 2020-05-20 Rmd/abstract.Rmd
@@
 
-We describe methods for collecting tweets from Twitter. With these methods, we 
-curate a large collection of tweets.
-We present methods for social media event detection through text analysis of tweets. 
-We describe our use of tweets in mentored undergraduate research. 
-We also offer other directions for other text analysis of tweets.
+We present a framework for undergraduate research projects with Twitter data. 
+Drawing on backward design principles, we share our learning objectives and
+summative assessments. We draw on experiences mentoring two students during 
+an academic year. To illustrate the value of Twitter as a data source, 
+we detail methods for collecting and analyzing tweets. We conclude by emphasizing
+how Twitter text analysis projects enable students to formulate 
+original research questions, collect and analyze the needed data, and communicate
+findings and their implications.

### 2020-06-01 Rmd/jse-special-issue-2020.Rmd
fixed capitalization
fixed abstract

@@
 ---
-title: "What is happening on Twitter? Tweet analysis and student research projects"
+title: "What is happening on Twitter? A framework for student research projects with tweets"
 author: "Frederick J. Boehm and Bret M. Hanlon"
@@
 
+Our mentoring in data science aligns with others' calls to reconsider training in statistics
+and data science [@nolan2010computing;@carver2016guidelines]. @hicks2018guide argue for
+incorporating three concepts into data science classes: computing, connecting and creating. 
+They use the terms "connecting" and "creating" to describe the processes of applying 
+quantitative methods to real data and real research questions and of research question 
+formulation, respectively. Our tweet analysis projects offer students opportunities in all
+three skills sets. Our students first formulated research questions, then collected and 
+analyzed data to address the questions. Throughout the projects, students 
+drew heavily on computing, both to acquire data and to analyze it.
+
 Tweet analysis gives students practical experience in the data science process of 
@@
 # Appendix 2: R analysis code
+
+We include our R code in supplementary files.

### 2020-06-01 Rmd/jse-special-issue-2020.Rmd
updated manuscript with summ assessments table.

@@
 
-In future iterations, we will use a written rubric to grade student theses. We'll
+```{r, summative, echo = FALSE, fig.cap = "Rubric for summative assessment of learning objectives."}
+summative <- readr::read_csv("../data/summative-assessments.csv")
+knitr::kable(summative, caption = "Rubric for summative assessment of learning objectives.")
+```
+
+In future iterations, we will use a written rubric to grade student theses (Table \@ref(tab:summative) ). We'll
 share the rubric with our students at the start of the academic year.

### 2020-06-02 Rmd/jse-special-issue-2020.Rmd
updated bib with blei ref on variational inference.

--- 2020-06-01 Rmd/jse-special-issue-2020.Rmd
@@
 
-Our two students conducted their research projects during the 2015-2016 academic year. 
+Our two statistics students conducted their research projects during the 2015-2016 academic year. 
 We recommend a full academic year for projects of this magnitude, although a one-semester 
@@
 undergraduate poster session near the end of the 2015-2016 academic year.
+
+We present below reproducible R code for analyzing data from May 2020. While these
+are not the same data that our students analyzed in 2015, the methods and code 
+are very similar.
 
@@
 to repeatedly execute R code to submit API queries. Each 
-query lasted five minutes. We timed the API queries so 
+query lasted five minutes and produced a text file of approximately 130 MB. We timed the API queries so 
 that there was no time lag between queries. We stored tweets resulting from API queries in their native JSON format.
-
-The R package `rtweet` 
-provides functions that parse tweet JSON to R data frames. We then 
-conducted all further analyses in R.
 
@@
 
+From our collection of tweets, we chose to analyze those sent on three consecutive 
+days from May 24, 2020 to May 26, 2020. We wanted to see if we could use text analysis 
+tools to detect a transient change in topics for Memorial Day (May 25, 2020). To minimize the computing requirements, we 
+limited our analysis to tweets sent during a five-minute period (12:00pm to 12:05pm Eastern time) every day. However, our methods are appropriate for much larger data sets.
+We then submitted API queries to Twitter
+to get the full content of tweets, including the tweet text.
+In supplementary files, we provide the R code that we used to 
+query the Twitter API to obtain full tweet content.
+
 Twitter API use agreements forbid users from sharing complete API query results. 
@@
 return data. This complicates our goal of making all analyses computationally 
-reproducible.
-
-From our collection of tweets, we chose to analyze those sent on three consecutive 
-days from May 24, 2020 to May 26, 2020. We wanted to see if we could use text analysis 
-tools to detect a transient change in topics for Memorial Day (May 25, 2020).
-
-To minimize the computing requirements, we 
-limited our analysis to tweets sent during a five-minute period (12:00pm to 12:05pm Eastern time) every day. However, our methods are appropriate for much larger data sets.
-We then submitted API queries to Twitter
-to get the full content of tweets, including the tweet text.
-In supplementary files, we provide the R code that we used to 
-query the Twitter API to obtain full tweet content.
+reproducible and motivates our decision to share the tweet IDs of those tweets that 
+we actually analyzed.
 
@@
 
-Tweets are available from the Twitter API as Javascript Object Notation (JSON) objects. Every tweet consists 
+Tweets are available from the Twitter API as Javascript Object Notation (JSON) objects [@json]. 
+Every tweet consists 
 of multiple key-value pairs. The number of fields per 
@@
 the date and time of the tweet. Every tweet has a unique identifier, the "id_str" value. The 
-"text" field contains the unicode representation of the message. For our topic modeling, a single day defines a single "corpus", and a single tweet 
-corresponds to a single "document".
+"text" field contains the unicode representation of the message. After creating a text file 
+with tweet JSON, our next step involved reading and parsing tweets with the R 
+packages `rtweet` and `tidytext`.
 
@@
 
-Latent Dirichlet allocation models require that the corpus be organized as a 
+Latent Dirichlet allocation model fitting requires that the corpus be organized as a 
 document by term matrix. In a document by term matrix, each row corresponds to a single document (a single 
@@
 sampling from the posterior distribution or through variational methods. Researchers
-have devised a variety of Gibbs sampling techniques for latent Dirichlet allocation models. 
+have devised a variety of Gibbs sampling techniques for these models. 
 Variational methods, while using approximations to the posterior distribution, offer the 
-advantage of computational speed. We used variational methods below.
+advantage of computational speed [@blei2017variational]. We used variational methods below.
 
@@
 24 and 48 hours later. 
-We defined each time period to be a single collection, or corpus, of tweets. We then fitted latent 
-Dirichlet allocation models to each corpus.
+We defined each time period to be a single collection, or corpus, of tweets.
 
@@
 our interest is in the transient appearance of a new topic, we don't attempt to
-assign meaning to topics in our models.
-
-We see that topic 7 from May 25 has several words that suggest Memorial Day: 
-memorial, remember, honor, country. A similar topic is not seen on May 24 or May 26.
-
-Some topics persist, possibly with distinct word probabilities, across the three days. For example, we see that President Trump features prominently in all three models.
-
-We also note, on May 26, topic 10, which reflects discussion of the Amy Cooper Central Park incident.
-
-We found evidence that latent Dirichlet allocation summarizes elements of 
-large collections of Twitter conversations.
+assign meaning to every topic in our models. We see that topic 7 from May 25 has several 
+words that suggest Memorial Day: memorial, remember, honor, country. A similar topic is not 
+seen on May 24 or May 26. Some topics persist, possibly with distinct word probabilities, 
+across the three days. For example, we see that President Trump features prominently in all 
+three models. We also note, on May 26, topic 10, which reflects discussion of the Amy Cooper Central Park incident.
 
@@
 
-In future iterations, we will use a written rubric to grade student theses. We'll
+```{r, summative, echo = FALSE, message = FALSE}
+summative <- readr::read_csv("../data/summative-assessments.csv")
+knitr::kable(summative, caption = "Rubric for summative assessment of learning objectives.") %>%
+  kableExtra::column_spec(column = 1:5, width = "10em") %>%
+  kableExtra::kable_styling(font_size = 6)
+
+```
+
+In future iterations, we will use a written rubric to grade student theses (Table \@ref(tab:summative)). We'll
 share the rubric with our students at the start of the academic year.

### 2020-06-05 Rmd/jse-special-issue-2020.Rmd
updated abstract

@@
 tools in guiding undergraduate research projects.
-We report mentored text analysis projects. 
-We share our data and computer code to encourage others to undertake 
-tweet text analysis research. We also describe our methods for creating a 
+We guided two statistics students in their senior research projects. 
+Both students used tweets to address novel research questions. We 
+share products of their research in supplementary files. Because 
+their data are no longer available, we present as a case study one 
+analysis with tweets from May 2020. We share our data and computer 
+code to encourage others to undertake 
+tweet text analysis research. We also describe methods for creating a 
 collection of tweets.
@@
 available through website application product interfaces (APIs). 
-By way of a streaming API, Twitter shares a sample of approximately one percent of all 
-tweets during an API query time period [@tweet_stream]. Any Twitter user can freely 
+By way of a streaming API, Twitter shares a sample of approximately 
+one percent of all 
+tweets during an API query time period [@tweet_stream]. Any Twitter 
+user can freely 
 access this one percent sample, whereas access to a larger selection is available to researchers for a fee.
 
-Using large collections of tweets, scholars have studied diverse research questions, including
+Studies of Twitter conversations have yielded valuable insights into 
+modern culture. Using large collections of tweets, scholars have 
+investigated diverse research questions, including
 the inference of relationships and social networks among Twitter users [@lin2011joint]; 
@@
 While our tweet collection enables us to address many possible research questions, the 
-dynamic content of tweets over time particularly piqued our interest. Together, students and 
-mentors hypothesized that high-profile social media events would generate a high volume of 
+dynamic content of tweets over time particularly piqued our interest. 
+We hypothesized that high-profile social media events would generate a high volume of 
 tweets, and that we would detect social media events through changes in tweet 
@@
 data science methods for collecting and analyzing tweets, analysis results, 
-and thoughts on assessment and advanced topics.
+and ideas on assessment and advanced topics.
 
@@
 
+We note that nearly all of the skills in Table \@ref(tab:circle-table) are 
+transferable in that they apply not merely to the thesis projects, but also to
+data science research in general.
+
 ## Learning objectives
@@
 students should be able to do, understand, and know (Table \@ref(tab:circle-table)). We 
-phrased learning objectives in a manner that enabled their subsequent assessment.
+phrased learning objectives in a manner that enabled their subsequent assessment 
+(Table \@ref(tab:summative)) with formative and summative strategies.
 
@@
 
-Each objective is amenable to formative or summative assessment.
-
 # Preliminary research mentoring considerations
 
-We collaboratively developed research goals with students in a series of discussions during the academic year.
-As trainees begin their senior research projects, we suggest that mentors discuss with them:
+We developed research goals with students in a series of discussions. As trainees began their senior research projects, we spoke in detail about:
 
@@
 expertise in linux computing and cluster computing, too.
+
+Our data science projects required that students possess elementary R computing
+skills. We guided students towards supplementary R computing resources, including 
+the online books "R for data science" [@wickham2016r] and "Advanced R" [@wickham2019advanced].
 

### 2020-06-05 Rmd/jse-special-issue-2020.Rmd
updated per comments of June 4 ms

@@
 Additionally, we anticipate that sharing completed student project reports will guide student 
-thinking about the scope of possible projects.
-
-We mentored one student whose interest in financial time series and tweet sentiment analysis 
-guided her project. A second student formulated a project around event detection from tweet 
-time series.
+thinking about the scope of possible projects (Supplemental files).
 
@@
 
-From our collection of tweets, we chose to analyze those sent on three consecutive 
-days from May 24, 2020 to May 26, 2020. We wanted to see if we could use text analysis 
-tools to detect a transient change in topics for Memorial Day (May 25, 2020). To minimize the computing requirements, we 
-limited our analysis to tweets sent during a five-minute period (12:00pm to 12:05pm Eastern time) every day. However, our methods are appropriate for much larger data sets.
-We then submitted API queries to Twitter
-to get the full content of tweets, including the tweet text.
-In supplementary files, we provide the R code that we used to 
-query the Twitter API to obtain full tweet content.
+We analyzed tweets sent on three consecutive 
+days from May 24, 2020 to May 26, 2020. We wanted to see if we could detect a transient change in topics for Memorial Day (May 25, 2020). To minimize the computing requirements, we 
+limited our analysis to tweets sent during a five-minute period (12:00pm to 12:05pm Eastern time) every day. However, our methods are also appropriate for much larger data sets.
 
 Twitter API use agreements forbid users from sharing complete API query results. 
-However, Twitter permits users to share tweet identification numbers. A user can 
-then query a Twitter API to obtain complete tweet data. In our experience, this 
+However, Twitter permits users to share tweet identification numbers. With a tweet 
+identification number, a user may query a Twitter API to obtain complete tweet data. In our 
+experience, this 
 process is incomplete; that is, many tweets submitted to the Twitter API return 
@@
 reproducible and motivates our decision to share the tweet IDs of those tweets that 
-we actually analyzed.
+we actually analyzed (Supplemental files).
 
@@
 
+The next task is to wrangle the tweet JSON into a data structure suitable for latent Dirichlet allocation modeling. 
 We used functions from the `rtweet` R package to parse tweet JSON into a data frame [@rtweet-package]. 
@@
 
-Inference for latent Dirichlet allocation model is performed by either 
+Inference for latent Dirichlet allocation models is performed by either 
 sampling from the posterior distribution or through variational methods. Researchers
-have devised a variety of Gibbs sampling techniques for these models. 
+have devised a variety of Gibbs sampling techniques for these models [@porteous2008fast]. 
 Variational methods, while using approximations to the posterior distribution, offer the 
@@
 media event by examining tweet topic content at distinct time periods. 
-As a proof of principle of our event detection strategy, we chose to 
-analyze tweets before, during, and after Memorial Day (May 25, 2020). We fitted latent Dirichlet allocation models for each of three
+As a proof of principle of our event detection strategy, we analyzed tweets before, during, and 
+after Memorial Day (May 25, 2020). We fitted latent Dirichlet allocation models for each of three
 distinct five-minute periods. The first period began at noon Eastern time on May 24, 2020. Subsequent time periods started 
@@
 
-We examined student learning through both formative and summative assessments.
+We examined student learning with both formative and summative assessments.
 We conducted formative assessments through weekly discussions with students.
-In these discussions, we developed action items to advance research progress and overcome challenges.
-
-We summatively assessed student achievement at the end of the academic year. Both
+In these discussions, we developed action items to advance research progress and overcome challenges. We summatively assessed student achievement at the end of the academic year. Both
 students wrote a thesis and presented a poster to our statistics department. 
@@
 
-In future iterations, we will use a written rubric to grade student theses (Table \@ref(tab:summative)). We'll
+With future students, we will use a written rubric to evaluate theses (Table \@ref(tab:summative)). We'll
 share the rubric with our students at the start of the academic year.
@@
 instructions to enable readers to 
-establish their own tweet collections. We also presented details for one 
-mentored research project that made use of our stored tweets.
+establish their own tweet collections. We also presented details for one analysis strategy.
 
-Our mentoring in data science aligns with others' calls to reconsider training in statistics
+Our mentoring in data science aligns with others' calls to reconsider the role of computing in statistics
 and data science [@nolan2010computing;@carver2016guidelines]. @hicks2018guide argue for

### 2020-06-08 Rmd/jse-special-issue-2020.Rmd
added supp files

@@
 topic content over time. We discuss in detail below one approach to studying this
-question.
-
-In the sections that follow, we detail our backward design-inspired approach to 
+question. In the sections that follow, we detail our backward design-inspired approach to 
 writing learning objectives, preliminary research mentoring considerations, 
@@
 
-# Methods
+# Case study methods
+
+To illustrate the value of Twitter data, we present below a 
+reproducible case study. In it, we aim to detect a social media 
+event by examining topic content over time. We use latent Dirichlet 
+allocation models to infer topics on three consecutive days centered 
+on Memorial Day 2020. We chose to use this example case study, 
+instead of the student projects, because of limited data 
+availability for the student projects. Despite this, the case study 
+illustrates the strategy and methods for one of the student 
+projects.
+
+## Case study design
+
+We sought to validate our hypothesis that we could detect a social 
+media event by examining tweet topic content at distinct time periods. 
+As a proof of principle of our event detection strategy, 
+we analyzed tweets before, during, and 
+after Memorial Day (May 25, 2020). 
+We fitted latent Dirichlet allocation models for each of three
+distinct five-minute periods. The first period began at noon Eastern time 
+on May 24, 2020. Subsequent time periods started 
+24 and 48 hours later. 
+We defined each time period to be a single collection, or corpus, of 
+tweets.
 
@@
 
-We analyzed tweets sent on three consecutive 
-days from May 24, 2020 to May 26, 2020. We wanted to see if we could detect a transient change in topics for Memorial Day (May 25, 2020). To minimize the computing requirements, we 
-limited our analysis to tweets sent during a five-minute period (12:00pm to 12:05pm Eastern time) every day. However, our methods are also appropriate for much larger data sets.
-
 Twitter API use agreements forbid users from sharing complete API query results. 
@@
 experience, this 
-process is incomplete; that is, many tweets submitted to the Twitter API return 
-no data. Additionally, on repeated querying of the API, different sets of tweets 
+process is incomplete; that is, many tweet identification numbers submitted to the Twitter API return 
+no data. Additionally, on repeated querying of the API, different sets of tweet identification numbers
 return data. This complicates our goal of making all analyses computationally 
 reproducible and motivates our decision to share the tweet IDs of those tweets that 
-we actually analyzed (Supplemental files).
+we actually analyzed (Supplementary files). Should a reader wish to reproduce 
+our analysis, we anticipate that she will get complete tweet data
+for all or most of these tweet identification numbers from the API. We provide 
+R code for this task in the supplementary files.
 
@@
 The 31 tweet key-value pairs belong to
-`r length(unique(data_dictionary$Type))` distinct classes (Appendix 1). The 
+`r length(unique(data_dictionary$Type))` distinct classes (Supplementary files). The 
 classes are either vectors - numeric, logical, or character - or arrays assembled 
@@
 chosen topic. 
-The goal for latent Dirichlet allocation is to infer both the distribution over topics and the topics [@blei2003latent]. A topic, in this setting, is a distribution over 
+The goal for latent Dirichlet allocation is to infer both the distribution over topics and 
+the topics [@blei2003latent]. A topic, in this setting, is a distribution over 
 the vocabulary (the collection of all words in a corpus).
@@
 
-## Study design
+# Case study results
 
-We sought to validate our hypothesis that we could detect a social 
-media event by examining tweet topic content at distinct time periods. 
-As a proof of principle of our event detection strategy, we analyzed tweets before, during, and 
-after Memorial Day (May 25, 2020). We fitted latent Dirichlet allocation models for each of three
-distinct five-minute periods. The first period began at noon Eastern time on May 24, 2020. Subsequent time periods started 
-24 and 48 hours later. 
-We defined each time period to be a single collection, or corpus, of tweets.
+We identified the top ten most probable terms for each of ten topics in our models 
+(Figures \@ref(fig:may24), \@ref(fig:may25), \@ref(fig:may26)). 
+We plotted the within-topic word probabilities as bar graphs.
 
-# Results
-
-We identified the top ten terms for each of ten topics in our models. We plotted 
-the within-topic word probabilities as bar graphs.
-
-```{r, fig.cap = "Top terms for LDA model from May 24, 2020"}
+```{r, may24, fig.cap = "Top terms for LDA model from May 24, 2020"}
 knitr::include_graphics("../results/beta-2020-05-24.png")
@@
 
-```{r, fig.cap = "Top terms for LDA model from May 25, 2020 (Memorial Day)"}
+```{r, may25, fig.cap = "Top terms for LDA model from May 25, 2020 (Memorial Day)"}
 knitr::include_graphics("../results/beta-2020-05-25.png")
@@
 
-```{r, fig.cap = "Top terms for LDA model from May 26, 2020"}
+```{r, may26, fig.cap = "Top terms for LDA model from May 26, 2020"}
 knitr::include_graphics("../results/beta-2020-05-26.png")
@@
 words that suggest Memorial Day: memorial, remember, honor, country. A similar topic is not 
-seen on May 24 or May 26. Some topics persist, possibly with distinct word probabilities, 
+seen on May 24 or May 26. Some topics persist, with distinct word probabilities, 
 across the three days. For example, we see that President Trump features prominently in all 
@@
 We conducted formative assessments through weekly discussions with students.
-In these discussions, we developed action items to advance research progress and overcome challenges. We summatively assessed student achievement at the end of the academic year. Both
-students wrote a thesis and presented a poster to our statistics department. 
+In these discussions, we developed action items to advance research progress and overcome 
+challenges. We summatively assessed student achievement at the end of the academic year. 
+Both students wrote a thesis and presented a poster to our statistics department. 
 We asked questions at the poster session to probe student understanding and critically
@@
 
-With future students, we will use a written rubric to evaluate theses (Table \@ref(tab:summative)). We'll
+With future students, we will use a written rubric to evaluate 
+theses (Table \@ref(tab:summative)). We'll
 share the rubric with our students at the start of the academic year.
@@
 
-Tweet collections over time are a rich, large, authentic data set that offer 
-many opportunities for student research projects. We provided 
-instructions to enable readers to 
-establish their own tweet collections. We also presented details for one analysis strategy.
-
-Our mentoring in data science aligns with others' calls to reconsider the role of computing in statistics
+Our mentoring in data science aligns with others' calls to reconsider 
+the role of computing in statistics
 and data science [@nolan2010computing;@carver2016guidelines]. @hicks2018guide argue for
-incorporating three concepts into data science classes: computing, connecting and creating. 
+incorporating three concepts into data science training: 
+computing, connecting and creating. 
 They use the terms "connecting" and "creating" to describe the processes of applying 
-quantitative methods to real data and real research questions and of research question 
-formulation, respectively. Our tweet analysis projects offer students opportunities in all
+quantitative methods to real data and research questions and of formulating research 
+questions, respectively. Our tweet analysis projects offer students opportunities in all
 three skills sets. Our students first formulated research questions, then collected and 
@@
 formulating a research question, gathering data to address it, summarizing the 
-data, visualizing results, and communicating findings. By considering first student 
-research interests and integrating them with our senior thesis learning objectives, 
-we successfully guided two undergraduate researchers in data science research with
-tweets.
+data, visualizing results, and communicating findings. Tweets over time are a rich, large, 
+authentic data set that offer many opportunities. 
+We provided instructions to enable readers to establish their own tweet collections. 
+We also presented details for one analysis strategy.
+By considering first student research interests and integrating them with our 
+senior thesis learning objectives, we successfully guided two undergraduate
+researchers in data science research with tweets.
 
@@
 
-# Appendix 1: Tweet data dictionary
+# Supplementary files
 
-Twitter shares a data dictionary for tweets (https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/tweet-object,  (Accessed: May 23, 2020)).
-We have saved it as a supplementary file, "tweets-data-dictionary.csv".
+## Student projects
 
-\newpage
-
-# Appendix 2: R analysis code
-
-We include our R code in supplementary files.
+1. Jinyu poster
+1. Jinyu report
+2. useR 2016 slides
+3. political methods poster 2016

### 2020-06-08 supplementary/report.tex
added supp files

\documentclass[10pt,letterpaper]{article}
\usepackage{listings}
\usepackage{geometry}
\usepackage{courier,amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{amssymb,amsmath}
\usepackage{blindtext}
\usepackage{color} %red, green, blue, yellow, cyan, magenta, black, white
\usepackage{graphicx}
\usepackage{sidecap}
\usepackage{threeparttable}
\usepackage{subfig}
\usepackage{ textcomp }
\usepackage{ wasysym }

\begin{document}
\font\myfont= cmr12 at 16pt
\title{\myfont Stock Index Prediction Using the Analysis of Twitter Sentiment}
\author{Jinyu Xia}
\maketitle

\begin{abstract}
This paper explores the correlation between Twitter sentiment and the stock index. We first generate Twitter sentiment valences, and examine the potential weekly pattern in the sentiment valence time series by using SLT. Based on the Twitter sentiment analysis, we then apply multiple statistical and machine learning techniques to predict the stock market crashes and the S\&P 500's daily trends.  
\end{abstract}

\section{Introduction} 
In Economics, Efficient Market Hypothesis suggests that the stock market is efficient enough to instantly react to all information. However, like most traditional economics theories, this hypothesis is based on one important assumption -- that all individuals are rational, which in fact, is unlikely to be satisfied in most of the real-world situations. Behavioral economics believe that people's decision-making process is affected by their emotions. In other words, in the real-world economy system, individuals are not always rational. Therefore, studying and quantifying people's emotions would add to methods currently available to predict economic trend such as the stock index.

Former researchers have measured the collective public mood states derived from large-scale Twitter feeds\cite{bollen2011twitter}. It is found that Twitter mood states significantly improve the accuracy of Dow Jones Industrial Average (DJIA) prediction. The results show that emotions can greatly affect individual decision-making.

This paper will derive Twitter sentiment by based on the word emotion lexicon developed by Mohammad et al\cite{mohammad2010emotions}. Based on Word Emotion Association Lexicon, developed by National Research Council (NRC) of Canada. The Twitter sentiment analysis in this paper contains three parts, which are sentiment valence generation, the weekly pattern recognition, and anomaly detection. The sentiment information will then be used not only to predict the stock index trend, but also to identify several corresponding stock crashes in 2015.

In addition, instead of using DJIA, we use the Standard\&Poor's 500 (S\&P 500) as the index of the stock market. The main difference between the S\&P 500 and DJIA is that the S\&P 500 is a market value weighted index of 500 North American Stocks that have a range of sector representations, whereas DJIA calculate only 30 North American stocks by a simple method of mathematics average. In consideration of the number of stocks covered, we select S\&P 500.

\section{Unsupervised Learning of Postings on Twitter}
\subsection{Twitter Data Raw Process}
The corpus of tweets used in this project is downloaded via the Twitter API in R, which includes 8,640,000 tweets per day for almost a year. Much of the raw data is unreadable, so the first step is to clean all non-letter symbols, and filter out non-English tweets by using pattern matching and replacements in R. The second step is to filter tweets by stock-related keywords.

Tweets contents cover a wild range of topics. Note that it is meaningless to use sentiments of tweets talking about sports or TV shows to predict the stock index, and there are lots of tweets that do not even have real meaning, despite using English words. Therefore, to make the prediction of the stock index is more precise, we filter everyday tweets by keywords ``stock'', and several other words closely related to stock market (DJIA, Wall Streets etc.). We expect that the sentiment of tweets posted by people who care about stock market could more accurately reflect the emotions of individuals of our economic system.

After the raw filtering, about one-third of the original downloaded data remained. For each day, there are no more than 1000 tweets left. The Twitter feeds are dated from March 2015, to September 2015. About 10 days of Twitter data are missing due to the failure of a program used to download Twitter data. There are 205 days of tweets in total. Since stock market is closed on weekends and a certain holidays, there are 143 corresponding days of the S\&P 500 value.

\subsection{Generating Twitter Sentiment} 
To get the sentiment of each tweet, we have looked into two R packages which can generate Twitter sentiment. One is called \texttt{sentiment}, the other one is called \texttt{syuzhet}. At first, we prefered the package \texttt{sentiment} more since it claims to apply Bayes' Algorithm to calculate Twitter sentiment. However, we find there are several mistakes in the package source code. Hence, the reliability of the results obtained from this package cannot be guaranteed. With a more careful analysis, we believe Bayes' Theorem is a limited approach to be applied to derive the Twitter sentiment, which makes the mistakes in \texttt{sentiment} unavoidable. By comparison, using a simple voter algorithm based on a reliable dictionary might be a more solid approach, although it is a very simple algorithm.

\subsubsection{Bayes' Theorem}
 The R pacakge \texttt{sentiment} has a lexicon of 6518 common English words. It assigns two parameters to each of the words. One tells whether this word is positive, negative, or neutral, and the other one indicates whether this word represents a strong or weak topic. Based on this lexicon, a function called  \texttt{classify\_polarity} calculates the polarity of a posting, by employing Bayes' theorem.

Given an input tweet, let us call it T. Suppose that we want to calculate the probability that T is positive, the formula of the Bayes' theorem is shown in Eq.\ref{e1}.

\begin{align}\label{e1}
\mathbb{P}(\text{T is positive}|\text{T}) &= \frac{\mathbb{P}(\text{T is positive, } \text{T})}{\mathbb{P}(\text{T})}  
\end{align}
\\
where $\mathbb{P}$(T) is approximated as the proportion of this words in the tweet T that are in the lexicon. The calculation of the numerator on the right hand side of Eq.\ref{e1} is more complicated. It involves calculating the conditional probability:

\begin{align}\label{e2}
\mathbb{P}(\text{T is positive, T}) = \mathbb{P}(\text{T}|\text{T is positive}) \times \mathbb{P}(\text{T is positive})
\end{align}
\\
\noindent The package function assumes that the appearance of each words in the tweet is independent. And thus the equation can be further transformed in to Eq.\ref{e3}.

\begin{align}\label{e3}
\prod_{\text{word}\in \text{T}} \mathbb{P}(\text{ith word}|\text{T is positive})\times \mathbb{P}(\text{T is positive}) 
\end{align}
\\
\noindent Nonetheless, this package function makes several invalid assumptions in these steps of calculation. First, the assumption of words' independence of each other is not valid in many situations, since in most text, the appearance of words are correlated with each other. Second, in the last step the package approximates the probability of tweet T being a positive tweet by using the proportion of positive words in the lexicon, overlooking the number of words in a tweet and the commonness of each word. Similarly, note that for Eq.\ref{e1}, the package approximates $\mathbb{P}(\text{T})$ by using the percentage of words in the lexicon of a tweet, which is also not an accurate calculation. In addition, it does not introduce the source of the lexicon used in the package, and thus we feel that the results obtained by \texttt{sentiment} are not reliable enough.

The problems of \texttt{sentiment} package reflect some difficulties of using Bayes' algorithm to derive Twitter sentiment. First of all, it is hard to calculate the probability of the appearance of a given tweet, and the probability of it being positive or negative. Additionally, to simplify the calculation we have to assume that the independence of each word, while it is a common sense that many words appears in a context dependently.

Despite that \texttt{sentiment} is a problematic package, the Bayes' algorithm is a practical approach to generate Twitter sentiment. More details of using Bayes' theorem to calculate Twitter sentiment is worthy of exploration in the future.

\subsubsection{Emotion Lexicon}
Our second approach for generating Twitter sentiment is to use voter algorithm based on the NRC Emotion Lexicon developed by Mohammad et al\cite{mohammad2010emotions}. The NRC Emotion Lexicon a high-quality emotion lexicon based on manual evluation conducted by using Mechanical Turk\cite{callison2009fast}. The researchers create Human Intelligence Tasks (HITs) for terms in the lexicon, and each HIT has a set of emotion survey questions. The lexicon now include 14183 English words, and each word corresponde with eight basic emotions (anger, fear, anticipation, trust, surprise, sadness, joy, and disgust) and two sentiments (negative and positive).

\begin{figure}[!ht]
    \centering
    \subfloat[Data Fitting]{{\includegraphics[width=.46\textwidth]{histogram1.eps} }}%
    \qquad
    \subfloat[Sparsity Tradeoff]{{\includegraphics[width=.46\textwidth]{histogram2.eps} }}%
    \caption{\small Emotion and sentiment pie charts of tweets on August 24, 2015. This day is reported as the Wall Street's worst day in four years.}
    \label{f1}%
\end{figure}

\begin{figure} [!ht]
	\centering
    \includegraphics[width=\textwidth]{nrc}
    \caption{\small Twitter sentiment valence time series}
    \label{f2}
\end{figure}

The emotion scores and sentiment valence calculation of each tweet is done by using R package \texttt{syuzhet}. It uses an algorithm similar to the Boyer–Moore majority vote algorithm. Since in the Twitter data clean process, non-letter symbols, including punctuations are excluded, and thus we need to treat each tweet as one sentence. Given one tweet, the presence of eight different emotions as well as a positive and negative valence in it are counted. The emotions scores and sentiment valence of a single day are the mean emotions scores and sentiment valence of hundreds of tweets on that day, respectively.

To give a simple overview of the emotion statistical data, we present an example in which the Twitter feeds on Monday, Aug 24, 2015 are analyzed. This day is an important day in 2015 stock market selloff. On this day, the Wall Street suffered a major starting loss as the Dow opened 1,000 points down, and it was also the S\&P 500's worst day since 2011. The pie charts of the Twitter emotion show that there is more than 75\% Twitter feeds present negative sentiment, and about 60\% present emotions such fear, anger, sadness, and disgust. The statistical data suggest people were expressing their negative sentiment in their Twitter posts on August 15, which is an evidence showing the existence of a relation  between stock market states and Twitter sentiments.

The Twitter sentiments of Twitter feeds of 205 days are shown in Fig.\ref{f2}. At first glance, we can see that this time series not a homogeneous. There is fierce fluctuation in the last half of sentiment valence data. We will explore more about it in the section of anomaly detection.

\section{Discovering Anomalies in Twitter Sentiment Valences}

Before conducting the S\&P 500 index prediction, one thing we notice is that the stock market crashes occationally. At the time when we were collecting Twitter data, a stock market crash happened to the Chinese stock market. This crash started in June 2015 and continued into July and August. It only caused a significant loss of paper wealth in China, but also set off a global rout. According to the British national daily newspaper, the Guardian, on August 24, 2015, China sent the world market into a spin, and the US stocks suffered a biggest one-day fall since 2011. The Dow Jones has ended the day down 588 points, or 3.5\%; the S\&P 500 lost around 3.9\%, and the tech-heavy Nasdaq shed 3.8\%. This 2015 market selloff inspired us to examine if the tweets we collected could convey any useful information that could be used as a warning sign of the future stock market crashes. The basic idea we have is to detect sudden spikes or dips in the Twitter sentiment time series, and probe the association between the detected Twitter sentiment ``crash'' and stock market crash.

\subsection{What is Anomalies}
The definition of anomaly, or outlier, is an element which does not follow the behavior of the majority. A concept similar to anomaly is change point, whereas the change point is not the same as an anomaly. Change point detection is the problem of discovering time points at which properties of time-series data change. Since the dataset of this project is not large enough, and also the interest of this section is to discover a certain days in which people's sentiments have violent changes, but not to find which parts of the given time series have different mean or variance.

Note that we have two time series data: the S\&P 500, and Twitter sentiment calculated by NRC method. However, the interest of this section is only to detect the anomalies in the Twitter sentiment data, but not the S\&P 500. There are two reasons for doing it. First, the general public has objective opinions towards what is an anomaly in the stock market, since news media is very sensitive to any abnormal violent changes in the stock market. Additionally, overall the S\&P 500 is relatively a stable time series compared with most dataset. Hence, it is hard to use algorithms to detect anomalies in it.

Nonetheless, it is necessary to use academic approaches to detect anomalies of Twitter sentiment. As opposed to stock indexes, Twitter sentiment is a new concept for people, it is difficult for them to conclude on which days sentiments data have substantial changes. Thus, using professional methods to detect anomalies of the Twitter sentiment scores is essential.

\subsection{Seasonal Hybrid Extreme Studentized Deviate test}
In this section, we use an R package called \texttt{AnomalyDetection}, which could automatically detect anomalies in big data in a practical and robust way. The algorithm used to detect anomalies in Twitter sentiment valence is called Seasonal Hybrid Extreme Studentized Deviate (ESD) test. ESD test is also known as Grubb's test, which is a statistical test used to detect outliers in a univariate data set assumed to come from a normally distribued population. The null hypothesis of EST test is that there is no outliers in the data set. EST test detects one outlier at a time, until there are no more outlier detected. The EST test's statistic is defined as: 
\begin{align}\label{e4}
	G = \frac{\max\limits_{i=1,\ldots,N} \vert Y_i - \bar{Y} \vert} {s}
\end{align}

where $\bar{Y}$ and $s$ denoting the sample mean and standard deviation, respectively. The hypothesis of no outliers is rejected at significance level $\alpha$ if: 
\begin{align}\label{e5}
	G > \frac{N-1}{\sqrt{N}} \sqrt{\frac{t^2_{\alpha/(2N),N-2}}{N-2 + t^2_{\alpha/(2N),N-2}}}
\end{align}

Notice that ESD test is based on the assumption of normality, but many real data are not normally distributed. Real data, however, may have some seasonality. Therefore, one approach to deal with non-normal data is to first fit its seasonal adjustment. This is where Seasonal-Trend Decomposition (STL) by Loess comes in\cite{cleveland1990stl}. STL decomposes seasonal time series into trend, seasonal, and remainder, by using locally-weighted regression, or LOESS. STL has a simple design that consists of a sequence of applications of the loess smoother. Other features of STL are specification of amounts of seasonal and tread smoothing that range.

The \texttt{AnomalyDetection} package uses the Seasonal Hybrid ESD algorithm, which combines seasonal decomposition with robust statistical methods to idnetify local and global anomalies. This is might be an ideal package applied to our Twitter sentiment time series, since at Twitter, a distinct seasonal patterns could be observed in most of the time series. An anomaly can be detected when this point is extreme enough to extend beyond the natural seasonal variation in the time seires (Twitter calls these ``global anomalies''). Also this algoirthm could be used to idenfiy ``local anomalies'', which are small variations on the seasonal trend, but which do not extend the beyong the usual range of values.

\subsection{Detection: Warning Sign of Stock Market Crashes}
We applied the \texttt{AnomalyDetection} package to both Twitter positive and negative valence time series we obtained before. Interestingly, for the positive valence time series, there is no anomaly found, but there are four anomalies detected in the negative valence time series. The date of the anomalies are listed as below:

\begin{table}[!ht]
\centering
\caption{Twitter Negative Sentiment Valence, Anomalies}
\label{t1}
\begin{tabular}{l|llll}
    &Date        & timestamp     & anomalies          & expected value  \\ \hline
1   &2015-06-29  &01:00:00       &1.181598            &0                 \\
2   &2015-07-27  &01:00:00       &1.259012            &0                 \\
3   &2015-08-23  &01:00:00       &1.076377            &0                 \\
4   &2015-08-24  &01:00:00       &1.166600            &0                 \\

\end{tabular}
\end{table}

We mark these four anomalies in the negative valence time series. As we can see, these four points do have obvious deviation from the behavior of the majority. The corresponding points of the S\&P 500 is also market, and as we can see, on all of the days of the sentiment anomalies, there are a sudden drop of the S\&P 500 index. Also notes that the distribution of Twitter negative valence time series starting from July 2015 becomes very strange and unusual. In regard of observation, a further change-point analysis is probably worth conducting on the twitter sentiment time series, which may be helpful for predicting stock market crashes.

\begin{figure}[!ht]
    \centering
    \includegraphics[width=\textwidth]{anomaly}
    \caption{\small Comparison}
    \label{f3}

\end{figure}

\section{Stock Market Index Trend Prediction} 
In this section, we will explore several machine learning supervised models to examine the potential relationships between the derived Twitter sentiment and the S\&P 500 index. We treat the problem as a classification task, whereas the label is the daily up and down of the S\&P 500 index. In section 4.1, we conduct Least features that may help to predict the up the daily trend of the stock index. In section 4.2, we use regularized methods, Lasso and Rigide, to improve the generalization of the learned model. We've also tried logistic regression and support vector machines (SVM) on the S\&P 500 index prediction. Logictisc regressio yields error rates slight worse than LS, and SVM yield a very undesirable results, which is probably due to the small size of features we have.

\subsection{Least Square Binary Classification}
We first use the Least Square (LS) binary classification. The label $y$ is the daily trend of the S\&P 500 index. A daily up and down is labeled as 1 and -1, respectively. The original features contain the positive and negative Twitter sentiment valence from the current day to 5 days ago; also included in the feature sets are the labels from one day ago to 5 days ago. Thus, the number of features $n$ is 17.

A common method for estimating the performance of a classifier is cross-validation (CV). We divide the dataset into 9 equal sized subsets, each of 14 data (so that we only use 126 out of 138 days' of Twitter sentiment). Note that if the dataset is not time series, we can simply use 8 sets of the data to choose the weights, then use the weights to predict the labels of the remaining ``hold-out'' set, and compute the number of mistakes made on this hold-out set and divide that number by 14 (the size of the set). The last step is to repeat this process 9 times (for the 8 different choices of the hold-out set) and average the error rates to obtain a final estimate.

 However, since our dataset is a time series, we cannot use the future data as the training set while the past data make up the test set. Hence, the estimation of the model performance is slightly different here. We still use 9 folders, but we cannot cross validate the error rate by repeating the process of randomly splitting the set of data into 9 folders of size 14. Alternatively, we can only use the folders that contain the past data to predict a hold-out folder that contains the future data. An approach that is sometimes more principled for time series is forward chaining, where the procedure is shown below:

\begin{itemize}
\item folder 1: training [1], test [2]
\item folder 2: training [1 2], test [3]
\item folder 3: training [1 2 3], test [4]
\item  ...
\item folder 8: training [1 2 3 ... 8], test [9]
\end{itemize}

This forward chaining method enables us to model on past data and predict on forward-looking data. The error rates of the 8 test folders are: 0.5000, 0.4286, 0.2143, 0.3571, 0.3571, 0.2871, 0.2143 and 0.1429. As we can tell, the overall trend is that as the size of the training set grows, the error rate becomes smaller. When the test set is the last fold, and all the ``past'' folders are used as the training set, the error rate is the smallest. This makes sense since more training set give us more information for the model fitting. In the best situation, there are 2 misclassifications of the daily trend of the S\&P 500 index out of 14 days.

\subsection{Refinement: Regularized Least Square Optimization}
One challenging part of our prediction problem is how to choose good features. In section 4.1, the size of the initial feature sets $n$ is 17, whereas 12 of them are Twitter positive and negative valence, and 5 of them are past daily trends of the stock index. Among the 12 valence features, there may exist potential data dependency. The correlation matrix of the 12 features shows that the positive valence score of current day and one day ago have the largest correlation, which is 0.47. A few other couple of features also have correlation higher than 0.4. Although they are not desperately high correlations, but fitting 12 similar features together need to certainly be handled carefully, so that we do not overfit the training data.  
In this section, we will discuss two common regularized least squares (RLS) to improve the generalization of our model by constraining it at training time.

\subsubsection{$l_2$ regularizer: Ridge}

\begin{figure}[!ht]
    \centering
    \subfloat[LSData Fitting]{{\includegraphics[width=.46\textwidth]{LS_FC.eps} }}%
    \qquad
    \subfloat[Performance of Ridge Regression]{{\includegraphics[width=.46\textwidth]{Ridge_FC.eps} }}%
    \caption{Least Squares vs. Ridge}%
    \label{f4}%
\end{figure}

We will first try Ridge (aka Tikhonov) regularization. Ridge is a RLS problem with $\ell_2$ regularizer: 
\begin{equation}\label{e6}
\min\limits_{\boldsymbol \beta} \| \boldsymbol{y} - \mathbf{X}\mathbf{\boldsymbol \beta}\|^2 + \lambda \|\mathbf{\boldsymbol \beta}\|_2
\end{equation}

\noindent where $\lambda$ is for penalizing the model if the sum of the 2-norms of the fitted weights get too high. In other words, Ridge regression can avoid over-fitting by regularizing the weights to keep them small. Eq.\ref{e6} has a closed-form algebraic solution $\mathbf{\hat{\boldsymbol\beta}} = (\boldsymbol{X^{T}X} + \lambda \boldsymbol{I})^{-1}\boldsymbol{y}$. Since $n$ is not too large, we can compute the solution directly. The performance of the ridge parameters is shown in Fig.\ref{f4}(b).

Fig.\ref{f4} shows 7 lines of error rate, corresponding with different choices of $\lambda$ ranging from 1/8 to 4. The overall trend of these lines is similar to the shape of the error rate line in Fig.\ref{f4}(a). The regularization achieves its best performance when we all 8 folders are used as the training set, and $\lambda = 2$ yields the best error rate 0.1429, which the same as the LS method.

As we can see, Ridge regression does not improve the performance of our model. This is due to two reasons. First, the error rate of the LS method has already achieved a small error rate that is hard to exceed. Second, the size of our features is not desperately large; thus Ridge does not serve a useful purpose here. However, the weights fitted by Ridge when $\lambda = 2$ is a good candidate set of parameters for future prediction, since it can avoid potential overheating.

\subsubsection{$l_1$ Regularizer: Lasso}
The regularized least squares optimization problem with $\ell_1$ regularizer is called the Lasso in machine learning: 
\begin{equation}\label{e7}
\min\limits_{\boldsymbol \beta} \| \boldsymbol{y} - \mathbf{X}\mathbf{\boldsymbol \beta}\|^2 + \lambda \|\mathbf{\boldsymbol \beta}\|_1
\end{equation}

The Lasso has some very interesting properties. Most notably, the solution to the optimization above usually has many of the elements of $\boldsymbol \beta$ set to zero. It tends to put weight only on elements of $\boldsymbol \beta$ that correspond to the columns of $\boldsymbol{X}$ that are most predictive of $\boldsymbol{y}$. The larger $\lambda$ is, the sparser the solution $\hat{\boldsymbol \beta}$. Unlike Ridge, we do not have a closed-form solution for the Lasso. The solution to Eq.\ref{e7} is given by the ``soft-threshold'' operation: 
\begin{equation}\label{e8}
\beta_j = sign(y_i)(\|y_i\| - \lambda/2)_{+},
\end{equation}

\begin{figure}[!ht]
    \centering
    \subfloat[Data Fitting]{{\includegraphics[width=.46\textwidth]{Lasso_FC.eps} }}%
    \qquad
    \subfloat[Sparsity Tradeoff]{{\includegraphics[width=.46\textwidth]{Lasso_Sparsity.eps} }}%
    \caption{Performance of Lasso Regression}%
    \label{f5}%
\end{figure}

Despite of the nonexistence of the closed-form solution to the Lasso, iterative solvers can be used to obtain good approximate to the solution at a lower cost. We choose to use the Landweber iteration to the optimization. It could be proved that the bound optimization algorithm to approximate the lasso solution is given by 
\begin{equation}\label{e9}
\beta_{k+1} = sign(\boldsymbol{z}_k)(\|\boldsymbol{z}_k\| - \tau\lambda/2)_{+},
\end{equation}

\noindent where $\boldsymbol{z_k} = \boldsymbol{\beta_k} - \tau \boldsymbol{X^{T}}(\boldsymbol{X\beta_k} - \boldsymbol{y})$. The performance and the weights sparsity of the Lasso is shown in Fig.\ref{f5}. The figure indicates that $\lambda < 1$ yield a lower error rate. When $\lambda$ is 3/4 and 1, the error rate is 0.0714 (i.e. There is only 1 misclassification in the test set), which is a result not achieved by LS method. In addition, Fig.\ref{f5}(b) shows that the Lasso with $\lambda = 1$ or $3/4$ select only half of the original features. Based on this information, we can conclude that the Lasso help improving the generalization of our model and avoid overfitting.

\section{Conclusion}
In this paper, we have described our exploration of the association between sentiment in Twitter posts and economic indicators, S\&P 500 market index. We discusseed several potential methods of retrieving Twitter sentiment, and use the one based on NRC Emotion Lexicon developed by Saif Mohammad et al. We have seen that the anomalies of the negative twitter valence is a good indication of great stock market crash. We have also explored the prediction performance of S\&P 500 daily trend by using machine learning techniques such as LS, Lasso, Rigid, logistic regression, and SVM, and presented the results of the models which yield decent results. The best prediction result is seen in Lasso regression ($\lambda$ = 3/4 or 1) by using all the first 8 folders as training set to predict the last folder of the data. In this setting, there is only one missclassification of the total 14 response variables. Based on our data analysis and conclusion, several improvement could be made to increase the prediction rate. One possibility is to increase the data set; due to the time constraint, the data set size of our project is no more than 200. Another promising adaptation to made is to find more features that might be helpful for stock market index prediction. Those could be features not related to twitter sentiment, such as unemployment rate and oil prices.

\newpage
\bibliographystyle{plain}
\bibliography{ref.bib}

\end{document}

### 2020-06-13 Rmd/tweets-one.Rmd
submitted to JSE

@@
 fn <- paste0("../data/", "{{date}}", "-analyzed-tweet-ids.csv")
-#fn <- paste0("../data/", "2020-05-10", "-analyzed-tweet-ids.csv")
 tr_fn <- paste0("../data/", "{{date}}", "-rehydrated-tweets.rds")
-#tr_fn <- paste0("../data/", "2020-05-10", "-rehydrated-tweets.rds")
 sleep <- FALSE
@@
   tweets <- recover_stream(path = "{{tweet_json_file}}", verbose = TRUE) 
-  #tweets <- recover_stream(path = "../data/2020-05-26 12:00:02-tweets.json", verbose = TRUE) 
   id_fn <- paste0("../data/tweets-status-ids-", "{{date}}", ".csv")
-  #id_fn <- paste0("../data/tweets-status-ids-", "2020-05-26", ".csv")
   tweets %>%

### 2020-08-14 Rmd/reply-to-reviewers.Rmd
added Rmd for reviewer comments

---
title: "Replies to reviewers"
output: bookdown::pdf_document2
---

Reviewer: 1

Comments to the Author
General Comments:

The revised manuscript “What is happening on Twitter? A framework for student research projects with tweets” provides a case study for facilitating undergraduate research projects using twitter and topic modeling. The manuscript is well-written and is of interest to JSE readers, and is a strong “datasets and stories” article.

I have only two general suggestions: first, it would be nice to have a brief description of the two undergraduate projects within the manuscript. While I understand the data is no longer available, and think it was the right decision to focus on the case study for methods and results, including a short paragraph about each of the undergraduate projects you supervised would help to give the reader a sense of the range of possibilities of a twitter-based project. This is briefly mentioned at the end of the manuscript but it would help to see the motivations for such a project earlier.

Second, while I appreciate the concise and clear writing, adding transitions between the sections would help the reader anticipate what is coming. As one example, prior to Sec 5.2, there is no indication that the authors will walk through (a) querying twitter API (b) accessing full tweets via ID numbers (c) parsing JSON.

Specific Comments:

- Page 2 Line 28: Should there be a reference for “rhetoric in recruiting political supporters”?
- Page 2 Line 52: I’d recommend adding a statement along the lines of: “different statistics and data science researchers may categorize these skills differently, but our assignment is based on our program/specific projects.” For example, I imagine some projects using twitter data may not categorize “Use text analysis tools to analyze tweets” as “Enduring Understanding”, but would categorize “Use data visualization to clarify and inform quantitative analysis” in such a way.
- Page 4 Sec 4.2: I agree with the authors that differences in student research interests and goals should be an important consideration and it would be nice to see some examples.
- Page 4 Line 82: Would this type of project also be appropriate for a summer research experience?
- Page 4 Line 88-92: This wording could be a little more clear: is the case study essentially a replicate of a student project? It would be helpful to see an overview of what the original project was, and then introduce the case study as a replicate.
- Page 5 Sec 5.2: would this process theoretically work for both windows and Mac machines? Is it necessary to use a crontab-like process, or could it be done entirely within R?
- Page 6 Sec 5.4: This overview of JSON tweet data is very useful.
- Page 7 Line 133: Include citations for rtweet and tidy text appear here, where they’re mentioned first, instead of in the following section.
- Page 8 Line 152: It would be helpful to see a brief model equation.
- Page 8 Figure 1:“beta” has not been mentioned before appearing in this graph (related to previous comment)
- Page 11 Line 162-163: “… our interest is in the transient appearance of a new topic” — this could be made more clear at the beginning of the case study. Another sentence or two motivating what to expect for results would be helpful.

Reviewer: 2

Comments to the Author
Summary: This paper describes a case study using Twitter data and latent Dirichlet allocation to find topics in social media data. Overall, I felt like this edit of the paper was much stronger than the previous submission, much more useful to me personally and to a broad audience of statistical educators. I particularly appreciate the grading rubric!

As always with case studies, I think this is useful at a variety of different levels. First, it provides a template for doing a similar project with a student (i.e., a year-long independent research project). Then, it gives example code for a task that might otherwise be complex. And finally, if an instructor is just looking for an interesting example dataset, they could use your code to grab (and perhaps clean) data for their students to analyze, without exposing them to the entire pipeline.

Major Issues: N/A

Minor Issues:

I think sections 3 and 4 could be combined into one section that is perhaps “Structure of mentored research” or something like that. Perhaps there would be fewer subsections under that section header, but more details on how the research was structured.

Throughout the existing sections, the verb tenses were strange. Since this is now a case study, I expected everything to be grounded in what you actually did. But, on page 3 it says, “These are our four learning objectives.” Why not, “were”? Did you develop the learning objectives after the fact?
At the bottom of the page there are some general statements about students in your statistics department, but not about the particular expertise of the two students who did the project. Did they have existing expertise in linux computing? It says that you guided students toward supplementary resources— did they not have previous R experience?

Section 4.2 says that “an initial brainstorming session may clarify.” Why not “helped”? Is an initial brainstorming session just something you recommend/plan to do in the future, or did you actually do it with those students? You say that you anticipate that sharing completed student project reports will guide future students, but this is from the 2015-2016 school year. Since this project, has it become easier to talk to students about goals, grounded in examples like this? Back to the students in 2015-2016, what goals did you eventually land on?

I assume that those discussions with students involved conversations about the learning objectives from the previous section, as well as the expectations for students (presentation at the end of the year, perhaps some amount of weekly work?).

I found the case study pretty easy to follow starting at 5.3, but I personally have never used crontab before, so that was the most intimidating piece of this. I’d guess that my experience is pretty similar to many JSE readers, who may be familiar with R but not unix/linux commands. Can you insert a short description of what crontab is, and a reference to how to learn more? I’d guess that the instructions you provided would work on Mac, and probably on an RStudio server or RStudio Cloud, but not on a Windows computer. You may want to explicitly state this.

I like that you’ve shared the results from the analysis in the paper! One of my first observations was that there are a lot of swear words in the topics. You may want to address this directly in the paper. Real world data is messy! And may not be “safe for work.” If faculty are going to supervise students in research like this, they need to be prepared for that.

I also audibly gasped when I saw the days you were collecting data. It doesn’t look like George Floyd’s death has made it into the topics, unless it is part of topic 10 from May 26, but again, may be worth an explicit mention. He was killed May 25, and protests began in Minnesota on the 26th.

I really liked the fully-reproducible analysis, but I think it would also be worthwhile to discuss the projects your students actually did. You talked about those projects in the previous version of the paper, and while I know the data is no longer available, perhaps you could incorporate the ideas into the paper somewhere. Either when you talk about question generation, or at the end about other topics. Having a variety of ideas for using tweets might help instructors see that they could either follow your code, or do something related but different.

Typos/copyediting:

p. 3 I think Table 1 would benefit from rules on the left and right to close off the table.
p. 3 “We translated our prioritized list of skills […] into learning objectives” rather than “we translated into learning objectives our prioritized list of skills”
p. 4 I think “R for Data Science” should have headline case
p. 13 - 14 many references need adjustment of capitalization. It looks like this was generated using bibTeX, so just in case this is useful— to force a capital letter you can use curly braces to surround something like {ACM SIGKDD} in the bibtex entry. I noted that Statistical need to be capitalized in the Blei reference, probably Gibbs and Dirichlet should be capitalized in the Porteous reference (as well as the conference name), R in the first Wickham reference.

Reviewer: 3

Comments to the Author
The authors describe a framework for student research using data available using the twitter API.  This is an interesting and relevant paper since many other uses of Twitter data have proven to be somewhat superficial in terms of the richness of data.  The paper is grounded in research on projects.

My main issue relates to the question being answered here: what actionable insights are being extracted from these data?  Is it really just that Memorial Day words show up on Memorial Day?  Are there other questions that could be considered (even if the authors don’t answer them)?  I’m left with the concern that Twitter data is cumbersome to deal with (in terms of scale and restrictions on use of the API) and not particularly information rich.¬

Suggestions:

1)      Table 1 is a nice approach to classify project skills.  I wonder whether it might be productive to connect these skills with other aspects of data science (e.g., the components of data acumen from the NASEM 2018 “Data Science for Undergraduates” report, https://nas.edu/envisioningds)?

2)      Need to add a ref for LDA (page 4): I would suggest capitalizing all words on first use, then using LDA.

3)      Page 5: I was confused by this sentence: “Additionally, on repeated querying of the API, different sets of tweet identification numbers return data.”  Can you clarify/rewrite?

4)      It will be important to note early and in section 6/figure 1 that some tweets may not be suitable for work in terms of inappropriate language.  This is a general issue that might be important for instructors to discuss.

5)      Please provide a link to the Amy Cooper Central Park incident.

6)      I really like the rubric.  How closely tied is it to using Twitter data?  What would be different if some other data source were provided?

7)      I wonder whether the title might be reconsidered.  Perhaps “A framework for student research projects in data science courses”?  The application to Twitter feels secondary to me.

Editor's Comments to Author:

Associate Editor
Comments to the Author:
The manuscript is a substantial revision and provides excellent information for JSE readers.  The referees are generally quite happy with the work, and they offer some general guidelines for additional improvements.

In particular, the referees all spoke about the way that "research question" was addressed in the manuscript.    Although your focus away from the student projects makes sense, bringing in their research question would be a way to address the referees suggestions that twitter data is hard to use for generating hypotheses.  Indeed, you might address the point by speaking to the difference between "data science research" and "political science research".  It seems as though the former was a learning goal and the latter would be difficult to do with such exploratory data as twitter.  Can you expand on that point?

In terms of George Floyd, my guess is that by May 26, the 1% of twitter hadn't get caught wind of the situation?  You might make a comment to that effect.

minor:
* line 128 “Below is an example of Tweet JSON.”  Is that phrasing right?  I might have said “Below is an example of one tweet in JSON format.”  Or “Below is information provided by the Tweet JSON script.”  I’m not sure.  But I guess I don’t know what it is an example *of*.

### 2020-08-14 Rmd/jse-2019-ms.tex

--- 2020-05-25 Rmd/jse-special-issue-2020.tex
@@
 \hypersetup{
-  pdftitle={What is happening on Twitter? Tweet analysis and undergraduate research projects},
+  pdftitle={Analyzing tweets to detect social media events},
   pdfauthor={Frederick J. Boehm and Bret M. Hanlon},
@@
 
-\title{What is happening on Twitter? Tweet analysis and undergraduate research
-projects}
+\title{Analyzing tweets to detect social media events}
 \author{Frederick J. Boehm and Bret M. Hanlon}
@@
 
-Last modified: 2020-05-24 19:58:05.
-
 \hypertarget{introduction}{%
-\subsection{Introduction}\label{introduction}}
+\section{Introduction}\label{introduction}}
 
 Twitter has profoundly changed how we communicate. In only 280
-characters, users instantly contribute to public conversations on
+characters, users can instantly contribute to public conversations on
 politics, current events, sports, media, and many other topics. Recent
@@
 topic content over time. We present below 1) an approach for collecting
-tweets in real time and 2) methods for detecting social media events via
-latent Dirichlet allocation modeling of tweets and 3) suggestions for
-using tweets in research mentoring of undergraduate students.
+tweets in real time and 2) statistical methods for detecting social
+media events via latent Dirichlet allocation modeling of collections of
+tweets and 3) reflections on using this data set in research mentoring
+of undergraduate students.
 
-\hypertarget{preliminary-research-mentoring-considerations}{%
-\subsection{Preliminary research mentoring
-considerations}\label{preliminary-research-mentoring-considerations}}
+\hypertarget{research-mentoring-considerations}{%
+\subsection{Research mentoring
+considerations}\label{research-mentoring-considerations}}
 
@@
 
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  Student experience with data analysis software
-\item
-  Student research interests and goals
-\item
-  Best practices for reproducibility
-\end{enumerate}
-
-\hypertarget{student-experience-with-data-analysis-software}{%
-\subsubsection{Student experience with data analysis
-software}\label{student-experience-with-data-analysis-software}}
-
-Student experience with data analysis software varies. In our statistics
-department, most students learn elementary R computing skills through
-class assignments. Some students, by concentrating in computer science,
-learn other data analysis software packages, such as Python. Those who
-do undergraduate statistics research often learn advanced topics in R
-computing, such as R package assembly, documentation, and testing. Many
-develop expertise in linux computing and cluster computing, too.
-
 \hypertarget{student-research-interests-and-goals}{%
@@
 goals}\label{student-research-interests-and-goals}}
+
+Student experience with data analysis and statistical analysis software
+Software for statistical analysis and best practices for reproducibility
+
+\hypertarget{student-research-interests-and-goals-1}{%
+\subsubsection{Student research interests and
+goals}\label{student-research-interests-and-goals-1}}
 
@@
 
-\hypertarget{best-practices-for-reproducibility}{%
-\subsubsection{Best practices for
-reproducibility}\label{best-practices-for-reproducibility}}
+\hypertarget{student-experience-with-data-analysis-and-statistical-software}{%
+\subsubsection{Student experience with data analysis and statistical
+software}\label{student-experience-with-data-analysis-and-statistical-software}}
+
+\hypertarget{software-for-data}{%
+\subsubsection{Software for data}\label{software-for-data}}
+
+We used the linux software package \texttt{crontab} to automate the
+execution of R code every five minutes. The tweets are available from
+the Twitter API as JSON. We saved the JSON in plain text files.
+
+We performed all analyses in the R statistical computing environment. We
+used the Rstudio software package to interface with R. We share our R
+code to enhance reproducibility of our findings. Users may also find our
+code helpful when creating their own tweet collections.
 
@@
 \item
-  Teach statistical thinking
+  Teach statistical thinking\\
 \item
-  Focus on conceptual understanding
+  Focus on conceptual understanding\\
 \item
-  Integrate real data with a context and purpose
+  Integrate real data with a context and purpose\\
 \item
-  Foster active learning
+  Foster active learning\\
 \item
-  Use technology to explore concepts and analyze data
+  Use technology to explore concepts and analyze data\\
 \item
@@
 
-Our two students conducted their research projects during the 2015-2016
-academic year. We recommend a full academic year for projects of this
-magnitude, although a one-semester project is possible. Our students
-presented their findings at the statistics department's undergraduate
-poster session near the end of the 2015-2016 academic year.
+We began collecting tweets in 2013. Our two students conducted their
+research projects during the 2015-2016 academic year. We recommend a
+full academic year for projects of this magnitude, although a
+one-semester project is possible. Our students presented their findings
+at the statistics department's undergraduate poster session near the end
+of the 2015-2016 academic year.
 
@@
 Twitter API use agreements forbid users from sharing complete API query
-results. However, Twitter permits users to share tweet identification
+results. However, Twitter enables users to share tweet identification
 numbers. A user can then query a Twitter API to obtain complete tweet
@@
   \StringTok{"id_str"}\NormalTok{: }\StringTok{"850006245121695744"}\NormalTok{,}
-  \StringTok{"text"}\NormalTok{: }\StringTok{"1\textbackslash{}/ Today we\textbackslash{}u2019re sharing our vision for the future of the Twitter API platform!"}\NormalTok{,}
+  \StringTok{"text"}\NormalTok{: }\StringTok{"1\textbackslash{}/ Today we\textbackslash{}u2019re sharing our vision for the future of the Twitter API platform!\textbackslash{}nhttps:\textbackslash{}/\textbackslash{}/t.co\textbackslash{}/XweGngmxlP"}\NormalTok{,}
   \StringTok{"user"}\NormalTok{: }\KeywordTok{\{}
@@
 We used functions from the \texttt{rtweet} R package to parse tweet JSON
-into a data frame. We divided JSON arrays into their component vectors
-and added them to the data frame.
+into a data frame. With \texttt{rtweet} functions, we parsed JSON arrays
+into their component vectors and added them to the data frame.
 
@@
 We then inspected topic contents at each of the four time points.
-
-\hypertarget{results}{%
-\subsubsection{Results}\label{results}}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{knitr}\OperatorTok{::}\KeywordTok{include_graphics}\NormalTok{(}\StringTok{"../results/beta-2020-05-10.png"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}
-\includegraphics[width=29.17in]{../results/beta-2020-05-10} \caption{Top terms for LDA model from May 10, 2020 (Mother's Day)}\label{fig:unnamed-chunk-5}
-\end{figure}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\NormalTok{knitr}\OperatorTok{::}\KeywordTok{include_graphics}\NormalTok{(}\StringTok{"../results/beta-2020-05-11.png"}\NormalTok{)}
-\end{Highlighting}
-\end{Shaded}
-
-\begin{figure}
-\includegraphics[width=29.17in]{../results/beta-2020-05-11} \caption{Top terms for LDA model from May 11, 2020}\label{fig:unnamed-chunk-6}
-\end{figure}
 

### 2020-08-14 Rmd/tweets-one.Rmd

--- 2020-05-29 Rmd/tweets-one.Rmd
@@
   tweets <- recover_stream(path = "{{tweet_json_file}}", verbose = TRUE) 
-  #tweets <- recover_stream(path = "../data/2020-05-26 12:00:02-tweets.json", verbose = TRUE) 
+  #tweets <- recover_stream(path = "../data/2020-05-10 12:00:01-tweets.json", verbose = TRUE) 
   id_fn <- paste0("../data/tweets-status-ids-", "{{date}}", ".csv")
-  #id_fn <- paste0("../data/tweets-status-ids-", "2020-05-26", ".csv")
+  #id_fn <- paste0("../data/tweets-status-ids-", "2020-05-10", ".csv")
   tweets %>%

### 2020-08-14 Rmd/abstract.Rmd
updated abstract
added supp files

--- 2020-05-20 Rmd/abstract.Rmd
@@
 
-We describe methods for collecting tweets from Twitter. With these methods, we 
-curate a large collection of tweets.
-We present methods for social media event detection through text analysis of tweets. 
-We describe our use of tweets in mentored undergraduate research. 
-We also offer other directions for other text analysis of tweets.
+We draw on our experiences with mentoring two students to develop a framework for undergraduate research projects with Twitter data.
+Leveraging backward design principles, we share our learning objectives and
+rubric for summative assessments. To illustrate the value of Twitter as a data source, 
+we detail methods for collecting and analyzing tweets. We conclude by emphasizing
+how Twitter text analysis projects enable students to formulate 
+original research questions, collect and analyze data, and communicate
+findings and their implications.

### 2020-08-14 Rmd/jse-special-issue-2020-anon.Rmd

--- 2020-06-08 Rmd/jse-special-issue-2020.Rmd
@@
 title: "What is happening on Twitter? A framework for student research projects with tweets"
-author: "Frederick J. Boehm and Bret M. Hanlon"
+author: ""
 date: "June 15, 2020"
@@
 ---
-
-Last modified: `r params$lastmod`.
 
@@
 authorship of specific tweets when multiple persons share a single account [@drob]; and 
-rhetoric in recruiting political supporters [@pelled2018little;@wells2016trump].
+rhetoric in recruiting political supporters.
 Recognizing the potential utility of tweets for data science research and teaching, 
@@
 and worth being familiar with [@wiggins2005understanding] (Table \@ref(tab:circle-table)). 
+Nearly all of the skills in Table \@ref(tab:circle-table) are 
+transferable. They apply not merely to thesis projects, but also to
+data science research in general.
 
@@
 
-We note that nearly all of the skills in Table \@ref(tab:circle-table) are 
-transferable in that they apply not merely to the thesis projects, but also to
-data science research in general.
-
 ## Learning objectives
@@
 phrased learning objectives in a manner that enabled their subsequent assessment 
-(Table \@ref(tab:summative)) with formative and summative strategies.
+(Table \@ref(tab:summative)) with formative and summative strategies. These are 
+our four learning objectives:
 
@@
 
-In our experience, student interests vary, and students' ability to articulate research goals 
+Student interests vary, and students' ability to articulate research goals 
 may be limited. An initial brainstorming session may clarify their interests and encourage 
@@
 Additionally, we anticipate that sharing completed student project reports will guide student 
-thinking about the scope of possible projects (Supplemental files).
+thinking about the scope of possible projects (Supplementary files).
 
@@
 project is possible. Our students presented their findings at the statistics department’s 
-undergraduate poster session near the end of the 2015-2016 academic year.
-
-We present below reproducible R code for analyzing data from May 2020. While these
+undergraduate poster session near the end of the 2015-2016 academic year (Supplementary files). We present below reproducible R code for analyzing data from May 2020. While these
 are not the same data that our students analyzed in 2015, the methods and code 
@@
 allocation models to infer topics on three consecutive days centered 
-on Memorial Day 2020. We chose to use this example case study, 
+on Memorial Day 2020. We chose this example case study, 
 instead of the student projects, because of limited data 
@@
 new account on Twitter. With these user credentials, we used the 
-R package `rtweet` to query the API. Because we work with linux operating systems, we used the `crontab` software 
+R package `rtweet` to query the API. We used the linux `crontab` software 
 to repeatedly execute R code to submit API queries. Each 
@@
 opened a text file containing user-specified tasks. We added the following line to the 
-bottom of the file:
+bottom of the file before saving and closing the text file.
 
@@
 
-Users may need to slightly amend the above line to conform to requirements of their operating system's `crontab`.
+Readers may need to slightly amend the above line to conform to requirements of their operating system's `crontab`.
 
@@
 (Figures \@ref(fig:may24), \@ref(fig:may25), \@ref(fig:may26)). 
-We plotted the within-topic word probabilities as bar graphs.
+We plotted the within-topic word probabilities as bar graphs. We see that topic-specific 
+word probabilities seldom exceed 0.05. We also note that some words are heavily weighted in 
+multiple topics. This observation complicates semantic topic interpretation.
 
@@
 
-Twitter data over time offers a wealth of potential research projects. Supplementing 
+Twitter data over time inspires a variety of research projects. Supplementing 
 tweets with public data from other sources multiplies the possibilities. For example, 
@@
 She studied sentiment of finance-related tweets and daily stock market index
-closing prices.
+closing prices (Supplementary files).
 
@@
 data, visualizing results, and communicating findings. Tweets over time are a rich, large, 
-authentic data set that offer many opportunities. 
+authentic data set that offers many opportunities. 
 We provided instructions to enable readers to establish their own tweet collections. 
@@
 
-The authors thank Betsy Colby Davie and Rick Nordheim for helpful discussions and 
-feedback on preliminary versions of the manuscript. We thank the special issue 
+We thank the special issue 
 editors and anonymous reviewers for their constructive comments and suggestions.
@@
 
+## Tweets data dictionary
+
+1. Data dictionary
+
+## R code to reproduce the case study
+
+1. tweets.Rmd
+1. tweets-one.Rmd
+1. recover\_tweets.R
+
 ## Student projects
 
-1. Jinyu poster
-1. Jinyu report
-2. useR 2016 slides
-3. political methods poster 2016
+1. Student 1 poster
+1. Student 1 report
+1. Student 2 useR 2016 slides
+1. Student 2 poster
+
+## Github repository
+
+1.

### 2020-08-22 Rmd/jse-special-issue-2020-october.Rmd
created a new Rmd for revised text

--- 2020-08-14 Rmd/jse-special-issue-2020-anon.Rmd
@@
 title: "What is happening on Twitter? A framework for student research projects with tweets"
-author: ""
+author: "Frederick J. Boehm and Bret M. Hanlon"
 date: "June 15, 2020"
@@
 authorship of specific tweets when multiple persons share a single account [@drob]; and 
-rhetoric in recruiting political supporters.
+rhetoric in recruiting political supporters [@pelled2018little;@wells2016trump].
 Recognizing the potential utility of tweets for data science research and teaching, 
@@
 
-We thank the special issue 
+The authors thank Betsy Colby Davie and Rick Nordheim for helpful discussions and 
+feedback on preliminary versions of the manuscript. We thank the special issue 
 editors and anonymous reviewers for their constructive comments and suggestions.
@@
 
-1. Data dictionary
+1. [Data dictionary](https://github.com/fboehm/jse-2019/blob/master/data/tweets-data-dictionary.csv)
 
@@
 
-1. tweets.Rmd
-1. tweets-one.Rmd
-1. recover\_tweets.R
+1. [tweets.Rmd](https://raw.githubusercontent.com/fboehm/jse-2019/master/Rmd/tweets.Rmd)
+1. [tweets-one.Rmd](https://raw.githubusercontent.com/fboehm/jse-2019/master/Rmd/tweets-one.Rmd)
+1. [recover\_tweets.R](https://raw.githubusercontent.com/fboehm/jse-2019/master/Rmd/recover_tweets.R)
 
@@
 
-1. Student 1 poster
-1. Student 1 report
-1. Student 2 useR 2016 slides
-1. Student 2 poster
+1. Student 1 poster: [Project\_Poster.pdf](https://github.com/fboehm/jse-2019/blob/master/supplementary/Project_Poster.pdf)
+1. Student 1 report: [report.pdf](https://github.com/fboehm/jse-2019/blob/master/supplementary/report.pdf)
+1. Student 2 useR 2016 slides: [user2016boehm.pdf](https://github.com/fboehm/jse-2019/blob/master/supplementary/user2016boehm.pdf) 
+1. Student 2 poster: [warfdiscovery2016boehm.tiff](https://github.com/fboehm/jse-2019/blob/master/supplementary/warfdiscovery2016boehm.tiff)
 
@@
 
-1.
+1. https://github.com/fboehm/jse-2019

### 2020-08-22 Rmd/jse-special-issue-2020-october.Rmd
added text to describe Jinyu's project.

@@
 author: "Frederick J. Boehm and Bret M. Hanlon"
-date: "June 15, 2020"
+date: ""
 params:

### 2020-08-22 Rmd/student-projects.Rmd
added text to describe Jinyu's project.

We briefly describe the two student projects to give readers a better sense of research possibilities with tweets.

One student examined relationships between 
stock market index prices and tweet sentiment. For each day in her 12-month study period, 
she identified stock market related tweets with a key word search. With the complete
texts of stock market-related tweets for each day, she calculated a daily sentiment score and plotted it over time. 
Her sentiment score reflected presence of emotion-associated terms (eg., "happy", "sad", "mad", "scared") 
in tweet texts. Days with more net positive emotion words in the collected tweets received a higher (positive) daily sentiment score, while days with more net negative words received a negative daily sentiment score.

For her final project, she presented parallel plots over time of her daily sentiment scores and daily closing prices of the Standard and Poor's 500 index. She also explored time series analysis methods to quantify relationships between daily closing index prices and daily sentiment scores.

### 2020-08-25 Rmd/jse-special-issue-2020-october.Rmd
added Amy Cooper NYTIMES link
fixed wording of example JSON

@@
 
-Below is an example of Tweet JSON.
+Below is an example of JSON for one tweet.
 
-```{bash, eval = FALSE, fig.cap = "Example Tweet JSON", echo = TRUE}
+```{bash, eval = FALSE, fig.cap = "JSON example for one tweet", echo = TRUE}
 {
@@
 across the three days. For example, we see that President Trump features prominently in all 
-three models. We also note, on May 26, topic 10, which reflects discussion of the Amy Cooper Central Park incident.
+three models. We also note, on May 26, topic 10, which reflects discussion of the Amy Cooper Central Park incident (https://www.nytimes.com/2020/05/26/nyregion/amy-cooper-dog-central-park.html).
 

### 2020-08-25 Rmd/jse-special-issue-2020-october.Rmd
added latent Dirichlet allocation ref (blei 2003) to first occurence of LDA

@@
 event by examining topic content over time. We use latent Dirichlet 
-allocation models to infer topics on three consecutive days centered 
+allocation models [@blei2003latent] to infer topics on three consecutive days centered 
 on Memorial Day 2020. We chose this example case study, 
@@
 with tweet JSON, our next step involved reading and parsing tweets with the R 
-packages `rtweet` and `tidytext`.
+packages `rtweet` [@rtweet-package] and `tidytext` [@tidytext].
 
@@
 The next task is to wrangle the tweet JSON into a data structure suitable for latent Dirichlet allocation modeling. 
-We used functions from the `rtweet` R package to parse tweet JSON into a data frame [@rtweet-package]. 
-We then divided tweet text into words with functions from the `tidytext` R package [@tidytext]. We 
+We used functions from the `rtweet` R package to parse tweet JSON into a data frame. 
+We then divided tweet text into words with functions from the `tidytext` R package. We 
 discarded commonly used "stop words" and emojis.

### 2020-08-27 Rmd/reply-to-reviewers.Rmd
added markdown formatting to reply document

@@
 
-Reviewer: 1
+> Reviewer: 1
 
-Comments to the Author
+> Comments to the Author
 General Comments:
 
-The revised manuscript “What is happening on Twitter?
+> The revised manuscript “What is happening on Twitter?
 A framework for student research projects with tweets” provides a case study for facilitating undergraduate research projects using twitter and topic modeling.
@@
 
-I have only two general suggestions: first, it would be nice to have a brief description of the two undergraduate projects within the manuscript.
+> I have only two general suggestions: first, it would be nice to have a brief description of the two undergraduate projects within the manuscript.
 While I understand the data is no longer available, and think it was the right decision to focus on the case study for methods and results, including a short paragraph about each of the undergraduate projects you supervised would help to give the reader a sense of the range of possibilities of a twitter-based project.
@@
 
-Second, while I appreciate the concise and clear writing, adding transitions between the sections would help the reader anticipate what is coming.
+> Second, while I appreciate the concise and clear writing, adding transitions between the sections would help the reader anticipate what is coming.
 As one example, prior to Sec 5.2, there is no indication that the authors will walk through (a) querying twitter API (b) accessing full tweets via ID numbers (c) parsing JSON.
 
-Specific Comments:
+> Specific Comments:
 
-- Page 2 Line 28: Should there be a reference for “rhetoric in recruiting political supporters”?
-- Page 2 Line 52: I’d recommend adding a statement along the lines of: “different statistics and data science researchers may categorize these skills differently, but our assignment is based on our program/specific projects.” For example, I imagine some projects using twitter data may not categorize “Use text analysis tools to analyze tweets” as “Enduring Understanding”, but would categorize “Use data visualization to clarify and inform quantitative analysis” in such a way.
-- Page 4 Sec 4.2: I agree with the authors that differences in student research interests and goals should be an important consideration and it would be nice to see some examples.
-- Page 4 Line 82: Would this type of project also be appropriate for a summer research experience?
-- Page 4 Line 88-92: This wording could be a little more clear: is the case study essentially a replicate of a student project?
+> - Page 2 Line 28: Should there be a reference for “rhetoric in recruiting political supporters”?
+> - Page 2 Line 52: I’d recommend adding a statement along the lines of: “different statistics and data science researchers may categorize these skills differently, but our assignment is based on our program/specific projects.” For example, I imagine some projects using twitter data may not categorize “Use text analysis tools to analyze tweets” as “Enduring Understanding”, but would categorize “Use data visualization to clarify and inform quantitative analysis” in such a way.
+
+> - Page 4 Sec 4.2: I agree with the authors that differences in student research interests and goals should be an important consideration and it would be nice to see some examples.
+> - Page 4 Line 82: Would this type of project also be appropriate for a summer research experience?
+> - Page 4 Line 88-92: This wording could be a little more clear: is the case study essentially a replicate of a student project?
 It would be helpful to see an overview of what the original project was, and then introduce the case study as a replicate.
-- Page 5 Sec 5.2: would this process theoretically work for both windows and Mac machines? Is it necessary to use a crontab-like process, or could it be done entirely within R?
-- Page 6 Sec 5.4: This overview of JSON tweet data is very useful.
-- Page 7 Line 133: Include citations for rtweet and tidy text appear here, where they’re mentioned first, instead of in the following section.
-- Page 8 Line 152: It would be helpful to see a brief model equation.
-- Page 8 Figure 1:“beta” has not been mentioned before appearing in this graph (related to previous comment)
-- Page 11 Line 162-163: “… our interest is in the transient appearance of a new topic” — this could be made more clear at the beginning of the case study.
+> - Page 5 Sec 5.2: would this process theoretically work for both windows and Mac machines? Is it necessary to use a crontab-like process, or could it be done entirely within R?
+
+**Thanks for this question.
+We've added text to clarify that this process is possible from Windows, Mac, and linux operating systems.
+We've also included text to say that there are R packages that interact with operating system-specific scheduling software, so that a user would only interact with the R GUI.**
+
+> - Page 6 Sec 5.4: This overview of JSON tweet data is very useful.
+
+**Thank you.**
+
+> - Page 7 Line 133: Include citations for rtweet and tidy text appear here, where they’re mentioned first, instead of in the following section.
+
+**We've now moved the citations to the first mention of these packages.**
+
+> - Page 8 Line 152: It would be helpful to see a brief model equation.
+
+**We agree! We've added the equations that describe latent Dirichlet allocation.**
+
+> - Page 8 Figure 1:“beta” has not been mentioned before appearing in this graph (related to previous comment)
+
+**You make a good point. We think that the equations that we've added (to address the previous comment) also resolve this.**
+
+> - Page 11 Line 162-163: “… our interest is in the transient appearance of a new topic” — this could be made more clear at the beginning of the case study.
 Another sentence or two motivating what to expect for results would be helpful.
 
-Reviewer: 2
+**We agree with this suggestion, too.**
 
-Comments to the Author
+> Reviewer: 2
+
+> Comments to the Author
 Summary: This paper describes a case study using Twitter data and latent Dirichlet allocation to find topics in social media data.
@@
 
-As always with case studies, I think this is useful at a variety of different levels.
+> As always with case studies, I think this is useful at a variety of different levels.
 First, it provides a template for doing a similar project with a student (i.e., a year-long independent research project).
@@
 
-Major Issues: N/A
+> Major Issues: N/A
 
-Minor Issues:
+> Minor Issues:
 
-I think sections 3 and 4 could be combined into one section that is perhaps “Structure of mentored research” or something like that.
+> I think sections 3 and 4 could be combined into one section that is perhaps “Structure of mentored research” or something like that.
 Perhaps there would be fewer subsections under that section header, but more details on how the research was structured.
 
-Throughout the existing sections, the verb tenses were strange.
+> Throughout the existing sections, the verb tenses were strange.
 Since this is now a case study, I expected everything to be grounded in what you actually did.
@@
 Did you develop the learning objectives after the fact?
-At the bottom of the page there are some general statements about students in your statistics department, but not about the particular expertise of the two students who did the project.
+
+> At the bottom of the page there are some general statements about students in your statistics department, but not about the particular expertise of the two students who did the project.
 Did they have existing expertise in linux computing?
@@
 
-Section 4.2 says that “an initial brainstorming session may clarify.” Why not “helped”?
+> Section 4.2 says that “an initial brainstorming session may clarify.” Why not “helped”?
 Is an initial brainstorming session just something you recommend/plan to do in the future, or did you actually do it with those students?
@@
 
-I assume that those discussions with students involved conversations about the learning objectives from the previous section, as well as the expectations for students (presentation at the end of the year, perhaps some amount of weekly work?).
+> I assume that those discussions with students involved conversations about the learning objectives from the previous section, as well as the expectations for students (presentation at the end of the year, perhaps some amount of weekly work?).
 
-I found the case study pretty easy to follow starting at 5.3, but I personally have never used crontab before, so that was the most intimidating piece of this.
+> I found the case study pretty easy to follow starting at 5.3, but I personally have never used crontab before, so that was the most intimidating piece of this.
 I’d guess that my experience is pretty similar to many JSE readers, who may be familiar with R but not unix/linux commands.
@@
 
-I like that you’ve shared the results from the analysis in the paper!
+> I like that you’ve shared the results from the analysis in the paper!
 One of my first observations was that there are a lot of swear words in the topics.
@@
 
-I also audibly gasped when I saw the days you were collecting data.
+> I also audibly gasped when I saw the days you were collecting data.
 It doesn’t look like George Floyd’s death has made it into the topics, unless it is part of topic 10 from May 26, but again, may be worth an explicit mention.
@@
 
-I really liked the fully-reproducible analysis, but I think it would also be worthwhile to discuss the projects your students actually did.
+> I really liked the fully-reproducible analysis, but I think it would also be worthwhile to discuss the projects your students actually did.
 You talked about those projects in the previous version of the paper, and while I know the data is no longer available, perhaps you could incorporate the ideas into the paper somewhere.
@@
 
-Typos/copyediting:
+> Typos/copyediting:
 
-p. 3 I think Table 1 would benefit from rules on the left and right to close off the table.
-p. 3 “We translated our prioritized list of skills […] into learning objectives” rather than “we translated into learning objectives our prioritized list of skills”
-p. 4 I think “R for Data Science” should have headline case
-p.
+> p. 3 I think Table 1 would benefit from rules on the left and right to close off the table.
+
+> p. 3 “We translated our prioritized list of skills […] into learning objectives” rather than “we translated into learning objectives our prioritized list of skills”
+
+> p. 4 I think “R for Data Science” should have headline case
+
+> p.
 13 - 14 many references need adjustment of capitalization.
@@
 
-Reviewer: 3
+> Reviewer: 3
 
-Comments to the Author
+> Comments to the Author
 The authors describe a framework for student research using data available using the twitter API.
@@
 
-My main issue relates to the question being answered here: what actionable insights are being extracted from these data?
+> My main issue relates to the question being answered here: what actionable insights are being extracted from these data?
 Is it really just that Memorial Day words show up on Memorial Day?
@@
 
-Suggestions:
+> Suggestions:
 
-1)      Table 1 is a nice approach to classify project skills.
+> 1)      Table 1 is a nice approach to classify project skills.
 I wonder whether it might be productive to connect these skills with other aspects of data science (e.g., the components of data acumen from the NASEM 2018 “Data Science for Undergraduates” report, https://nas.edu/envisioningds)?
 
-2)      Need to add a ref for LDA (page 4): I would suggest capitalizing all words on first use, then using LDA.
+> 2)      Need to add a ref for LDA (page 4): I would suggest capitalizing all words on first use, then using LDA.
 
-3)      Page 5: I was confused by this sentence: “Additionally, on repeated querying of the API, different sets of tweet identification numbers return data.”  Can you clarify/rewrite?
+> 3)      Page 5: I was confused by this sentence: “Additionally, on repeated querying of the API, different sets of tweet identification numbers return data.”  Can you clarify/rewrite?
 
-4)      It will be important to note early and in section 6/figure 1 that some tweets may not be suitable for work in terms of inappropriate language.
+> 4)      It will be important to note early and in section 6/figure 1 that some tweets may not be suitable for work in terms of inappropriate language.
 This is a general issue that might be important for instructors to discuss.
 
-5)      Please provide a link to the Amy Cooper Central Park incident.
+> 5)      Please provide a link to the Amy Cooper Central Park incident.
 
-6)      I really like the rubric.  How closely tied is it to using Twitter data?  What would be different if some other data source were provided?
+> 6)      I really like the rubric.  How closely tied is it to using Twitter data?  What would be different if some other data source were provided?
 
-7)      I wonder whether the title might be reconsidered.  Perhaps “A framework for student research projects in data science courses”?  The application to Twitter feels secondary to me.
+> 7)      I wonder whether the title might be reconsidered.  Perhaps “A framework for student research projects in data science courses”?  The application to Twitter feels secondary to me.
 
-Editor's Comments to Author:
+> Editor's Comments to Author:
 
-Associate Editor
-Comments to the Author:
+> Associate Editor
+> Comments to the Author:
 The manuscript is a substantial revision and provides excellent information for JSE readers.
@@
 
-In particular, the referees all spoke about the way that "research question" was addressed in the manuscript.
+> In particular, the referees all spoke about the way that "research question" was addressed in the manuscript.
 Although your focus away from the student projects makes sense, bringing in their research question would be a way to address the referees suggestions that twitter data is hard to use for generating hypotheses.
@@
 
-In terms of George Floyd, my guess is that by May 26, the 1% of twitter hadn't get caught wind of the situation?  You might make a comment to that effect.
+> In terms of George Floyd, my guess is that by May 26, the 1% of twitter hadn't get caught wind of the situation?  You might make a comment to that effect.
 
-minor:
-* line 128 “Below is an example of Tweet JSON.”  Is that phrasing right?
+> minor:
+> * line 128 “Below is an example of Tweet JSON.”  Is that phrasing right?
 I might have said “Below is an example of one tweet in JSON format.”  Or “Below is information provided by the Tweet JSON script.”  I’m not sure.

### 2020-08-28 Rmd/reply-to-reviewers.Rmd
added more detail to reply to reviewers.

@@
 
+**We agree. We've removed many of the subsection headings and used your suggested section title.**
+
 > Throughout the existing sections, the verb tenses were strange.
@@
 
+**We didn't write out learning objectives at the time of working with these students. Back then, we were new to research mentoring. We now are more explicit while designing projects with students.**
+
 > At the bottom of the page there are some general statements about students in your statistics department, but not about the particular expertise of the two students who did the project.
@@
 It says that you guided students toward supplementary resources— did they not have previous R experience?
+
+**Great point! We now discuss this point, and we think that it strengthens the manuscript.**
 
@@
 
+**Thank you for raising this issue. We agree that it needs clarification. We've added text to address this at lines ___ on page .**
+
 > I assume that those discussions with students involved conversations about the learning objectives from the previous section, as well as the expectations for students (presentation at the end of the year, perhaps some amount of weekly work?).
+
+**Yes, although we were new research mentors when we began working with these two students. Regretably, we didn't structure our conversations as we would now.**
 
@@
 
+**Yes, we completely agree.
+We've added text to explain what crontab is and what it does.
+We've also pointed readers to R packages, for Windows and Mac, that allow one to circumvent the need to directly interact with crontab.**
+
 > I like that you’ve shared the results from the analysis in the paper!
@@
 
+**We completely agree. We've now added text to prepare readers for this finding.**
+
 > I also audibly gasped when I saw the days you were collecting data.
@@
 He was killed May 25, and protests began in Minnesota on the 26th.
+
+**We agree. We've added an explicit mention of this issue in the body of the text.**
 
@@
 
+**Thank you.
+We've now added text about the student projects.
+We've also explicitly pointed out what you say in the last sentence - that there are a variety of ways in which an instructor could choose to use this article.
+We've also added more references to point readers to other researchers' work with Twitter.**
+
 > Typos/copyediting:
@@
 
+**We've added rules as you suggested.**
+
 > p. 3 “We translated our prioritized list of skills […] into learning objectives” rather than “we translated into learning objectives our prioritized list of skills”
 
+**We completely agree. The original wording was awkward, so we've fixed this as you suggested.**
+
 > p. 4 I think “R for Data Science” should have headline case
+
+**We agree, and we've corrected this.**
 
@@
 I noted that Statistical need to be capitalized in the Blei reference, probably Gibbs and Dirichlet should be capitalized in the Porteous reference (as well as the conference name), R in the first Wickham reference.
+
+**Thank you for catching this error. We've now corrected the capitalization issues by using curly braces as you suggested.**
 
@@
 Are there other questions that could be considered (even if the authors don’t answer them)?
-I’m left with the concern that Twitter data is cumbersome to deal with (in terms of scale and restrictions on use of the API) and not particularly information rich.¬
+I’m left with the concern that Twitter data is cumbersome to deal with (in terms of scale and restrictions on use of the API) and not particularly information rich.
+
+**We've attempted to address these concerns by writing more about the original two student projects.**
 
@@
 
+**This is a really nice suggestion! Thank you! we've now added text to connect Table 1 with the data acumen components that you reference.**
+
 > 2)      Need to add a ref for LDA (page 4): I would suggest capitalizing all words on first use, then using LDA.
 
+**We've now added the original 2003 Blei, et al. article as the reference on page 4. We've also capitalized all three words in the first mention, and used LDA subsequently.**
+
 > 3)      Page 5: I was confused by this sentence: “Additionally, on repeated querying of the API, different sets of tweet identification numbers return data.”  Can you clarify/rewrite?
+
+**Thanks for pointing this out. We've now rewritten the sentence in efforts to clarify the point that querying the Twitter API is not fully reproducible.**
 
@@
 
+**We've added a link to a New York Times article.**
+
 > 6)      I really like the rubric.  How closely tied is it to using Twitter data?  What would be different if some other data source were provided?
+
+**Thank you! This is a great point that you make, and, as you seem to suggest, much would not differ with other data sources. We've added text to the manuscript to expand on this point.**
 
@@
 
+**This is an excellent point.
+We've now added text to specify that although our May 26, 2020 tweets occurred after George Floyd's murder, we didn't detect evidence of a high volume of tweets about George Floyd at 12pm Eastern Daylight Time on May 26.
+We suggest that the George Floyd Twitter activity likely increased after our collection time as more viewers watched the video.**
+
 > minor:
@@
 But I guess I don’t know what it is an example *of*.
+
+**Thank you for the suggestion. We've tried to clarify this by rephrasing.**

### 2020-09-05 Rmd/jse-special-issue-2020-october.Rmd
updates per reviewers

@@
 
-We translated into learning objectives our prioritized list of skills that 
-students should be able to do, understand, and know (Table \@ref(tab:circle-table)). We 
-phrased learning objectives in a manner that enabled their subsequent assessment 
-(Table \@ref(tab:summative)) with formative and summative strategies. These are 
+We translatedour prioritized list of skills that 
+students should be able to do, understand, and know into learning objectives (Table \@ref(tab:circle-table)). 
+We phrased learning objectives in a manner that enabled their subsequent assessment 
+(Table \@ref(tab:summative)) with formative and summative strategies. These were 
 our four learning objectives:
@@
 
-We developed research goals with students in a series of discussions. As trainees began their senior research projects, we spoke in detail about:
+We developed research goals with students in a series of discussions.
+As trainees began their senior research projects, we spoke in detail about both their research interests and goals and their experience with data analysis software.
 
-1. Student experience with data analysis software
-1. Student research interests and goals
-
-## Student experience with data analysis software
-
-Student experience with data analysis software varies. In our statistics department, 
+Student experience with statistical software varies. In our statistics department, 
 most students learn elementary R computing skills through class assignments. Some 
@@
 
+One of our two students had extensive experience in statistical computing. In 
+addition to R computing skills, she also worked in Python and excelled in shell scripting. 
+She first learned Python in computer science courses. Our second student had 
+extensive experience with R from his statistics courses. His background 
+enabled him to write an R package as part of his senior project.
+
 Our data science projects required that students possess elementary R computing
 skills. We guided students towards supplementary R computing resources, including 
-the online books "R for data science" [@wickham2016r] and "Advanced R" [@wickham2019advanced].
+the online books "R for Data Science" [@wickham2016r] and "Advanced R" [@wickham2019advanced].
 
@@
 thinking about the scope of possible projects (Supplementary files).
+
+We briefly describe the two student projects to give readers a better sense of research possibilities with tweets. Our first student examined relationships over time between 
+stock market index prices and tweet sentiment. For each day in her 12-month study period, 
+she identified stock market-related tweets with a key word search. With the complete
+texts of stock market-related tweets for each day, she calculated a daily sentiment score and plotted it over time. 
+Her sentiment score reflected presence of emotion-associated terms (eg., "happy", "sad", "mad", "scared") 
+in tweet texts.
+Days with more net positive emotion words in the collected tweets received a higher (positive) daily sentiment score, while days with more net negative words received a negative daily sentiment score.
+
+For her final project, she presented plots over time of her daily sentiment scores and daily closing prices of the Standard and Poor's 500 index.
+She also explored time series analysis methods to quantify relationships between closing index prices and sentiment scores.
 
@@
 To illustrate the value of Twitter data, we present below a 
-reproducible case study. In it, we aim to detect a social media 
-event by examining topic content over time. We use latent Dirichlet 
-allocation models [@blei2003latent] to infer topics on three consecutive days centered 
+reproducible case study. It is essentially a reproduction of our second student's project, but at a 
+distinct time period. In it, we aim to detect a social media 
+event by examining topic content over time. We use Latent Dirichlet 
+Allocation (LDA) models [@blei2003latent] to infer topics on three consecutive days centered 
 on Memorial Day 2020. We chose this example case study, 
@@
 Readers may need to slightly amend the above line to conform to requirements of their operating system's `crontab`.
+Readers who use Mac OS may proceed as we did, while those with Windows operating systems may consider using the R package "taskscheduleR" to schedule API queries via the Windows task scheduler [@taskscheduleR].
 

### 2020-09-06 Rmd/jse-special-issue-2020-october.Rmd
added taskSCheduleR note

--- 2020-08-25 Rmd/jse-special-issue-2020-october.Rmd
@@
 
-Readers may need to slightly amend the above line to conform to requirements of their operating system's `crontab`.
+Mac users may need to slightly amend the above line to conform to requirements of their operating system's `crontab`.
+Windows users may want to schedule task with the R package `taskscheduleR`, which interacts with the Windows task scheduler.
 

### 2020-09-06 Rmd/reply-to-reviewers.Rmd
updates per reviewers

@@
 > - Page 2 Line 28: Should there be a reference for “rhetoric in recruiting political supporters”?
+
+**Yes, we mistakenly omitted them. We've now added key references for this point.**
+
 > - Page 2 Line 52: I’d recommend adding a statement along the lines of: “different statistics and data science researchers may categorize these skills differently, but our assignment is based on our program/specific projects.” For example, I imagine some projects using twitter data may not categorize “Use text analysis tools to analyze tweets” as “Enduring Understanding”, but would categorize “Use data visualization to clarify and inform quantitative analysis” in such a way.
 
+**We've now added a statement like this to the manuscript to reflect this point.**
+
 > - Page 4 Sec 4.2: I agree with the authors that differences in student research interests and goals should be an important consideration and it would be nice to see some examples.
+
+**We've added two examples from our students.**
+
 > - Page 4 Line 82: Would this type of project also be appropriate for a summer research experience?
@@
 
-**We agree! We've added the equations that describe latent Dirichlet allocation.**
+**We agree! We've added the equations that describe Latent Dirichlet Allocation.**
 

### 2020-09-06 Rmd/jse-special-issue-2020-october.Rmd

--- 2020-09-05 Rmd/jse-special-issue-2020-october.Rmd
@@
 we created a collection of tweets over time by repeated querying of the Twitter streaming API. 
+We envisioned this collection as a rich resource for data science research projects. This vision
+grew into two mentored undergraduate student research projects in the 2015-2016 academic year.
 
@@
 
-# Backward design and learning objectives
+# Structure of mentored research
 
@@
 and worth being familiar with [@wiggins2005understanding] (Table \@ref(tab:circle-table)). 
+While other researchers may categorize these skills differently, 
+our assignments reflect our projects' priorities.
 Nearly all of the skills in Table \@ref(tab:circle-table) are 
@@
 readr::read_csv("../data/concepts-backward-design.csv") %>%
-  knitr::kable(caption = "Classifying project skills")
+  knitr::kable(caption = "Classifying project skills") %>%
+  kableExtra::column_spec(column = 1, border_left = TRUE) %>%
+  kableExtra::column_spec(column = 2, border_right = TRUE)
 ```
@@
 
-We translatedour prioritized list of skills that 
+We translated our prioritized list of skills that 
 students should be able to do, understand, and know into learning objectives (Table \@ref(tab:circle-table)). 
@@
 
-# Preliminary research mentoring considerations
+## Preliminary research mentoring considerations
 
@@
 extensive experience with R from his statistics courses. His background 
-enabled him to write an R package as part of his senior project.
-
-Our data science projects required that students possess elementary R computing
-skills. We guided students towards supplementary R computing resources, including 
-the online books "R for Data Science" [@wickham2016r] and "Advanced R" [@wickham2019advanced].
+enabled him to write an R package as part of his senior project. To encourage 
+further development of R computing skills in our two students, we guided them 
+towards the online books "R for Data Science" [@wickham2016r] and "Advanced R" [@wickham2019advanced].
 
@@
 
-We briefly describe the two student projects to give readers a better sense of research possibilities with tweets. Our first student examined relationships over time between 
+We briefly describe the two student projects to give readers a better sense 
+of research possibilities with tweets. Our first student examined relationships over time between 
 stock market index prices and tweet sentiment. For each day in her 12-month study period, 
@@
 texts of stock market-related tweets for each day, she calculated a daily sentiment score and plotted it over time. 
-Her sentiment score reflected presence of emotion-associated terms (eg., "happy", "sad", "mad", "scared") 
-in tweet texts.
-Days with more net positive emotion words in the collected tweets received a higher (positive) daily sentiment score, while days with more net negative words received a negative daily sentiment score.
+Her sentiment score reflected presence of emotion-associated terms 
+(*eg*., "happy", "sad", "mad", "scared") 
+in tweet texts. Days with more net positive emotion words in the collected tweets received a higher 
+(positive) daily sentiment score, while days with more net negative words received a negative daily 
+sentiment score. 
+For her final project, she presented plots over time of her daily sentiment scores and daily closing 
+prices of the Standard and Poor's 500 index. She also explored time series analysis methods to quantify 
+relationships between index prices and sentiment scores.
 
-For her final project, she presented plots over time of her daily sentiment scores and daily closing prices of the Standard and Poor's 500 index.
-She also explored time series analysis methods to quantify relationships between closing index prices and sentiment scores.
+Our second student developed social media event detection methods with topic 
+models. He hypothesized that tweet content changes over time, and that we 
+might detect these changes by comparing inferred tweet topics from distinct 
+time periods. To validate his hypothesis, he examined tweet content before, during, and after 
+the National Football League's Super Bowl game in 2015. He reasoned that 
+because the Super Bowl is widely discussed on Twitter, we might detect 
+Super Bowl-related topics from tweets sent during the game, but that the football-related
+topics would be short-lived in the continuous Twitter stream. We discovered
+evidence to support his strategy, and we ultimately presented our findings 
+at international and local research meetings. We share below a case study based
+on this approach.
 
@@
 Our two statistics students conducted their research projects during the 2015-2016 academic year. 
-We recommend a full academic year for projects of this magnitude, although a one-semester 
+We recommend a full academic year for projects of this magnitude, although a summer or one-semester 
 project is possible. Our students presented their findings at the statistics department’s 
-undergraduate poster session near the end of the 2015-2016 academic year (Supplementary files). We present below reproducible R code for analyzing data from May 2020. While these
+undergraduate poster session near the end of the 2015-2016 academic year (Supplementary files). We 
+present below reproducible R code for analyzing data from May 2020. While these
 are not the same data that our students analyzed in 2015, the methods and code 
-are very similar.
+are very similar to that of our second student's project.
 
@@
 across the three days. For example, we see that President Trump features prominently in all 
-three models. We also note, on May 26, topic 10, which reflects discussion of the Amy Cooper Central Park incident (https://www.nytimes.com/2020/05/26/nyregion/amy-cooper-dog-central-park.html).
+three models. We also note, on May 26, topic 10, which reflects discussion of the Amy 
+Cooper Central Park incident (https://www.nytimes.com/2020/05/26/nyregion/amy-cooper-dog-central-park.html).
+
+We also note that the murder of George Floyd occurred on May 25. Our last 
+examined time period, from 12:00 pm to 12:05 pm (Eastern USA time zone) on May 26, occurred after Floyd's murder, 
+yet we didn't detect this event in our ten-topic LDA model. Several considerations 
+may account for this. While outrage at the murder eventually spread worldwide, there may have been few 
+Floyd-related tweets during our collection time on May 26, less than 24 hours after the murder and video release. 
+Had we extended our analysis to May 27 and beyond, we may have identified George Floyd-related topics.
 

### 2020-09-06 Rmd/student-projects.Rmd

@@
 
-One student examined relationships between 
+One student examined relationships over time between 
 stock market index prices and tweet sentiment. For each day in her 12-month study period, 
-she identified stock market related tweets with a key word search. With the complete
+she identified stock market-related tweets with a key word search. With the complete
 texts of stock market-related tweets for each day, she calculated a daily sentiment score and plotted it over time. 

### 2020-09-07 Rmd/jse-special-issue-2020-october.Rmd
added LDA generative process math. notation.

@@
 chosen topic.
+
+In mathematical notation, we write the generative process assumed by LDA [@blei2003latent]:
+
+1. Choose $N \sim \text{Poisson}(\xi)$  
+1. Choose $\theta \sim \text{Dirichlet}(\alpha)$  
+1. For each word, $w_n$ with $n = 1, ..., N$, 
+  a. Choose a topic $z_n \sim \text{Multinomial}(\theta)$  
+  b. Choose a word $w_n$ from $p(w_n | z_n, \beta)$, a multinomial probability
+
+$\beta$ refers to the $k$ by $V$ matrix of topic-specific word probabilities, where $k$
+is the number of topics and $V$ is the size of the vocabulary, *i.e.*, the number of unique words in the corpus.
+
 The goal for latent Dirichlet allocation is to infer both the distribution over topics and 
@@
 Variational methods, while using approximations to the posterior distribution, offer the 
-advantage of computational speed [@blei2017variational]. We used variational methods below.
+advantage of computational speed [@blei2017variational]. We used variational methods below in our case study.
 

### 2020-09-11 Rmd/jse-special-issue-2020-october.Rmd
replaced "latent Dirichlet allocation" with "LDA" abbreviation

@@
 after Memorial Day (May 25, 2020). 
-We fitted latent Dirichlet allocation models for each of three
+We fitted LDA models for each of three
 distinct five-minute periods. The first period began at noon Eastern time 
@@
 
-Readers may need to slightly amend the above line to conform to requirements of their operating system's `crontab`.
-Readers who use Mac OS may proceed as we did, while those with Windows operating systems may consider using the R package "taskscheduleR" to schedule API queries via the Windows task scheduler [@taskscheduleR].
+Readers may need to slightly amend the above line to conform to requirements of 
+their operating system's software. Readers who use Mac OS may proceed as we did, 
+while those with Windows operating systems may consider using the R package 
+"taskscheduleR" to schedule API queries via the Windows task scheduler [@taskscheduleR].
 
@@
 
-The next task is to wrangle the tweet JSON into a data structure suitable for latent Dirichlet allocation modeling. 
+The next task is to wrangle the tweet JSON into a data structure suitable for LDA modeling. 
 We used functions from the `rtweet` R package to parse tweet JSON into a data frame. 
@@
 
-Latent Dirichlet allocation model fitting requires that the corpus be organized as a 
+LDA model fitting requires that the corpus be organized as a 
 document by term matrix. In a document by term matrix, each row corresponds to a single document (a single 
@@
 
-## Latent Dirichlet allocation
+## Latent Dirichlet Allocation
 
-Latent Dirichlet allocation is a statistical method for inferring latent (unobservable) topics (or themes)
+LDA is a statistical method for inferring latent (unobservable) topics (or themes)
 from a large corpus (or collection) of documents [@blei2003latent].
@@
 
-The goal for latent Dirichlet allocation is to infer both the distribution over topics and 
+The goal for LDA is to infer both the distribution over topics and 
 the topics [@blei2003latent]. A topic, in this setting, is a distribution over 
@@
 
-Latent Dirichlet allocation modeling and related methods are a major research area in the 
+LDA modeling and related methods are a major research area in the 
 quantitative social sciences. Advanced students with interest in statistical computing might

### 2020-09-11 Rmd/jse-special-issue-2020-october.Rmd
added rtweet reference at first mention
added mention of curse words
fixed "We also note" issue

@@
 new account on Twitter. With these user credentials, we used the 
-R package `rtweet` to query the API. We used the linux `crontab` software 
+R package `rtweet` to query the API [@rtweet-package]. We used the linux `crontab` software 
 to repeatedly execute R code to submit API queries. Each 
@@
 word probabilities seldom exceed 0.05. We also note that some words are heavily weighted in 
-multiple topics. This observation complicates semantic topic interpretation.
+multiple topics. This observation complicates semantic topic interpretation. We 
+also caution that the results display expletives (that appeared on Twitter) and may 
+not be "safe for work".
 
@@
 across the three days. For example, we see that President Trump features prominently in all 
-three models. We also note, on May 26, topic 10, which reflects discussion of the Amy 
+three models. On May 26, topic 10 reflects discussion of the Amy 
 Cooper Central Park incident (https://www.nytimes.com/2020/05/26/nyregion/amy-cooper-dog-central-park.html).
 
-We also note that the murder of George Floyd occurred on May 25. Our last 
+The murder of George Floyd occurred on May 25. Our last 
 examined time period, from 12:00 pm to 12:05 pm (Eastern USA time zone) on May 26, occurred after Floyd's murder, 

### 2020-09-11 Rmd/jse-special-issue-2020-october.Rmd
minor text edits

@@
 
-We developed research goals with students in a series of discussions.
-As trainees began their senior research projects, we spoke in detail about both their research interests and goals and their experience with data analysis software.
+We developed research goals with students in a series of brainstorming sessions and discussions. 
+As trainees began their senior 
+research projects, we spoke in detail about both their research interests and goals and their 
+experience with data analysis software. When possible, we encouraged them to incorporate their 
+existing academic interests into their senior research projects.
 
-Student experience with statistical software varies. In our statistics department, 
-most students learn elementary R computing skills through class assignments. Some 
-students, by concentrating in computer science, learn other data analysis software 
+In our statistics department, most students learn elementary R computing skills through class 
+assignments. Some students, by concentrating in computer science, learn other data analysis software 
 packages, such as Python. 
@@
 further development of R computing skills in our two students, we guided them 
-towards the online books "R for Data Science" [@wickham2016r] and "Advanced R" [@wickham2019advanced].
+towards the free, online books "R for Data Science" [@wickham2016r] and "Advanced R" [@wickham2019advanced].
 
@@
 
-Student interests vary, and students' ability to articulate research goals 
-may be limited. An initial brainstorming session may clarify their interests and encourage 
-them to think critically about goals under the time constraints of their academic schedules. 
-Additionally, we anticipate that sharing completed student project reports will guide student 
-thinking about the scope of possible projects (Supplementary files).
+Our two students had diverse interests, and, initially, they had little experience in 
+articulating research goals. We engaged each in a brainstorming session to 
+clarify their interests and encourage them to think critically about research goals under 
+the time constraints of their academic schedules. 
+Additionally, we anticipate that sharing completed student project reports with 
+future trainees will guide thinking about the scope of possible projects (Supplementary files).
 

### 2020-09-15 Rmd/reply-to-reviewers.Rmd
added text on NASEM report and data acumen.

@@
 
+**Thank you for this suggestion! We've now added text that describes the two projects early in the manuscript.**
+
 > Second, while I appreciate the concise and clear writing, adding transitions between the sections would help the reader anticipate what is coming.
 As one example, prior to Sec 5.2, there is no indication that the authors will walk through (a) querying twitter API (b) accessing full tweets via ID numbers (c) parsing JSON.
+
+**We appreciate this suggestion, too. We've worked to add transitions between sections with the goal of better guiding the reader's expectations.**
 
@@
 > - Page 4 Line 82: Would this type of project also be appropriate for a summer research experience?
+
+**Yes, we've amended the text to reflect this.**
+
 > - Page 4 Line 88-92: This wording could be a little more clear: is the case study essentially a replicate of a student project?
 It would be helpful to see an overview of what the original project was, and then introduce the case study as a replicate.
+
+**We've reworded this to clarify the point. Yes, the case study essentially replicates a student project.**
+
 > - Page 5 Sec 5.2: would this process theoretically work for both windows and Mac machines? Is it necessary to use a crontab-like process, or could it be done entirely within R?
@@
 
-**We agree with this suggestion, too.**
+**We agree with this suggestion, too. We have tried to clarify it by adding a sentence to this section.**
 
@@
 
-**We didn't write out learning objectives at the time of working with these students. Back then, we were new to research mentoring. We now are more explicit while designing projects with students.**
+**We didn't write out learning objectives at the time of working with these students, but we did informally discuss them.
+Back then (2015), we were new to research mentoring.
+We now are more explicit while designing projects with students.**
 
@@
 
-**Thank you for raising this issue. We agree that it needs clarification. We've added text to address this at lines ___ on page .**
+**Thank you for raising this issue. We agree that it needs clarification. We've added text to address this at lines.**
 
@@
 
-**Yes, although we were new research mentors when we began working with these two students. Regretably, we didn't structure our conversations as we would now.**
+**Yes, although we were new research mentors when we began working with these two students. Regrettably, we didn't structure our conversations as we would now.**
 
@@
 
-**Thank you.
-We've now added text about the student projects.
-We've also explicitly pointed out what you say in the last sentence - that there are a variety of ways in which an instructor could choose to use this article.
-We've also added more references to point readers to other researchers' work with Twitter.**
+**Thank you. We've now added text about the student projects. We've also added more references to point readers to other researchers' work with Twitter.**
 
@@
 
-**We've attempted to address these concerns by writing more about the original two student projects.**
+**Thank you. We've attempted to address these concerns by writing more about the original two student projects.**
 

### 2020-09-15 Rmd/jse-special-issue-2020-october.Rmd
deleted blank line to combine paragraphs
added text on NASEM report and data acumen.

@@
 
+In particular, our project skills reflect elements of "data acumen", as defined 
+in a report from the U.S.A. National Academies of Sciences, Engineering, and Medicine (NASEM) [@national2018data]. For example, our skills "Acquire data from internet sources" and 
+"Develop data science strategies to address a research question" implicitly require 
+a trainee, in the language of the NASEM report [@national2018data], to "ingest, clean, and then 
+wrangle data into reliable and useful forms" and to "combine many existing programs or codes
+into a workflow that will accomplish some important task". Additionally, we tailored our list 
+of project skills with the assumption that students would work in R. R use is not required for such projects, but it is a 
+convenience for many of our students.
+
 ## Learning objectives
@@
 1. Develop data science strategies to address a scientific research question.
+
+Having decided on four learning objectives for students, we next established
+mentoring relationships with each student.
 
@@
 the time constraints of their academic schedules. 
-Additionally, we anticipate that sharing completed student project reports with 
-future trainees will guide thinking about the scope of possible projects (Supplementary files).
+We briefly describe the two student projects to give readers a better sense 
+of research possibilities with tweets.
 
-We briefly describe the two student projects to give readers a better sense 
-of research possibilities with tweets. Our first student examined relationships over time between 
+Our first student examined relationships over time between 
 stock market index prices and tweet sentiment. For each day in her 12-month study period, 
@@
 availability for the student projects. Despite this, the case study 
-illustrates the strategy and methods for one of the student 
-projects.
+illustrates the strategy and methods for one student 
+project.
+
+We next discuss case study design, tweet collection, and tweet structure, before turning 
+to statistical methods for the case study.
 
@@
 the vocabulary (the collection of all words in a corpus). 
-
 Inference for latent Dirichlet allocation models is performed by either 
@@
 our interest is in the transient appearance of a new topic, we don't attempt to
-assign meaning to every topic in our models. We see that topic 7 from May 25 has several 
+assign meaning to every topic in our models.
+Instead, we anticipate that discussions on Twitter are a mixture of topics that endure over weeks or months and subjects that appear and disappear quickly.
+We see that topic 7 from May 25 has several
 words that suggest Memorial Day: memorial, remember, honor, country. A similar topic is not 
@@
 across the three days. For example, we see that President Trump features prominently in all 
-three models. On May 26, topic 10 reflects discussion of the Amy 
+three models' results. On May 26, topic 10 reflects discussion of the Amy 
 Cooper Central Park incident (https://www.nytimes.com/2020/05/26/nyregion/amy-cooper-dog-central-park.html).
 
-The murder of George Floyd occurred on May 25. Our last 
+The murder of George Floyd occurred on May 25, 2020. Our last 
 examined time period, from 12:00 pm to 12:05 pm (Eastern USA time zone) on May 26, occurred after Floyd's murder, 

### 2020-09-16 Rmd/reply-to-reviewers.Rmd
incorporated reviewer comments.

@@
 
-**Thank you for raising this issue. We agree that it needs clarification. We've added text to address this at lines.**
+**Thank you for raising this issue. We agree that it needs clarification. We've added text to address this.**
 
@@
 
-**Yes, we completely agree.
-We've added text to explain what crontab is and what it does.
-We've also pointed readers to R packages, for Windows and Mac, that allow one to circumvent the need to directly interact with crontab.**
+**Yes, we completely agree. We've added text to explain what crontab is and what it does. We've also pointed readers to an R package for Windows**
 
@@
 
+**Thank you. We've added text to warn readers about this.**
+
 > 5)      Please provide a link to the Amy Cooper Central Park incident.
@@
 > 7)      I wonder whether the title might be reconsidered.  Perhaps “A framework for student research projects in data science courses”?  The application to Twitter feels secondary to me.
+
+**We appreciate the suggestion.
+We chose the title with hopes of promoting the use of tweets and text analysis in student research projects.
+With all of our supplementary materials explicitly dealing with tweets, we feel more comfortable with the current title.**
 

### 2020-09-16 Rmd/jse-special-issue-2020-october.Rmd
incorporated reviewer comments.

@@
 process is incomplete; that is, many tweet identification numbers submitted to the Twitter API return 
-no data. Additionally, on repeated querying of the API, different sets of tweet identification numbers
-return data. This complicates our goal of making all analyses computationally 
+no data. Additionally, some tweet identification numbers return data on the first query, but don't return data on subsequent queries. This complicates our goal of making all analyses computationally 
 reproducible and motivates our decision to share the tweet IDs of those tweets that 
@@
 theses (Table \@ref(tab:summative)). We'll
-share the rubric with our students at the start of the academic year.
+share the rubric with our students at the start of the academic year. With only minor
+modifications, the rubric may be suitable for projects that don't use tweets.
 

### 2020-09-16 Rmd/jse-special-issue-2020-october.Rmd
added text about Box 1976

@@
 user can freely 
-access this one percent sample, whereas access to a larger selection is available to researchers for a fee.
+access this one percent sample, whereas access to a larger selection is available to 
+researchers for a fee.
+
+Through our work with tweets, we demonstrate
+that Twitter data is a rich source of new data science research questions. 
+@box1976science described a 
+positive feedback loop for the interplay of discipline-specific research
+and quantitative methods research. The two components, "science" and "statistics" in the 
+language of @box1976science, iteratively fuel research questions in each other. 
+A new statistical 
+method enables new discipline-specific questions to be addressed, while a new 
+scientific question motivates new data science methods. We describe below two
+data science research questions that mentored students addressed with tweets.
 
@@
 
-To illustrate the value of Twitter data, we present below a 
+To illustrate the value of Twitter data and to encourage readers to envision other uses for tweets, we present below a 
 reproducible case study. It is essentially a reproduction of our second student's project, but at a 
 distinct time period. In it, we aim to detect a social media 
-event by examining topic content over time. We use Latent Dirichlet 
+event by examining tweet topic content over time. We use Latent Dirichlet 
 Allocation (LDA) models [@blei2003latent] to infer topics on three consecutive days centered 
@@
 illustrates the strategy and methods for one student 
-project.
-
-We next discuss case study design, tweet collection, and tweet structure, before turning 
-to statistical methods for the case study.
+project. Below, we discuss case study design, tweet collection, and tweet structure, 
+before turning to statistical methods for the case study.
 

### 2020-10-06 Rmd/reply-to-reviewers.Rmd
added text about Box 1976

@@
 ---
-title: "Replies to reviewers"
-output: bookdown::pdf_document2
+output: 
+  bookdown::pdf_document2:
+    toc: false
 ---
+
+Dear editors and reviewers:
+
+We present below our replies to the very helpful comments that you shared
+when reviewing the manuscript. We include both the text of the comments and (in bold) our 
+replies. 
+We are grateful for the constructive feedback that you've shared. Our manuscript
+is better because of it.
 
@@
 
-**We agree. We've removed many of the subsection headings and used your suggested section title.**
+**We agree. We now use your suggested section title.**
 
@@
 
+**Thanks for raising this point. We've now added text in the introduction that addresses the interplay between data science research and discipline-specific research, like that in political science.**
+
 > In terms of George Floyd, my guess is that by May 26, the 1% of twitter hadn't get caught wind of the situation?  You might make a comment to that effect.

### 2020-10-06 Rmd/reply-to-reviewers.Rmd
updated reply to reviewers per Bret's suggestions

@@
     toc: false
+    number_sections: false
 ---
+
+October 6, 2020
 
@@
 
-> Reviewer: 1
+Sincerely,
 
-> Comments to the Author
+Fred Boehm and Bret Hanlon
+
+## Editor's Comments to Author:
+
+## Associate Editor
+
+Comments to the Author:
+The manuscript is a substantial revision and provides excellent information for JSE readers.
+The referees are generally quite happy with the work, and they offer some general guidelines for additional improvements.
+
+In particular, the referees all spoke about the way that "research question" was addressed in the manuscript.
+Although your focus away from the student projects makes sense, bringing in their research question would be a way to address the referees suggestions that twitter data is hard to use for generating hypotheses.
+Indeed, you might address the point by speaking to the difference between "data science research" and "political science research".
+It seems as though the former was a learning goal and the latter would be difficult to do with such exploratory data as twitter.
+Can you expand on that point?
+
+> **Thanks for raising this point.
+We've now added text in the introduction that addresses the interplay between data science research and discipline-specific research, like that in political science.**
+
+In terms of George Floyd, my guess is that by May 26, the 1% of twitter hadn't get caught wind of the situation?  You might make a comment to that effect.
+
+> **This is an excellent point.
+We've now added text to specify that although our May 26, 2020 tweets occurred after George Floyd's murder, we didn't detect evidence of a high volume of tweets about George Floyd at 12pm Eastern Daylight Time on May 26.
+We suggest that the George Floyd Twitter activity likely increased after our collection time as more viewers watched the video.**
+
+minor:
+* line 128 “Below is an example of Tweet JSON.”  Is that phrasing right?
+I might have said “Below is an example of one tweet in JSON format.”  Or “Below is information provided by the Tweet JSON script.”  I’m not sure.
+But I guess I don’t know what it is an example *of*.
+
+> **Thank you for the suggestion. We've tried to clarify this by rephrasing.**
+
+## Reviewer: 1
+
+Comments to the Author
 General Comments:
 
-> The revised manuscript “What is happening on Twitter?
+The revised manuscript “What is happening on Twitter?
 A framework for student research projects with tweets” provides a case study for facilitating undergraduate research projects using twitter and topic modeling.
@@
 
-> I have only two general suggestions: first, it would be nice to have a brief description of the two undergraduate projects within the manuscript.
+I have only two general suggestions: first, it would be nice to have a brief description of the two undergraduate projects within the manuscript.
 While I understand the data is no longer available, and think it was the right decision to focus on the case study for methods and results, including a short paragraph about each of the undergraduate projects you supervised would help to give the reader a sense of the range of possibilities of a twitter-based project.
@@
 
-**Thank you for this suggestion! We've now added text that describes the two projects early in the manuscript.**
+> **Thank you for this suggestion! We've now added text that describes the two projects early in the manuscript.**
 
-> Second, while I appreciate the concise and clear writing, adding transitions between the sections would help the reader anticipate what is coming.
+Second, while I appreciate the concise and clear writing, adding transitions between the sections would help the reader anticipate what is coming.
 As one example, prior to Sec 5.2, there is no indication that the authors will walk through (a) querying twitter API (b) accessing full tweets via ID numbers (c) parsing JSON.
 
-**We appreciate this suggestion, too. We've worked to add transitions between sections with the goal of better guiding the reader's expectations.**
+> **We appreciate this suggestion, too. We've worked to add transitions between sections with the goal of better guiding the reader's expectations.**
 
-> Specific Comments:
+Specific Comments:
 
-> - Page 2 Line 28: Should there be a reference for “rhetoric in recruiting political supporters”?
+- Page 2 Line 28: Should there be a reference for “rhetoric in recruiting political supporters”?
 
-**Yes, we mistakenly omitted them. We've now added key references for this point.**
+> **Yes, we mistakenly omitted them. We've now added key references for this point.**
 
-> - Page 2 Line 52: I’d recommend adding a statement along the lines of: “different statistics and data science researchers may categorize these skills differently, but our assignment is based on our program/specific projects.” For example, I imagine some projects using twitter data may not categorize “Use text analysis tools to analyze tweets” as “Enduring Understanding”, but would categorize “Use data visualization to clarify and inform quantitative analysis” in such a way.
+- Page 2 Line 52: I’d recommend adding a statement along the lines of: “different statistics and data science researchers may categorize these skills differently, but our assignment is based on our program/specific projects.” For example, I imagine some projects using twitter data may not categorize “Use text analysis tools to analyze tweets” as “Enduring Understanding”, but would categorize “Use data visualization to clarify and inform quantitative analysis” in such a way.
 
-**We've now added a statement like this to the manuscript to reflect this point.**
+> **We've now added a statement like this to the manuscript to reflect this point.**
 
-> - Page 4 Sec 4.2: I agree with the authors that differences in student research interests and goals should be an important consideration and it would be nice to see some examples.
+- Page 4 Sec 4.2: I agree with the authors that differences in student research interests and goals should be an important consideration and it would be nice to see some examples.
 
-**We've added two examples from our students.**
+> **We've added two examples from our students.**
 
-> - Page 4 Line 82: Would this type of project also be appropriate for a summer research experience?
+- Page 4 Line 82: Would this type of project also be appropriate for a summer research experience?
 
-**Yes, we've amended the text to reflect this.**
+> **Yes, we've amended the text to reflect this.**
 
-> - Page 4 Line 88-92: This wording could be a little more clear: is the case study essentially a replicate of a student project?
+- Page 4 Line 88-92: This wording could be a little more clear: is the case study essentially a replicate of a student project?
 It would be helpful to see an overview of what the original project was, and then introduce the case study as a replicate.
 
-**We've reworded this to clarify the point. Yes, the case study essentially replicates a student project.**
+> **We've reworded this to clarify the point. Yes, the case study essentially replicates a student project.**
 
-> - Page 5 Sec 5.2: would this process theoretically work for both windows and Mac machines? Is it necessary to use a crontab-like process, or could it be done entirely within R?
+- Page 5 Sec 5.2: would this process theoretically work for both windows and Mac machines? Is it necessary to use a crontab-like process, or could it be done entirely within R?
 
-**Thanks for this question.
+> **Thanks for this question.
 We've added text to clarify that this process is possible from Windows, Mac, and linux operating systems.
@@
 
-> - Page 6 Sec 5.4: This overview of JSON tweet data is very useful.
+- Page 6 Sec 5.4: This overview of JSON tweet data is very useful.
 
-**Thank you.**
+> **Thank you.**
 
-> - Page 7 Line 133: Include citations for rtweet and tidy text appear here, where they’re mentioned first, instead of in the following section.
+- Page 7 Line 133: Include citations for rtweet and tidy text appear here, where they’re mentioned first, instead of in the following section.
 
-**We've now moved the citations to the first mention of these packages.**
+> **We've now moved the citations to the first mention of these packages.**
 
-> - Page 8 Line 152: It would be helpful to see a brief model equation.
+- Page 8 Line 152: It would be helpful to see a brief model equation.
 
-**We agree! We've added the equations that describe Latent Dirichlet Allocation.**
+> **We agree! We've added the equations that describe Latent Dirichlet Allocation.**
 
-> - Page 8 Figure 1:“beta” has not been mentioned before appearing in this graph (related to previous comment)
+- Page 8 Figure 1:“beta” has not been mentioned before appearing in this graph (related to previous comment)
 
-**You make a good point. We think that the equations that we've added (to address the previous comment) also resolve this.**
+> **You make a good point. We think that the equations that we've added (to address the previous comment) also resolve this.**
 
-> - Page 11 Line 162-163: “… our interest is in the transient appearance of a new topic” — this could be made more clear at the beginning of the case study.
+- Page 11 Line 162-163: “… our interest is in the transient appearance of a new topic” — this could be made more clear at the beginning of the case study.
 Another sentence or two motivating what to expect for results would be helpful.
 
-**We agree with this suggestion, too. We have tried to clarify it by adding a sentence to this section.**
+> **We agree with this suggestion, too. We have tried to clarify it by adding a sentence to this section.**
 
-> Reviewer: 2
+## Reviewer: 2
 
-> Comments to the Author
+Comments to the Author
 Summary: This paper describes a case study using Twitter data and latent Dirichlet allocation to find topics in social media data.
@@
 
-> As always with case studies, I think this is useful at a variety of different levels.
+As always with case studies, I think this is useful at a variety of different levels.
 First, it provides a template for doing a similar project with a student (i.e., a year-long independent research project).
@@
 
-> Major Issues: N/A
+Major Issues: N/A
 
-> Minor Issues:
+Minor Issues:
 
-> I think sections 3 and 4 could be combined into one section that is perhaps “Structure of mentored research” or something like that.
+I think sections 3 and 4 could be combined into one section that is perhaps “Structure of mentored research” or something like that.
 Perhaps there would be fewer subsections under that section header, but more details on how the research was structured.
 
-**We agree. We now use your suggested section title.**
+> **We agree. We now use your suggested section title.**
 
-> Throughout the existing sections, the verb tenses were strange.
+Throughout the existing sections, the verb tenses were strange.
 Since this is now a case study, I expected everything to be grounded in what you actually did.
@@
 
-**We didn't write out learning objectives at the time of working with these students, but we did informally discuss them.
+> **We didn't write out learning objectives at the time of working with these students, but we did informally discuss them.
 Back then (2015), we were new to research mentoring.
@@
 
-> At the bottom of the page there are some general statements about students in your statistics department, but not about the particular expertise of the two students who did the project.
+At the bottom of the page there are some general statements about students in your statistics department, but not about the particular expertise of the two students who did the project.
 Did they have existing expertise in linux computing?
@@
 
-**Great point! We now discuss this point, and we think that it strengthens the manuscript.**
+> **Great point! We now discuss this point, and we think that it strengthens the manuscript.**
 
-> Section 4.2 says that “an initial brainstorming session may clarify.” Why not “helped”?
+Section 4.2 says that “an initial brainstorming session may clarify.” Why not “helped”?
 Is an initial brainstorming session just something you recommend/plan to do in the future, or did you actually do it with those students?
@@
 
-**Thank you for raising this issue. We agree that it needs clarification. We've added text to address this.**
+> **Thank you for raising this issue. We agree that it needs clarification. We've added text to address this.**
 
-> I assume that those discussions with students involved conversations about the learning objectives from the previous section, as well as the expectations for students (presentation at the end of the year, perhaps some amount of weekly work?).
+I assume that those discussions with students involved conversations about the learning objectives from the previous section, as well as the expectations for students (presentation at the end of the year, perhaps some amount of weekly work?).
 
-**Yes, although we were new research mentors when we began working with these two students. Regrettably, we didn't structure our conversations as we would now.**
+> **Yes, although we were new research mentors when we began working with these two students. Regrettably, we didn't structure our conversations as we would now.**
 
-> I found the case study pretty easy to follow starting at 5.3, but I personally have never used crontab before, so that was the most intimidating piece of this.
+I found the case study pretty easy to follow starting at 5.3, but I personally have never used crontab before, so that was the most intimidating piece of this.
 I’d guess that my experience is pretty similar to many JSE readers, who may be familiar with R but not unix/linux commands.
@@
 
-**Yes, we completely agree. We've added text to explain what crontab is and what it does. We've also pointed readers to an R package for Windows**
+> **Yes, we completely agree. We've added text to explain what crontab is and what it does. We've also pointed readers to an R package for Windows**
 
-> I like that you’ve shared the results from the analysis in the paper!
+I like that you’ve shared the results from the analysis in the paper!
 One of my first observations was that there are a lot of swear words in the topics.
@@
 
-**We completely agree. We've now added text to prepare readers for this finding.**
+> **We completely agree. We've now added text to prepare readers for this finding.**
 
-> I also audibly gasped when I saw the days you were collecting data.
+I also audibly gasped when I saw the days you were collecting data.
 It doesn’t look like George Floyd’s death has made it into the topics, unless it is part of topic 10 from May 26, but again, may be worth an explicit mention.
@@
 
-**We agree. We've added an explicit mention of this issue in the body of the text.**
+> **We agree. We've added an explicit mention of this issue in the body of the text.**
 
-> I really liked the fully-reproducible analysis, but I think it would also be worthwhile to discuss the projects your students actually did.
+I really liked the fully-reproducible analysis, but I think it would also be worthwhile to discuss the projects your students actually did.
 You talked about those projects in the previous version of the paper, and while I know the data is no longer available, perhaps you could incorporate the ideas into the paper somewhere.
@@
 
-**Thank you. We've now added text about the student projects. We've also added more references to point readers to other researchers' work with Twitter.**
+> **Thank you. We've now added text about the student projects. We've also added more references to point readers to other researchers' work with Twitter.**
 
-> Typos/copyediting:
+Typos/copyediting:
 
-> p. 3 I think Table 1 would benefit from rules on the left and right to close off the table.
+p. 3 I think Table 1 would benefit from rules on the left and right to close off the table.
 
-**We've added rules as you suggested.**
+> **We've added rules as you suggested.**
 
-> p. 3 “We translated our prioritized list of skills […] into learning objectives” rather than “we translated into learning objectives our prioritized list of skills”
+p. 3 “We translated our prioritized list of skills […] into learning objectives” rather than “we translated into learning objectives our prioritized list of skills”
 
-**We completely agree. The original wording was awkward, so we've fixed this as you suggested.**
+> **We completely agree. The original wording was awkward, so we've fixed this as you suggested.**
 
-> p. 4 I think “R for Data Science” should have headline case
+p. 4 I think “R for Data Science” should have headline case
 
-**We agree, and we've corrected this.**
+> **We agree, and we've corrected this.**
 
-> p.
+p.
 13 - 14 many references need adjustment of capitalization.
@@
 
-**Thank you for catching this error. We've now corrected the capitalization issues by using curly braces as you suggested.**
+> **Thank you for catching this error. We've now corrected the capitalization issues by using curly braces as you suggested.**
 
-> Reviewer: 3
+## Reviewer: 3
 
-> Comments to the Author
+Comments to the Author
 The authors describe a framework for student research using data available using the twitter API.
@@
 
-> My main issue relates to the question being answered here: what actionable insights are being extracted from these data?
+My main issue relates to the question being answered here: what actionable insights are being extracted from these data?
 Is it really just that Memorial Day words show up on Memorial Day?
@@
 
-**Thank you. We've attempted to address these concerns by writing more about the original two student projects.**
+> **Thank you. We've attempted to address these concerns by writing more about the original two student projects.**
 
-> Suggestions:
+Suggestions:
 
-> 1)      Table 1 is a nice approach to classify project skills.
+1.
+Table 1 is a nice approach to classify project skills.
 I wonder whether it might be productive to connect these skills with other aspects of data science (e.g., the components of data acumen from the NASEM 2018 “Data Science for Undergraduates” report, https://nas.edu/envisioningds)?
 
-**This is a really nice suggestion! Thank you! we've now added text to connect Table 1 with the data acumen components that you reference.**
+> **This is a really nice suggestion! Thank you! we've now added text to connect Table 1 with the data acumen components that you reference.**
 
-> 2)      Need to add a ref for LDA (page 4): I would suggest capitalizing all words on first use, then using LDA.
+2. Need to add a ref for LDA (page 4): I would suggest capitalizing all words on first use, then using LDA.
 
-**We've now added the original 2003 Blei, et al. article as the reference on page 4. We've also capitalized all three words in the first mention, and used LDA subsequently.**
+> **We've now added the original 2003 Blei, et al. article as the reference on page 4. We've also capitalized all three words in the first mention, and used LDA subsequently.**
 
-> 3)      Page 5: I was confused by this sentence: “Additionally, on repeated querying of the API, different sets of tweet identification numbers return data.”  Can you clarify/rewrite?
+3. Page 5: I was confused by this sentence: “Additionally, on repeated querying of the API, different sets of tweet identification numbers return data.”  Can you clarify/rewrite?
 
-**Thanks for pointing this out. We've now rewritten the sentence in efforts to clarify the point that querying the Twitter API is not fully reproducible.**
+> **Thanks for pointing this out. We've now rewritten the sentence in efforts to clarify the point that querying the Twitter API is not fully reproducible.**
 
-> 4)      It will be important to note early and in section 6/figure 1 that some tweets may not be suitable for work in terms of inappropriate language.
+4.
+It will be important to note early and in section 6/figure 1 that some tweets may not be suitable for work in terms of inappropriate language.
 This is a general issue that might be important for instructors to discuss.
 
-**Thank you. We've added text to warn readers about this.**
+> **Thank you. We've added text to warn readers about this.**
 
-> 5)      Please provide a link to the Amy Cooper Central Park incident.
+5. Please provide a link to the Amy Cooper Central Park incident.
 
-**We've added a link to a New York Times article.**
+> **We've added a link to a New York Times article.**
 
-> 6)      I really like the rubric.  How closely tied is it to using Twitter data?  What would be different if some other data source were provided?
+6. I really like the rubric.  How closely tied is it to using Twitter data?  What would be different if some other data source were provided?
 
-**Thank you! This is a great point that you make, and, as you seem to suggest, much would not differ with other data sources. We've added text to the manuscript to expand on this point.**
+> **Thank you! This is a great point that you make, and, as you seem to suggest, much would not differ with other data sources. We've added text to the manuscript to expand on this point.**
 
-> 7)      I wonder whether the title might be reconsidered.  Perhaps “A framework for student research projects in data science courses”?  The application to Twitter feels secondary to me.
+7. I wonder whether the title might be reconsidered.  Perhaps “A framework for student research projects in data science courses”?  The application to Twitter feels secondary to me.
 
-**We appreciate the suggestion.
+> **We appreciate the suggestion.
 We chose the title with hopes of promoting the use of tweets and text analysis in student research projects.
 With all of our supplementary materials explicitly dealing with tweets, we feel more comfortable with the current title.**
-
-> Editor's Comments to Author:
-
-> Associate Editor
-> Comments to the Author:
-The manuscript is a substantial revision and provides excellent information for JSE readers.
-The referees are generally quite happy with the work, and they offer some general guidelines for additional improvements.
-
-> In particular, the referees all spoke about the way that "research question" was addressed in the manuscript.
-Although your focus away from the student projects makes sense, bringing in their research question would be a way to address the referees suggestions that twitter data is hard to use for generating hypotheses.
-Indeed, you might address the point by speaking to the difference between "data science research" and "political science research".
-It seems as though the former was a learning goal and the latter would be difficult to do with such exploratory data as twitter.
-Can you expand on that point?
-
-**Thanks for raising this point. We've now added text in the introduction that addresses the interplay between data science research and discipline-specific research, like that in political science.**
-
-> In terms of George Floyd, my guess is that by May 26, the 1% of twitter hadn't get caught wind of the situation?  You might make a comment to that effect.
-
-**This is an excellent point.
-We've now added text to specify that although our May 26, 2020 tweets occurred after George Floyd's murder, we didn't detect evidence of a high volume of tweets about George Floyd at 12pm Eastern Daylight Time on May 26.
-We suggest that the George Floyd Twitter activity likely increased after our collection time as more viewers watched the video.**
-
-> minor:
-> * line 128 “Below is an example of Tweet JSON.”  Is that phrasing right?
-I might have said “Below is an example of one tweet in JSON format.”  Or “Below is information provided by the Tweet JSON script.”  I’m not sure.
-But I guess I don’t know what it is an example *of*.
-
-**Thank you for the suggestion. We've tried to clarify this by rephrasing.**

### 2020-10-06 Rmd/jse-special-issue-2020-october.Rmd
re-make-ed Rmd manuscript file.

@@
 project. Below, we discuss case study design, tweet collection, and tweet structure, 
-before turning to statistical methods for the case study.
+before turning to quantitative methods for the case study.
 
@@
 feedback on preliminary versions of the manuscript. We thank the special issue 
-editors and anonymous reviewers for their constructive comments and suggestions.
+editors and anonymous reviewers for their constructive comments and suggestions. Finally, this work wouldn't have been possible without the keen and enthusiastic students, Jinyu Xia and Robert Turner.
 

### 2020-10-07 Rmd/jse-special-issue-2020-october.Rmd
fixed yaml to keep tex file

@@
     number_sections: TRUE
+    keep_tex: TRUE
     includes:

### 2020-10-07 Rmd/jse-special-issue-2020-october.tex
fixed yaml to keep tex file

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={What is happening on Twitter? A framework for student research projects with tweets},
  pdfauthor={Frederick J. Boehm and Bret M. Hanlon},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=2.5cm,right=2.5cm,top=2.5cm,bottom=2.5cm]{geometry}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.77,0.63,0.00}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
%\usepackage{lineno}
%\linenumbers
\usepackage{setspace}
\doublespacing
\usepackage{caption}

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}
%\usepackage[utf8x]{inputenc}
\usepackage{rotating}
\usepackage{wrapfig}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\title{What is happening on Twitter? A framework for student research projects with tweets}
\author{Frederick J. Boehm and Bret M. Hanlon}
\date{}

\begin{document}
\maketitle

\hypertarget{abstract}{%
\section{Abstract}\label{abstract}}

We draw on our experiences with mentoring two students to develop a framework for undergraduate research projects with Twitter data.
Leveraging backward design principles, we share our learning objectives and
rubric for summative assessments. To illustrate the value of Twitter as a data source,
we detail methods for collecting and analyzing tweets. We conclude by emphasizing
how Twitter text analysis projects enable students to formulate
original research questions, collect and analyze data, and communicate
findings and their implications.

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Twitter has profoundly changed how we communicate.
In only 280 characters, users instantly contribute to public conversations on politics,
current events, sports, media, and many other topics.
Recent development of accessible statistical methods for large-scale text
analysis now enable instructors to use tweets as contemporary pedagogical
tools in guiding undergraduate research projects.
We guided two statistics students in their senior research projects.
Both students used tweets to address novel research questions. We
share products of their research in supplementary files. Because
their data are no longer available, we present as a case study one
analysis with tweets from May 2020. We share our data and computer
code to encourage others to undertake
tweet text analysis research. We also describe methods for creating a
collection of tweets.

Some social media data, including tweets from Twitter, is
available through website application product interfaces (APIs).
By way of a streaming API, Twitter shares a sample of approximately
one percent of all
tweets during an API query time period (``Sampled stream'' 2019). Any Twitter
user can freely
access this one percent sample, whereas access to a larger selection is available to
researchers for a fee.

Through our work with tweets, we demonstrate
that Twitter data is a rich source of new data science research questions.
Box (1976) described a
positive feedback loop for the interplay of discipline-specific research
and quantitative methods research. The two components, ``science'' and ``statistics'' in the
language of Box (1976), iteratively fuel research questions in each other.
A new statistical
method enables new discipline-specific questions to be addressed, while a new
scientific question motivates new data science methods. We describe below two
data science research questions that mentored students addressed with tweets.

Studies of Twitter conversations have yielded valuable insights into
modern culture. Using large collections of tweets, scholars have
investigated diverse research questions, including
the inference of relationships and social networks among Twitter users (Lin et al. 2011);
authorship of specific tweets when multiple persons share a single account (Robinson 2016); and
rhetoric in recruiting political supporters (Pelled et al. 2018; Wells et al. 2016).
Recognizing the potential utility of tweets for data science research and teaching,
we created a collection of tweets over time by repeated querying of the Twitter streaming API.
We envisioned this collection as a rich resource for data science research projects. This vision
grew into two mentored undergraduate student research projects in the 2015-2016 academic year.

In line with recent calls for students to work with real data (Carver et al. 2016; Nolan and Temple Lang 2010), our
collection of tweets has served as a valuable resource in our mentoring
of undergraduate data science research.
Working with real data allows students to develop proficiency not
only in statistical analysis,
but also in related data science skills, including data
transfer from online sources, data storage, using data
from multiple file formats, and communicating findings.
Collaboratively asking and addressing novel questions with
our collection of tweets gave mentored students
opportunities to develop competency in all of these areas.

While our tweet collection enables us to address many possible research questions, the
dynamic content of tweets over time particularly piqued our interest.
We hypothesized that high-profile social media events would generate a high volume of
tweets, and that we would detect social media events through changes in tweet
topic content over time. We discuss in detail below one approach to studying this
question. In the sections that follow, we detail our backward design-inspired approach to
writing learning objectives, preliminary research mentoring considerations,
data science methods for collecting and analyzing tweets, analysis results,
and ideas on assessment and advanced topics.

\hypertarget{structure-of-mentored-research}{%
\section{Structure of mentored research}\label{structure-of-mentored-research}}

\hypertarget{backward-design}{%
\subsection{Backward design}\label{backward-design}}

Backward design principles guided our planning and informed the writing of
learning objectives (Wiggins and McTighe 2005). Following Wiggins and McTighe (2005),
we began by listing
what students, at the end of their thesis research, should be able to do,
understand, and know. We then classified each of these items into
one of three categories: enduring understanding, important to know and do,
and worth being familiar with (Wiggins and McTighe 2005) (Table \ref{tab:circle-table}).
While other researchers may categorize these skills differently,
our assignments reflect our projects' priorities.
Nearly all of the skills in Table \ref{tab:circle-table} are
transferable. They apply not merely to thesis projects, but also to
data science research in general.

\begin{table}

\caption{\label{tab:circle-table}Classifying project skills}
\centering
\begin{tabular}[t]{|>{}l|>{}l|}
\hline
Skill & Category\\
\hline
Structure research project files as R package & Worth being familiar with\\
\hline
Communicate results in speaking and in writing & Enduring understanding\\
\hline
Formulate a research question & Enduring understanding\\
\hline
Develop data science strategies to address research question & Enduring understanding\\
\hline
Use Github to share code and documentation & Important to know and do\\
\hline
Use git for version control & Important to know and do\\
\hline
Use text analysis tools to analyze tweets & Enduring understanding\\
\hline
Use cluster computing as needed & Worth being familiar with\\
\hline
Use data visualization to clarify and inform quantitative analyses & Important to know and do\\
\hline
Translate analysis results into scientific conclusions & Enduring understanding\\
\hline
Incorporate supplementary data sources into analysis & Important to know and do\\
\hline
Acquire data from internet sources & Important to know and do\\
\hline
Describe assumptions and limitations of statistical analyses & Enduring understanding\\
\hline
\end{tabular}
\end{table}

In particular, our project skills reflect elements of ``data acumen'', as defined
in a report from the U.S.A. National Academies of Sciences, Engineering, and Medicine (NASEM) (National Academies of Sciences et al. 2018). For example, our skills ``Acquire data from internet sources'' and
``Develop data science strategies to address a research question'' implicitly require
a trainee, in the language of the NASEM report (National Academies of Sciences et al. 2018), to ``ingest, clean, and then
wrangle data into reliable and useful forms'' and to ``combine many existing programs or codes
into a workflow that will accomplish some important task''. Additionally, we tailored our list
of project skills with the assumption that students would work in R. R use is not required for such projects, but it is a
convenience for many of our students.

\hypertarget{learning-objectives}{%
\subsection{Learning objectives}\label{learning-objectives}}

We translated our prioritized list of skills that
students should be able to do, understand, and know into learning objectives (Table \ref{tab:circle-table}).
We phrased learning objectives in a manner that enabled their subsequent assessment
(Table \ref{tab:summative}) with formative and summative strategies. These were
our four learning objectives:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Write R code to perform text analysis of large volumes of tweets (R Core Team 2019).
\item
  Communicate results in a written report and poster presentation.
\item
  Translate statistical findings into scientific conclusions.
\item
  Develop data science strategies to address a scientific research question.
\end{enumerate}

Having decided on four learning objectives for students, we next established
mentoring relationships with each student.

\hypertarget{preliminary-research-mentoring-considerations}{%
\subsection{Preliminary research mentoring considerations}\label{preliminary-research-mentoring-considerations}}

We developed research goals with students in a series of brainstorming sessions and discussions.
As trainees began their senior
research projects, we spoke in detail about both their research interests and goals and their
experience with data analysis software. When possible, we encouraged them to incorporate their
existing academic interests into their senior research projects.

In our statistics department, most students learn elementary R computing skills through class
assignments. Some students, by concentrating in computer science, learn other data analysis software
packages, such as Python.
Those who do undergraduate statistics research often learn advanced topics in R
computing, such as R package assembly, documentation, and testing. Many develop
expertise in linux computing and cluster computing, too.

One of our two students had extensive experience in statistical computing. In
addition to R computing skills, she also worked in Python and excelled in shell scripting.
She first learned Python in computer science courses. Our second student had
extensive experience with R from his statistics courses. His background
enabled him to write an R package as part of his senior project. To encourage
further development of R computing skills in our two students, we guided them
towards the free, online books ``R for Data Science'' (Wickham and Grolemund 2016) and ``Advanced R'' (Wickham 2019).

\hypertarget{student-research-interests-and-goals}{%
\subsection{Student research interests and goals}\label{student-research-interests-and-goals}}

Our two students had diverse interests, and, initially, they had little experience in
articulating research goals. We engaged each in a brainstorming session to
clarify their interests and encourage them to think critically about research goals under
the time constraints of their academic schedules.
We briefly describe the two student projects to give readers a better sense
of research possibilities with tweets.

Our first student examined relationships over time between
stock market index prices and tweet sentiment. For each day in her 12-month study period,
she identified stock market-related tweets with a key word search. With the complete
texts of stock market-related tweets for each day, she calculated a daily sentiment score and plotted it over time.
Her sentiment score reflected presence of emotion-associated terms
(\emph{eg}., ``happy'', ``sad'', ``mad'', ``scared'')
in tweet texts. Days with more net positive emotion words in the collected tweets received a higher
(positive) daily sentiment score, while days with more net negative words received a negative daily
sentiment score.
For her final project, she presented plots over time of her daily sentiment scores and daily closing
prices of the Standard and Poor's 500 index. She also explored time series analysis methods to quantify
relationships between index prices and sentiment scores.

Our second student developed social media event detection methods with topic
models. He hypothesized that tweet content changes over time, and that we
might detect these changes by comparing inferred tweet topics from distinct
time periods. To validate his hypothesis, he examined tweet content before, during, and after
the National Football League's Super Bowl game in 2015. He reasoned that
because the Super Bowl is widely discussed on Twitter, we might detect
Super Bowl-related topics from tweets sent during the game, but that the football-related
topics would be short-lived in the continuous Twitter stream. We discovered
evidence to support his strategy, and we ultimately presented our findings
at international and local research meetings. We share below a case study based
on this approach.

\hypertarget{time-period}{%
\subsection{Time period}\label{time-period}}

Our two statistics students conducted their research projects during the 2015-2016 academic year.
We recommend a full academic year for projects of this magnitude, although a summer or one-semester
project is possible. Our students presented their findings at the statistics department's
undergraduate poster session near the end of the 2015-2016 academic year (Supplementary files). We
present below reproducible R code for analyzing data from May 2020. While these
are not the same data that our students analyzed in 2015, the methods and code
are very similar to that of our second student's project.

\hypertarget{case-study-methods}{%
\section{Case study methods}\label{case-study-methods}}

To illustrate the value of Twitter data and to encourage readers to envision other uses for tweets, we present below a
reproducible case study. It is essentially a reproduction of our second student's project, but at a
distinct time period. In it, we aim to detect a social media
event by examining tweet topic content over time. We use Latent Dirichlet
Allocation (LDA) models (Blei et al. 2003) to infer topics on three consecutive days centered
on Memorial Day 2020. We chose this example case study,
instead of the student projects, because of limited data
availability for the student projects. Despite this, the case study
illustrates the strategy and methods for one student
project. Below, we discuss case study design, tweet collection, and tweet structure,
before turning to quantitative methods for the case study.

\hypertarget{case-study-design}{%
\subsection{Case study design}\label{case-study-design}}

We sought to validate our hypothesis that we could detect a social
media event by examining tweet topic content at distinct time periods.
As a proof of principle of our event detection strategy,
we analyzed tweets before, during, and
after Memorial Day (May 25, 2020).
We fitted LDA models for each of three
distinct five-minute periods. The first period began at noon Eastern time
on May 24, 2020. Subsequent time periods started
24 and 48 hours later.
We defined each time period to be a single collection, or corpus, of
tweets.

\hypertarget{collecting-tweets-over-time}{%
\subsection{Collecting tweets over time}\label{collecting-tweets-over-time}}

We include here instructions for creating a tweet collection. First, we created a
new account on Twitter. With these user credentials, we used the
R package \texttt{rtweet} to query the API (Kearney 2019). We used the linux \texttt{crontab} software
to repeatedly execute R code to submit API queries. Each
query lasted five minutes and produced a text file of approximately 130 MB. We timed the API queries so
that there was no time lag between queries. We stored tweets resulting from API queries in their native JSON format.

Setting up the query task with \texttt{crontab} is straightforward. On our computer, with
Ubuntu 20.04 linux operating system, we opened a terminal and typed \texttt{crontab\ -e}. This
opened a text file containing user-specified tasks. We added the following line to the
bottom of the file before saving and closing the text file.

\begin{Shaded}
\begin{Highlighting}[]
\ExtensionTok{*/5}\NormalTok{ * * * * R -e }\StringTok{'rtweet::stream_tweets(timeout = (60 * 5), }
\StringTok{parse = FALSE, file_name = paste0("~/work/mentoring/mentoring-framework/data/",}
\StringTok{lubridate::now(), "-tweets"))'}
\end{Highlighting}
\end{Shaded}

Readers may need to slightly amend the above line to conform to requirements of
their operating system's software. Readers who use Mac OS may proceed as we did,
while those with Windows operating systems may consider using the R package
``taskscheduleR'' to schedule API queries via the Windows task scheduler (Wijffels and Belmans 2018).

\hypertarget{querying-twitter-api-to-get-complete-tweets}{%
\subsection{Querying Twitter API to get complete tweets}\label{querying-twitter-api-to-get-complete-tweets}}

Twitter API use agreements forbid users from sharing complete API query results.
However, Twitter permits users to share tweet identification numbers. With a tweet
identification number, a user may query a Twitter API to obtain complete tweet data. In our
experience, this
process is incomplete; that is, many tweet identification numbers submitted to the Twitter API return
no data. Additionally, some tweet identification numbers return data on the first query, but don't return data on subsequent queries. This complicates our goal of making all analyses computationally
reproducible and motivates our decision to share the tweet IDs of those tweets that
we actually analyzed (Supplementary files). Should a reader wish to reproduce
our analysis, we anticipate that she will get complete tweet data
for all or most of these tweet identification numbers from the API. We provide
R code for this task in the supplementary files.

\hypertarget{tweet-structure}{%
\subsection{Tweet structure}\label{tweet-structure}}

Tweets are available from the Twitter API as Javascript Object Notation (JSON) objects (``Introducing JSON'' 2020).
Every tweet consists
of multiple key-value pairs. The number of fields per
tweet depends on user settings, retweet status, and other factors (``Introduction to Tweet JSON'' 2020).
The 31 tweet key-value pairs belong to
12 distinct classes (Supplementary files). The
classes are either vectors - numeric, logical, or character - or arrays assembled
from the vector classes.

Below is an example of JSON for one tweet.

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{\{}
  \StringTok{"created_at"}\NormalTok{: }\StringTok{"Thu Apr 06 15:24:15 +0000 2017"}\NormalTok{,}
  \StringTok{"id_str"}\NormalTok{: }\StringTok{"850006245121695744"}\NormalTok{,}
  \StringTok{"text"}\NormalTok{: }\StringTok{"1\textbackslash{}/ Today we\textbackslash{}u2019re sharing our vision for the future of the Twitter API platform!"}\NormalTok{,}
  \StringTok{"user"}\NormalTok{: }\KeywordTok{\{}
    \StringTok{"id"}\NormalTok{: }\ExtensionTok{2244994945}\NormalTok{,}
    \StringTok{"name"}\NormalTok{: }\StringTok{"Twitter Dev"}\NormalTok{,}
    \StringTok{"screen_name"}\NormalTok{: }\StringTok{"TwitterDev"}\NormalTok{,}
    \StringTok{"location"}\NormalTok{: }\StringTok{"Internet"}\NormalTok{,}
    \StringTok{"url"}\NormalTok{: }\StringTok{"https:\textbackslash{}/\textbackslash{}/dev.twitter.com\textbackslash{}/"}\NormalTok{,}
    \StringTok{"description"}\NormalTok{: }\StringTok{"Your official source for Twitter Platform news, updates & events. }
\StringTok{    Need technical help? Visit https:\textbackslash{}/\textbackslash{}/twittercommunity.com\textbackslash{}/ \textbackslash{}u2328\textbackslash{}ufe0f }
\StringTok{    #TapIntoTwitter"}
  \KeywordTok{\}}\NormalTok{,}
  \StringTok{"place"}\NormalTok{: }\KeywordTok{\{}   
  \KeywordTok{\}}\NormalTok{,}
  \StringTok{"entities"}\NormalTok{: }\KeywordTok{\{}
    \StringTok{"hashtags"}\NormalTok{:}\BuiltInTok{ [}      
\NormalTok{    ],}
    \StringTok{"urls"}\NormalTok{: [}
\NormalTok{      \{}
        \StringTok{"url"}\NormalTok{: }\StringTok{"https:\textbackslash{}/\textbackslash{}/t.co\textbackslash{}/XweGngmxlP"}\NormalTok{,}
        \StringTok{"unwound"}\NormalTok{: \{}
          \StringTok{"url"}\NormalTok{: }\StringTok{"https:\textbackslash{}/\textbackslash{}/cards.twitter.com\textbackslash{}/cards\textbackslash{}/18ce53wgo4h\textbackslash{}/3xo1c"}\NormalTok{,}
          \StringTok{"title"}\NormalTok{: }\StringTok{"Building the Future of the Twitter API Platform"}
\NormalTok{        \}}
\NormalTok{      \}}
\NormalTok{    ],}
    \StringTok{"user_mentions"}\NormalTok{: [     }
\NormalTok{    ]}
\NormalTok{  \}}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

Our analyses use three fields from each tweet: date (``created\_at''), tweet identifier
(``id\_str''), and tweet text (``text''). The ``created\_at'' field is a character string containing
the date and time of the tweet. Every tweet has a unique identifier, the ``id\_str'' value. The
``text'' field contains the unicode representation of the message. After creating a text file
with tweet JSON, our next step involved reading and parsing tweets with the R
packages \texttt{rtweet} (Kearney 2019) and \texttt{tidytext} (Silge and Robinson 2016).

\hypertarget{parsing-tweet-text}{%
\subsection{Parsing tweet text}\label{parsing-tweet-text}}

The next task is to wrangle the tweet JSON into a data structure suitable for LDA modeling.
We used functions from the \texttt{rtweet} R package to parse tweet JSON into a data frame.
We then divided tweet text into words with functions from the \texttt{tidytext} R package. We
discarded commonly used ``stop words'' and emojis.

LDA model fitting requires that the corpus be organized as a
document by term matrix. In a document by term matrix, each row corresponds to a single document (a single
tweet), and each column is a single term (or word). Each cell contains a
count (the number of
occurrences of a term in the specified document). We created a document by
term matrix with the R function \texttt{cast\_dtm} from the \texttt{tidytext} package.

\hypertarget{latent-dirichlet-allocation}{%
\subsection{Latent Dirichlet Allocation}\label{latent-dirichlet-allocation}}

LDA is a statistical method for inferring latent (unobservable) topics (or themes)
from a large corpus (or collection) of documents (Blei et al. 2003).
We pretend that there's an imaginary process for creating documents in the corpus. For each document,
we choose a discrete distribution over topics. For example, some tweets from Memorial Day may
refer to the holiday. This may constitute one topic in the corpus.
Having chosen a distribution over topics, we then select document words by first
drawing a topic from the distribution over topics, then drawing a word from the
chosen topic.

In mathematical notation, we write the generative process assumed by LDA (Blei et al. 2003):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Choose \(N \sim \text{Poisson}(\xi)\)\\
\item
  Choose \(\theta \sim \text{Dirichlet}(\alpha)\)\\
\item
  For each word, \(w_n\) with \(n = 1, ..., N\),
\end{enumerate}

\begin{enumerate}
\def\labelenumi{\alph{enumi}.}
\tightlist
\item
  Choose a topic \(z_n \sim \text{Multinomial}(\theta)\)\\
\item
  Choose a word \(w_n\) from \(p(w_n | z_n, \beta)\), a multinomial probability
\end{enumerate}

\(\beta\) refers to the \(k\) by \(V\) matrix of topic-specific word probabilities, where \(k\)
is the number of topics and \(V\) is the size of the vocabulary, \emph{i.e.}, the number of unique words in the corpus.

The goal for LDA is to infer both the distribution over topics and
the topics (Blei et al. 2003). A topic, in this setting, is a distribution over
the vocabulary (the collection of all words in a corpus).
Inference for latent Dirichlet allocation models is performed by either
sampling from the posterior distribution or through variational methods. Researchers
have devised a variety of Gibbs sampling techniques for these models (Porteous et al. 2008).
Variational methods, while using approximations to the posterior distribution, offer the
advantage of computational speed (Blei et al. 2017). We used variational methods below in our case study.

\hypertarget{case-study-results}{%
\section{Case study results}\label{case-study-results}}

We identified the top ten most probable terms for each of ten topics in our models
(Figures \ref{fig:may24}, \ref{fig:may25}, \ref{fig:may26}).
We plotted the within-topic word probabilities as bar graphs. We see that topic-specific
word probabilities seldom exceed 0.05. We also note that some words are heavily weighted in
multiple topics. This observation complicates semantic topic interpretation. We
also caution that the results display expletives (that appeared on Twitter) and may
not be ``safe for work''.

\begin{figure}
\includegraphics[width=29.17in]{../results/beta-2020-05-24} \caption{Top terms for LDA model from May 24, 2020}\label{fig:may24}
\end{figure}

\begin{figure}
\includegraphics[width=29.17in]{../results/beta-2020-05-25} \caption{Top terms for LDA model from May 25, 2020 (Memorial Day)}\label{fig:may25}
\end{figure}

\begin{figure}
\includegraphics[width=29.17in]{../results/beta-2020-05-26} \caption{Top terms for LDA model from May 26, 2020}\label{fig:may26}
\end{figure}

Assigning meaning to topics is an active research area (Chang et al. 2009). Since
our interest is in the transient appearance of a new topic, we don't attempt to
assign meaning to every topic in our models. Instead, we anticipate that discussions on Twitter are a mixture of topics that endure over weeks or months and subjects that appear and disappear quickly. We see that topic 7 from May 25 has several
words that suggest Memorial Day: memorial, remember, honor, country. A similar topic is not
seen on May 24 or May 26. Some topics persist, with distinct word probabilities,
across the three days. For example, we see that President Trump features prominently in all
three models' results. On May 26, topic 10 reflects discussion of the Amy
Cooper Central Park incident (\url{https://www.nytimes.com/2020/05/26/nyregion/amy-cooper-dog-central-park.html}).

The murder of George Floyd occurred on May 25, 2020. Our last
examined time period, from 12:00 pm to 12:05 pm (Eastern USA time zone) on May 26, occurred after Floyd's murder,
yet we didn't detect this event in our ten-topic LDA model. Several considerations
may account for this. While outrage at the murder eventually spread worldwide, there may have been few
Floyd-related tweets during our collection time on May 26, less than 24 hours after the murder and video release.
Had we extended our analysis to May 27 and beyond, we may have identified George Floyd-related topics.

\hypertarget{assessment-of-learning-exploring-more-advanced-topics-and-concluding-remarks}{%
\section{Assessment of learning, exploring more advanced topics, and concluding remarks}\label{assessment-of-learning-exploring-more-advanced-topics-and-concluding-remarks}}

\hypertarget{assessment-of-learning}{%
\subsection{Assessment of learning}\label{assessment-of-learning}}

We examined student learning with both formative and summative assessments.
We conducted formative assessments through weekly discussions with students.
In these discussions, we developed action items to advance research progress and overcome
challenges. We summatively assessed student achievement at the end of the academic year.
Both students wrote a thesis and presented a poster to our statistics department.
We asked questions at the poster session to probe student understanding and critically
evaluated the theses.

\begin{table}

\caption{\label{tab:summative}Rubric for summative assessment of learning objectives.}
\centering
\fontsize{6}{8}\selectfont
\begin{tabular}[t]{>{\raggedright\arraybackslash}p{10em}|>{\raggedright\arraybackslash}p{10em}|>{\raggedright\arraybackslash}p{10em}|>{\raggedright\arraybackslash}p{10em}|>{\raggedright\arraybackslash}p{10em}}
\hline
Learning objective & Assessment item & 2 points & 1 point & 0 points\\
\hline
Write R code to perform text analysis of large volumes of tweets. & R code performs intended analyses & Code contains few or no bugs & Code contains one or more errors & Code contains many errors\\
\hline
Write R code to perform text analysis of large volumes of tweets. & Uses literate programming tools, such as Sweave or knitr & Report is written using literate programming tools. It compiles easily when run by instructor. Time-consuming calculations are cached. & Report is written using literate programming tools, but compilation takes too long or fails. & Report is not written with literate programming tools.\\
\hline
Write R code to perform text analysis of large volumes of tweets. & Uses git for version control & Log reveals regular commits with informative commit messages & Log reveals intermittent commits and uninformative commit messages & Doesn't use git.\\
\hline
Write R code to perform text analysis of large volumes of tweets. & Shares code and data via Github & Instructor easily clones repository from Github. Contains share-able data and instructions for getting other data to reproduce analysis. & One or more needed files is missing from repository. & Doesn't use Github.\\
\hline
Communicate results in a written report and poster presentation. & Organizes poster to highlight main points & When prompted, can describe main points in less than one minute. & Less fluid presentation with periods of silence or confusion. & Disorganized presentation.\\
\hline
Communicate results in a written report and poster presentation. & Accurately presents study and findings during poster session & Fluently describes background, study goals, study design, approach, data, findings, and conclusions & At least one section is incomplete or is verbal explanation is incomplete. & At least one section is missing.\\
\hline
Communicate results in a written report and poster presentation. & Report structure mirrors a research manuscript & Contains abstract, introduction, methods, results, and discussion & At least one section is incomplete. & At least one section is missing.\\
\hline
Translate statistical findings into scientific conclusions. & Places statistical results in their scientific context & Demonstrates understanding of scientific context and integrates findings into it. & Incomplete scientific understanding or incomplete integration of findings. & Major gaps in scientific understanding or integration of findings.\\
\hline
Translate statistical findings into scientific conclusions. & Accurately portrays study limitations & Accurately describes, in writing and in speaking, limitations of the study & Incomplete or partially inaccurate description of limitations & Doesn't describe limitations.\\
\hline
Translate statistical findings into scientific conclusions. & Demonstrates familiarity with relevant literature & Fluent in both relevant data science literature and scientific literature. & Incomplete knowledge and understanding of relevant literature & Major gaps in knowledge and understanding\\
\hline
Develop data science strategies to address a scientific research question. & Presents an original research question & Presents, in writing and in speaking, a novel research question. Explains why it's novel, too. & Partially lacking in elements of question's background or novelty. & Doesn't present an original question.\\
\hline
Develop data science strategies to address a scientific research question. & Effectively uses data visualizations & Visualizations highlight main points of report. & Incomplete or omitted visualizations. & Doesn't use visualizations.\\
\hline
Develop data science strategies to address a scientific research question. & Presents accurate scientific conclusions & Effectively translates analysis results into their scientific context. & Minor inaccuracy in translation of findings into scientific context. & Major errors in translation of results.\\
\hline
\end{tabular}
\end{table}

With future students, we will use a written rubric to evaluate
theses (Table \ref{tab:summative}). We'll
share the rubric with our students at the start of the academic year. With only minor
modifications, the rubric may be suitable for projects that don't use tweets.

\hypertarget{exploring-more-advanced-topics}{%
\subsection{Exploring more advanced topics}\label{exploring-more-advanced-topics}}

Twitter data over time inspires a variety of research projects. Supplementing
tweets with public data from other sources multiplies the possibilities. For example,
one of our two students supplemented tweets with daily stock market index prices.
She studied sentiment of finance-related tweets and daily stock market index
closing prices (Supplementary files).

LDA modeling and related methods are a major research area in the
quantitative social sciences. Advanced students with interest in statistical computing might
compare inferential methods for topic models. Those with interests in event detection
and time series analysis could build on the findings of our student by explicitly accounting
for topic evolution with dynamic topic models (Blei and Lafferty 2006).

\hypertarget{concluding-remarks}{%
\subsection{Concluding remarks}\label{concluding-remarks}}

Our mentoring in data science aligns with others' calls to reconsider
the role of computing in statistics
and data science (Carver et al. 2016; Nolan and Temple Lang 2010). Hicks and Irizarry (2018) argue for
incorporating three concepts into data science training:
computing, connecting and creating.
They use the terms ``connecting'' and ``creating'' to describe the processes of applying
quantitative methods to real data and research questions and of formulating research
questions, respectively. Our tweet analysis projects offer students opportunities in all
three skills sets. Our students first formulated research questions, then collected and
analyzed data to address the questions. Throughout the projects, students
drew heavily on computing, both to acquire data and to analyze it.

Tweet analysis gives students practical experience in the data science process of
formulating a research question, gathering data to address it, summarizing the
data, visualizing results, and communicating findings. Tweets over time are a rich, large,
authentic data set that offers many opportunities.
We provided instructions to enable readers to establish their own tweet collections.
We also presented details for one analysis strategy.
By considering first student research interests and integrating them with our
senior thesis learning objectives, we successfully guided two undergraduate
researchers in data science research with tweets.

\hypertarget{acknowledgements}{%
\section{Acknowledgements}\label{acknowledgements}}

The authors thank Betsy Colby Davie and Rick Nordheim for helpful discussions and
feedback on preliminary versions of the manuscript. We thank the special issue
editors and anonymous reviewers for their constructive comments and suggestions. Finally, this work wouldn't have been possible without the keen and enthusiastic students, Jinyu Xia and Robert Turner.

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-blei2017variational}{}%
Blei, D. M., Kucukelbir, A., and McAuliffe, J. D. (2017), ``Variational inference: A review for statisticians,'' \emph{Journal of the American Statistical Association}, Taylor \& Francis, 112, 859--877.

\leavevmode\hypertarget{ref-blei2006dynamic}{}%
Blei, D. M., and Lafferty, J. D. (2006), ``Dynamic topic models,'' in \emph{Proceedings of the 23rd International Conference on Machine Learning}, pp. 113--120.

\leavevmode\hypertarget{ref-blei2003latent}{}%
Blei, D. M., Ng, A. Y., and Jordan, M. I. (2003), ``Latent Dirichlet Allocation,'' \emph{Journal of Machine Learning Research}, 3, 993--1022.

\leavevmode\hypertarget{ref-box1976science}{}%
Box, G. E. (1976), ``Science and statistics,'' \emph{Journal of the American Statistical Association}, Taylor \& Francis, 71, 791--799.

\leavevmode\hypertarget{ref-carver2016guidelines}{}%
Carver, R., Everson, M., Gabrosek, J., Horton, N., Lock, R., Mocko, M., Rossman, A., Roswell, G. H., Velleman, P., Witmer, J., and others (2016), ``Guidelines for assessment and instruction in statistics education (GAISE) college report 2016,'' AMSTAT.

\leavevmode\hypertarget{ref-chang2009reading}{}%
Chang, J., Gerrish, S., Wang, C., Boyd-Graber, J. L., and Blei, D. M. (2009), ``Reading tea leaves: How humans interpret topic models,'' in \emph{Advances in Neural Information Processing Systems}, pp. 288--296.

\leavevmode\hypertarget{ref-hicks2018guide}{}%
Hicks, S. C., and Irizarry, R. A. (2018), ``A guide to teaching data science,'' \emph{The American Statistician}, Taylor \& Francis, 72, 382--391.

\leavevmode\hypertarget{ref-json}{}%
``Introducing JSON'' (2020), \url{https://www.json.org/json-en.html}.

\leavevmode\hypertarget{ref-tweet_json}{}%
``Introduction to Tweet JSON'' (2020), \url{https://developer.twitter.com/en/docs/tweets/data-dictionary/overview/intro-to-tweet-json}.

\leavevmode\hypertarget{ref-rtweet-package}{}%
Kearney, M. W. (2019), ``rtweet: Collecting and analyzing Twitter data,'' \emph{Journal of Open Source Software}, 4, 1829. \url{https://doi.org/10.21105/joss.01829}.

\leavevmode\hypertarget{ref-lin2011joint}{}%
Lin, C. X., Mei, Q., Han, J., Jiang, Y., and Danilevsky, M. (2011), ``The joint inference of topic diffusion and evolution in social communities,'' in \emph{2011 IEEE 11th International Conference on Data Mining}, IEEE, pp. 378--387.

\leavevmode\hypertarget{ref-national2018data}{}%
National Academies of Sciences, Engineering, Medicine, and others (2018), \emph{Data science for undergraduates: Opportunities and options}, National Academies Press.

\leavevmode\hypertarget{ref-nolan2010computing}{}%
Nolan, D., and Temple Lang, D. (2010), ``Computing in the statistics curricula,'' \emph{The American Statistician}, Taylor \& Francis, 64, 97--107.

\leavevmode\hypertarget{ref-pelled2018little}{}%
Pelled, A., Lukito, J., Boehm, F., Yang, J., and Shah, D. (2018), ```Little Marco,'`Lyin'Ted,'`Crooked Hillary,' and the `Biased' media: How Trump used Twitter to attack and organize,'' in \emph{Digital Discussions}, Routledge, pp. 176--196.

\leavevmode\hypertarget{ref-porteous2008fast}{}%
Porteous, I., Newman, D., Ihler, A., Asuncion, A., Smyth, P., and Welling, M. (2008), ``Fast Collapsed Gibbs Sampling for Latent Dirichlet Allocation,'' in \emph{Proceedings of the 14th ACM SIGKDD International Conference on Knowledge Discovery and Data Mining}, pp. 569--577.

\leavevmode\hypertarget{ref-r}{}%
R Core Team (2019), \emph{R: A language and environment for statistical computing}, Vienna, Austria: R Foundation for Statistical Computing.

\leavevmode\hypertarget{ref-drob}{}%
Robinson, D. (2016), ``Text analysis of Trump's tweets confirms he writes only the (angrier) Android half,'' \url{http://varianceexplained.org/r/trump-tweets/}.

\leavevmode\hypertarget{ref-tweet_stream}{}%
``Sampled stream'' (2019), \url{https://developer.twitter.com/en/docs/labs/sampled-stream/overview}.

\leavevmode\hypertarget{ref-tidytext}{}%
Silge, J., and Robinson, D. (2016), ``tidytext: Text mining and analysis using tidy data principles in R,'' \emph{JOSS}, The Open Journal, 1. \url{https://doi.org/10.21105/joss.00037}.

\leavevmode\hypertarget{ref-wells2016trump}{}%
Wells, C., Shah, D. V., Pevehouse, J. C., Yang, J., Pelled, A., Boehm, F., Lukito, J., Ghosh, S., and Schmidt, J. L. (2016), ``How Trump drove coverage to the nomination: Hybrid media campaigning,'' \emph{Political Communication}, Taylor \& Francis, 33, 669--676.

\leavevmode\hypertarget{ref-wickham2019advanced}{}%
Wickham, H. (2019), \emph{Advanced R}, CRC press.

\leavevmode\hypertarget{ref-wickham2016r}{}%
Wickham, H., and Grolemund, G. (2016), \emph{R for Data Science: Import, tidy, transform, visualize, and model data}, O'Reilly Media, Inc.

\leavevmode\hypertarget{ref-wiggins2005understanding}{}%
Wiggins, G., and McTighe, J. (2005), \emph{Understanding by Design}.

\leavevmode\hypertarget{ref-taskscheduleR}{}%
Wijffels, J., and Belmans, O. (2018), \emph{taskscheduleR: Schedule R Scripts and Processes with the Windows Task Scheduler}.

\newpage

\hypertarget{supplementary-files}{%
\section{Supplementary files}\label{supplementary-files}}

\hypertarget{tweets-data-dictionary}{%
\subsection{Tweets data dictionary}\label{tweets-data-dictionary}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \href{https://github.com/fboehm/jse-2019/blob/master/data/tweets-data-dictionary.csv}{Data dictionary}
\end{enumerate}

\hypertarget{r-code-to-reproduce-the-case-study}{%
\subsection{R code to reproduce the case study}\label{r-code-to-reproduce-the-case-study}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \href{https://raw.githubusercontent.com/fboehm/jse-2019/master/Rmd/tweets.Rmd}{tweets.Rmd}
\item
  \href{https://raw.githubusercontent.com/fboehm/jse-2019/master/Rmd/tweets-one.Rmd}{tweets-one.Rmd}
\item
  \href{https://raw.githubusercontent.com/fboehm/jse-2019/master/Rmd/recover_tweets.R}{recover\_tweets.R}
\end{enumerate}

\hypertarget{student-projects}{%
\subsection{Student projects}\label{student-projects}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Student 1 poster: \href{https://github.com/fboehm/jse-2019/blob/master/supplementary/Project_Poster.pdf}{Project\_Poster.pdf}
\item
  Student 1 report: \href{https://github.com/fboehm/jse-2019/blob/master/supplementary/report.pdf}{report.pdf}
\item
  Student 2 useR 2016 slides: \href{https://github.com/fboehm/jse-2019/blob/master/supplementary/user2016boehm.pdf}{user2016boehm.pdf}
\item
  Student 2 poster: \href{https://github.com/fboehm/jse-2019/blob/master/supplementary/warfdiscovery2016boehm.tiff}{warfdiscovery2016boehm.tiff}
\end{enumerate}

\hypertarget{github-repository}{%
\subsection{Github repository}\label{github-repository}}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \url{https://github.com/fboehm/jse-2019}
\end{enumerate}

\end{document}

### 2020-10-08 october2020_uploads/zip_supplemental/Rmd/tweets.Rmd
added uploads for Oct 2020 submission

--- 2020-05-29 Rmd/tweets.Rmd
@@
 title: "Tweets"
-author: "Frederick J. Boehm"
+author: ""
 date: "4/9/2020"
@@
 
-Install `rtweet` package from CRAN.
+Install R packages from CRAN.
 
 ```{r, eval = FALSE}
-install.packages("rtweet")
-install.packages("lubridate")
+install.packages(c("rtweet", "lubridate", 
+                   "magrittr", "tidytext", 
+                   "qdapDictionaries", "purrr", "knitr", 
+                   "devtools", "git2r")
+                 )
 ```

### 2020-11-02 Rmd/jse-special-issue-2020-november.Rmd
added fixes for November 2020 typos

--- 2020-10-07 Rmd/jse-special-issue-2020-october.Rmd
@@
 Some social media data, including tweets from Twitter, is
-available through website application product interfaces (APIs). 
+available through website application programming interfaces (APIs). 
 By way of a streaming API, Twitter shares a sample of approximately 
@@
 readr::read_csv("../data/concepts-backward-design.csv") %>%
+  dplyr::arrange(Category) %>%
   knitr::kable(caption = "Classifying project skills") %>%
@@
 Her sentiment score reflected presence of emotion-associated terms 
-(*eg*., "happy", "sad", "mad", "scared") 
+(*e.g*., "happy", "sad", "mad", "scared") 
 in tweet texts. Days with more net positive emotion words in the collected tweets received a higher 
@@
 topics would be short-lived in the continuous Twitter stream. We discovered
-evidence to support his strategy, and we ultimately presented our findings 
-at international and local research meetings. We share below a case study based
-on this approach.
+evidence to support his ideas, and we ultimately presented our findings 
+at international and local research meetings. Below, we share a case study on a different, widely discussed topic which is analyzed using an approach similar to that from the Super Bowl tweets.
 
@@
 
-Below is an example of JSON for one tweet.
+Below is an example of a tweet in JSON format.
 
-```{bash, eval = FALSE, fig.cap = "JSON example for one tweet", echo = TRUE}
+```{bash, eval = FALSE, fig.cap = "JSON format for one tweet", echo = TRUE}
 {
@@
 LDA model fitting requires that the corpus be organized as a 
-document by term matrix. In a document by term matrix, each row corresponds to a single document (a single 
+document-term matrix. In a document-term matrix, each row corresponds to a single document (a single 
 tweet), and each column is a single term (or word). Each cell contains a 
 count (the number of 
-occurrences of a term in the specified document). We created a document by 
-term matrix with the R function `cast_dtm` from the `tidytext` package.
+occurrences of a term in the specified document). We created a document-term matrix with the R function `cast_dtm` from the `tidytext` package.
 
@@
 multiple topics. This observation complicates semantic topic interpretation. We 
-also caution that the results display expletives (that appeared on Twitter) and may 
-not be "safe for work".
+also caution that the results display expletives (that appeared on Twitter) and may be "not suitable for work (NSFW)".
 
@@
 
-```{r, may25, fig.cap = "Top terms for LDA model from May 25, 2020 (Memorial Day)"}
+```{r, may25, fig.cap = "Top terms for LDA model from May 25, 2020 (Memorial Day). Results contain expletives and may be 'not suitable for work (NSFW)'."}
 knitr::include_graphics("../results/beta-2020-05-25.png")

### 2020-11-02 Rmd/jse-special-issue-2020-october-anon.Rmd
added fixes for November 2020 typos

--- 2020-10-06 Rmd/jse-special-issue-2020-october.Rmd
@@
 title: "What is happening on Twitter? A framework for student research projects with tweets"
-author: "Frederick J. Boehm and Bret M. Hanlon"
+author: ""
 date: ""

### 2020-11-02 Rmd/reply-to-reviewers-nov2020.Rmd
updated ms AND reply to comments. Updated pdf of ms

---
output: 
  bookdown::pdf_document2:
    toc: false
    number_sections: false
---

November 2, 2020

Dear editors and reviewers:

We are writing to submit an updated manuscript that incorporates the suggested edits.

We include both the text of the comments and our replies.

## Editor's Comments to Author:

* unfortunately, I don't think JSE has formatting guidelines for packages. However, they should be internally consistent in
your paper (e.g., bold, quoted, without format, etc.): rtweet, tidytext, lubridate,...

> We now do this for all packages that we discuss in the text.

* API stands for application programming interface (not application product interface)

> Thank you for this correction. We've made the change.

* how was table 1 ordered? As is, the sorting is (seemingly?) arbitrary, and I think it would be easier to read if sorted. I’d
suggest one of two ways:
- sort by order of category
- sort by the order in which the skills are performed in a standard data analysis.

> Thank you. We now order the rows by category.

* page 10, line 12 “eg., “ should be “e.g.,”

> We've now made this change.

* page 10, line 32. You discovered evidence to “support his strategy” ... what is his strategy? Do you mean to “support his
hypothesis”? or maybe to “support his ideas” ?

> Thank you. We decided to use "support his ideas".

* page 10, line 34 “We share below a case study based on this approach.” I’d clarify to say “Below, we share a case study
on a different widely discussed topic which is analyzed using an approach similar to that from the Super Bowl tweets.”

> Thanks. We've made this change.

* page 12, line 32 ... still something is off as I read the word JSON. You write: “Below is an example of JSON for one
tweet.” Here are some possible alternatives:
- “Below is an example of a JSON entry for one tweet.”
- “Below is an example of a JSON file for one tweet.”
- “Below is an example of a tweet in JSON format.”

> Thank you. We've chosen to use the last of the three suggested wordings.

* “to wrangle the tweet JSON into” (instead change to: “to wrangle the JSON tweet into” or “to wrangle the tweet JSON
data into”) page 13, line 46

> Thanks. We've made this change by using the latter suggested wording.

* I think “document by term matrix” should be “document-term matrix”

> We agree. We've made the changes.

* page 10: suggest changing 'may not be "safe for work"' to "may be "not suitable for work (NSFW)" and add a similar
warning to the Figure 1 caption. Would an additional sentence noting that instructors could have students remove words
from a list of common expletives or inappropriate words?

> Thanks. We've fixed this wording.

* Would it make sense to have all of the supplementary files available on a new public GitHub repo? That would simplify
references and updates.

> Yes, all files are in a Github repo at https://github.com/fboehm/jse-2019. If an additional (new) repo is needed, we're happy to make one.

### 2020-11-02 Rmd/jse-special-issue-2020-november.Rmd
updated ms AND reply to comments. Updated pdf of ms

@@
 while those with Windows operating systems may consider using the R package 
-"taskscheduleR" to schedule API queries via the Windows task scheduler [@taskscheduleR].
+`taskscheduleR` to schedule API queries via the Windows task scheduler [@taskscheduleR].
 
@@
 
-The next task is to wrangle the tweet JSON into a data structure suitable for LDA modeling. 
+The next task is to wrangle the tweet JSON data into a structure suitable for LDA modeling. 
 We used functions from the `rtweet` R package to parse tweet JSON into a data frame. 
@@
 
-```{r, may24, fig.cap = "Top terms for LDA model from May 24, 2020"}
+```{r, may24, fig.cap = 'Top terms for LDA model from May 24, 2020. Results contain expletives and may be "not suitable for work (NSFW)".'}
 knitr::include_graphics("../results/beta-2020-05-24.png")
@@
 
-```{r, may25, fig.cap = "Top terms for LDA model from May 25, 2020 (Memorial Day). Results contain expletives and may be 'not suitable for work (NSFW)'."}
+```{r, may25, fig.cap = "Top terms for LDA model from May 25, 2020 (Memorial Day)."}
 knitr::include_graphics("../results/beta-2020-05-25.png")

### 2020-11-02 Rmd/jse-special-issue-2020-november.tex
updated ms AND reply to comments. Updated pdf of ms

--- 2020-10-07 Rmd/jse-special-issue-2020-october.tex
@@
 \makesavenoteenv{longtable}
-\usepackage{graphicx,grffile}
+\usepackage{graphicx}
 \makeatletter
@@
 \usepackage{xcolor}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newenvironment{cslreferences}%
+  {\setlength{\parindent}{0pt}%
+  \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces}%
+  {\par}
 
@@
 Some social media data, including tweets from Twitter, is
-available through website application product interfaces (APIs).
+available through website application programming interfaces (APIs).
 By way of a streaming API, Twitter shares a sample of approximately
@@
 \hline
-Structure research project files as R package & Worth being familiar with\\
-\hline
 Communicate results in speaking and in writing & Enduring understanding\\
@@
 \hline
+Use text analysis tools to analyze tweets & Enduring understanding\\
+\hline
+Translate analysis results into scientific conclusions & Enduring understanding\\
+\hline
+Describe assumptions and limitations of statistical analyses & Enduring understanding\\
+\hline
 Use Github to share code and documentation & Important to know and do\\
@@
 \hline
-Use text analysis tools to analyze tweets & Enduring understanding\\
-\hline
-Use cluster computing as needed & Worth being familiar with\\
-\hline
 Use data visualization to clarify and inform quantitative analyses & Important to know and do\\
-\hline
-Translate analysis results into scientific conclusions & Enduring understanding\\
 \hline
@@
 \hline
-Describe assumptions and limitations of statistical analyses & Enduring understanding\\
+Structure research project files as R package & Worth being familiar with\\
+\hline
+Use cluster computing as needed & Worth being familiar with\\
 \hline
@@
 Her sentiment score reflected presence of emotion-associated terms
-(\emph{eg}., ``happy'', ``sad'', ``mad'', ``scared'')
+(\emph{e.g}., ``happy'', ``sad'', ``mad'', ``scared'')
 in tweet texts. Days with more net positive emotion words in the collected tweets received a higher
@@
 topics would be short-lived in the continuous Twitter stream. We discovered
-evidence to support his strategy, and we ultimately presented our findings
-at international and local research meetings. We share below a case study based
-on this approach.
+evidence to support his ideas, and we ultimately presented our findings
+at international and local research meetings. Below, we share a case study on a different, widely discussed topic which is analyzed using an approach similar to that from the Super Bowl tweets.
 
@@
 \begin{Highlighting}[]
-\ExtensionTok{*/5}\NormalTok{ * * * * R -e }\StringTok{'rtweet::stream_tweets(timeout = (60 * 5), }
-\StringTok{parse = FALSE, file_name = paste0("~/work/mentoring/mentoring-framework/data/",}
-\StringTok{lubridate::now(), "-tweets"))'}
+\ExtensionTok{*/5}\NormalTok{ * * * * R {-}e }\StringTok{\textquotesingle{}rtweet::stream\_tweets(timeout = (60 * 5), }
+\StringTok{parse = FALSE, file\_name = paste0("\textasciitilde{}/work/mentoring/mentoring{-}framework/data/",}
+\StringTok{lubridate::now(), "{-}tweets"))\textquotesingle{}}
 \end{Highlighting}
@@
 while those with Windows operating systems may consider using the R package
-``taskscheduleR'' to schedule API queries via the Windows task scheduler (Wijffels and Belmans 2018).
+\texttt{taskscheduleR} to schedule API queries via the Windows task scheduler (Wijffels and Belmans 2018).
 
@@
 
-Below is an example of JSON for one tweet.
+Below is an example of a tweet in JSON format.
 
@@
 \KeywordTok{\{}
-  \StringTok{"created_at"}\NormalTok{: }\StringTok{"Thu Apr 06 15:24:15 +0000 2017"}\NormalTok{,}
-  \StringTok{"id_str"}\NormalTok{: }\StringTok{"850006245121695744"}\NormalTok{,}
+  \StringTok{"created\_at"}\NormalTok{: }\StringTok{"Thu Apr 06 15:24:15 +0000 2017"}\NormalTok{,}
+  \StringTok{"id\_str"}\NormalTok{: }\StringTok{"850006245121695744"}\NormalTok{,}
   \StringTok{"text"}\NormalTok{: }\StringTok{"1\textbackslash{}/ Today we\textbackslash{}u2019re sharing our vision for the future of the Twitter API platform!"}\NormalTok{,}
@@
     \StringTok{"name"}\NormalTok{: }\StringTok{"Twitter Dev"}\NormalTok{,}
-    \StringTok{"screen_name"}\NormalTok{: }\StringTok{"TwitterDev"}\NormalTok{,}
+    \StringTok{"screen\_name"}\NormalTok{: }\StringTok{"TwitterDev"}\NormalTok{,}
     \StringTok{"location"}\NormalTok{: }\StringTok{"Internet"}\NormalTok{,}
     \StringTok{"url"}\NormalTok{: }\StringTok{"https:\textbackslash{}/\textbackslash{}/dev.twitter.com\textbackslash{}/"}\NormalTok{,}
-    \StringTok{"description"}\NormalTok{: }\StringTok{"Your official source for Twitter Platform news, updates & events. }
+    \StringTok{"description"}\NormalTok{: }\StringTok{"Your official source for Twitter Platform news, updates \& events. }
 \StringTok{    Need technical help? Visit https:\textbackslash{}/\textbackslash{}/twittercommunity.com\textbackslash{}/ \textbackslash{}u2328\textbackslash{}ufe0f }
-\StringTok{    #TapIntoTwitter"}
+\StringTok{    \#TapIntoTwitter"}
   \KeywordTok{\}}\NormalTok{,}
@@
 \NormalTok{    ],}
-    \StringTok{"user_mentions"}\NormalTok{: [     }
+    \StringTok{"user\_mentions"}\NormalTok{: [     }
 \NormalTok{    ]}
@@
 
-The next task is to wrangle the tweet JSON into a data structure suitable for LDA modeling.
+The next task is to wrangle the tweet JSON data into a structure suitable for LDA modeling.
 We used functions from the \texttt{rtweet} R package to parse tweet JSON into a data frame.
@@
 LDA model fitting requires that the corpus be organized as a
-document by term matrix. In a document by term matrix, each row corresponds to a single document (a single
+document-term matrix. In a document-term matrix, each row corresponds to a single document (a single
 tweet), and each column is a single term (or word). Each cell contains a
 count (the number of
-occurrences of a term in the specified document). We created a document by
-term matrix with the R function \texttt{cast\_dtm} from the \texttt{tidytext} package.
+occurrences of a term in the specified document). We created a document-term matrix with the R function \texttt{cast\_dtm} from the \texttt{tidytext} package.
 
@@
 multiple topics. This observation complicates semantic topic interpretation. We
-also caution that the results display expletives (that appeared on Twitter) and may
-not be ``safe for work''.
+also caution that the results display expletives (that appeared on Twitter) and may be ``not suitable for work (NSFW)''.
 
 \begin{figure}
-\includegraphics[width=29.17in]{../results/beta-2020-05-24} \caption{Top terms for LDA model from May 24, 2020}\label{fig:may24}
+\includegraphics[width=29.17in]{../results/beta-2020-05-24} \caption{Top terms for LDA model from May 24, 2020. Results contain expletives and may be "not suitable for work (NSFW)".}\label{fig:may24}
 \end{figure}
@@
 \begin{figure}
-\includegraphics[width=29.17in]{../results/beta-2020-05-25} \caption{Top terms for LDA model from May 25, 2020 (Memorial Day)}\label{fig:may25}
+\includegraphics[width=29.17in]{../results/beta-2020-05-25} \caption{Top terms for LDA model from May 25, 2020 (Memorial Day).}\label{fig:may25}
 \end{figure}
@@
 \hypertarget{refs}{}
+\begin{cslreferences}
 \leavevmode\hypertarget{ref-blei2017variational}{}%
@@
 Pelled, A., Lukito, J., Boehm, F., Yang, J., and Shah, D.
-(2018), ```Little Marco,'`Lyin'Ted,'`Crooked Hillary,' and the `Biased' media: How Trump used Twitter to attack and organize,'' in \emph{Digital Discussions}, Routledge, pp.
+(2018), ```Little Marco,'\,`Lyin'Ted,'\,`Crooked Hillary,' and the `Biased' media: How Trump used Twitter to attack and organize,'' in \emph{Digital Discussions}, Routledge, pp.
 176--196.
@@
 Wijffels, J., and Belmans, O. (2018), \emph{taskscheduleR: Schedule R Scripts and Processes with the Windows Task Scheduler}.
+\end{cslreferences}
 

### 2020-11-02 Rmd/jse-special-issue-2020-november.Rmd
added requested sentence.

@@
 also caution that the results display expletives (that appeared on Twitter) and may be "not suitable for work (NSFW)". 
+Instructors may apply a filter to remove common expletives before LDA modeling of tweets.
 
-```{r, may24, fig.cap = 'Top terms for LDA model from May 24, 2020. Results contain expletives and may be "not suitable for work (NSFW)".'}
+```{r, may24, fig.cap = 'Top terms for LDA model from May 24, 2020. Results contain expletives and may be not suitable for work (NSFW).'}
 knitr::include_graphics("../results/beta-2020-05-24.png")

### 2020-11-03 Rmd/jse-special-issue-2020-november-anon.Rmd
updates prior to resubmission

--- 2020-11-02 Rmd/jse-special-issue-2020-november.Rmd
@@
 title: "What is happening on Twitter? A framework for student research projects with tweets"
-author: "Frederick J. Boehm and Bret M. Hanlon"
 date: ""

### 2020-11-29 november2020_uploads/zip_supplemental/recover_tweets.R
added november 2020 submission

@@ -1,0 +1,67 @@
+#' Recovers Twitter damaged stream data (JSON file) into parsed data frame.
+#'
+#' @param path Character, name of JSON file with data collected by
+#'   \code{\link{stream_tweets}}.
+#' @param dir Character, name of a directory where intermediate files are
+#'   stored.
+#' @param verbose Logical, should progress be displayed?
+#' 
+#' @family stream tweets
+# https://gist.githubusercontent.com/JBGruber/dee4c44e7d38d537426f57ba1e4f84ab/raw/ab87bebb8d020c2f96c71a40a483dc96a4c80e54/recover_stream.R
+recover_stream <- function(path, dir = NULL, verbose = TRUE) {
+  
+  # read file and split to tweets
+  lines <- readChar(path, file.info(path)$size, useBytes = TRUE)
+  tweets <- stringi::stri_split_fixed(lines, "\n{")[[1]]
+  tweets[-1] <- paste0("{", tweets[-1])
+  tweets <- tweets[!(tweets == "" | tweets == "{")]
+  
+  # remove misbehaving characters
+  tweets <- gsub("\r", "", tweets, fixed = TRUE)
+  tweets <- gsub("\n", "", tweets, fixed = TRUE)
+  
+  # write tweets to disk and try to read them in individually
+  if (is.null(dir)) {
+    dir <- paste0(tempdir(), "/tweets/")
+    dir.create(dir, showWarnings = FALSE)
+  }
+  
+  if (verbose) {
+    pb <- progress::progress_bar$new(
+      format = "Processing tweets [:bar] :percent, :eta remaining",
+      total = length(tweets), clear = FALSE
+    )
+    pb$tick(0)
+  }
+  
+  tweets_l <- lapply(tweets, function(t) {
+    pb$tick()
+    id <- unlist(stringi::stri_extract_first_regex(t, "(?<=id\":)\\d+(?=,)"))[1]
+    f <- paste0(dir, id, ".json")
+    writeLines(t, f, useBytes = TRUE)
+    out <- tryCatch(rtweet::parse_stream(f),
+                    error = function(e) {})
+    if ("tbl_df" %in% class(out)) {
+      return(out)
+    } else {
+      return(id)
+    }
+  })
+  
+  # test which ones failed
+  test <- vapply(tweets_l, is.character, FUN.VALUE = logical(1L))
+  bad_files <- unlist(tweets_l[test])
+  
+  # Let user decide what to do
[…]
